### 2013-06-10 paper/jneurosci2013.tex
First version of the paper, with figures.

%\documentclass[twocolumn]{article}
\documentclass[12pt]{article}
% Change "article" to "report" to get rid of page number on title page
\usepackage{amsmath,amsfonts,amsthm,amssymb}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{extramarks}
\usepackage{chngpage}
\usepackage{soul}
\usepackage[usenames,dvipsnames]{color}
\usepackage{graphicx,float,wrapfig}
\usepackage{listings}
\usepackage{framed}
\usepackage{inconsolata}
\usepackage{cite}

% Stuff to change, you know, if you want.
\setlength{\parindent}{1cm}
\setlength{\parskip}{3pt}

% In case you need to adjust margins:
\topmargin=-0.45in      %
\evensidemargin=0in     %
\oddsidemargin=0in      %
\textwidth=6.5in        %
\textheight=9.0in       %
\headsep=0.25in         %

\pagestyle{empty}

\newcommand{\framedscalefigone}[3]{
  \begin{figure}[ht!]
    % Requires \usepackage{graphicx}
    \centering
    \begin{framed}
      \begin{minipage}{\textwidth}
        \centering
        \includegraphics[width=#2\columnwidth]{#1}
        \caption{#3}
        \label{#1}
      \end{minipage}
    \end{framed}
  \end{figure}}

\newcommand{\framedscalefigtwo}[4]{
  \begin{figure}[ht!]
    % Requires \usepackage{graphicx}
    \centering
    \begin{framed}
      \begin{minipage}{0.49\textwidth}
        \includegraphics[width=#3\columnwidth]{#1}
      \end{minipage}
      \begin{minipage}{0.49\textwidth}
        \includegraphics[width=#3\columnwidth]{#2}
      \end{minipage}
      \caption{#4}
      \label{#1}
    \end{framed}
  \end{figure}}

\newcommand{\scalefigone}[3]{
  \begin{figure}[ht!]
    % Requires \usepackage{graphicx}
    \centering
    \includegraphics[width=#2\columnwidth]{#1}
    \caption{#3}
    \label{#1}
  \end{figure}}

\newcommand{\scalefigtwo}[4]{
  \begin{figure}[ht!]
    % Requires \usepackage{graphicx}
    \centering
    \includegraphics[width=#3\columnwidth]{#1}
    \includegraphics[width=#3\columnwidth]{#2}
    \caption{#4}
    \label{#1}
  \end{figure}}

\setlength\fboxrule{1pt}

\graphicspath{{../figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Make title
\title{}
\date{}
\author{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{center} \begin{Large}
  \textbf{A spiking neural circuit model of cortical function
          in simple reaction time tasks}
\end{Large} \end{center}

\begin{center}
  Trevor Bekolay, Benjamine Liu, Nandakumar Narayanan, Chris
  Eliasmith, Mark Laubach
\end{center}

\begin{abstract}
  todo
\end{abstract}

\section{Introduction}

Animals make complicated motor actions
with high spatial and temporal resolutions.
In temporally precise tasks,
the motor cortex and several afferents coordinate
to bring about motor actions at appropriate times.
For tasks with delay periods,
the medial prefrontal cortex is critically important
for suppressing motor action during the delay period.
This effect has been seen in ?.  %???primates humans rodents etc???.

In rodents, the role of the medial prefrontal cortex
has been investigated through ?.
%???previous work???.
Previously, we have used a simple reaction task
and principal component analysis
to investigate ?. %???summarize Narayana \& laubach jnp 2009???
These results point to neural integration
as playing a key role
in the mPFC's ability
to control the timing of motor actions.

In order to investigate the neural circuit
underlying the production and suppression of motor output,
we extend Singh \& Eliasmith's model to include
neural populations that directly affect the environment,
analogous to the function of Betz cells.
In doing so, we can relate the model to the experimental data
in both neural and behavioural contexts.

\section{Model}

The model presented in this study
is based on a conceptual model of PFC and MC
function in delay tasks proposed by Narayanan and Laubach.
Their conceptual model is based on
principal component analysis of neural data
in rats performing a simple reaction time task.

Figure 1 summarizes their PC analysis.
In PFC, two components account for over 35\%
of variance in population firing.
The first principal component increases slightly
at the time of the lever press,
and then decreases over the delay period.
The second principal component
also increases at the time of lever press,
but remains high until the time of release.

In motor cortex, three components account
for over 60\% of variance in population firing.
The first two principal components are
similar to those found in PFC;
the first steadily decreases over the delay period,
and the second remains high during the delay period.
The third principal component increases
sharply slightly before lever press
and lever release.

The identification of these principal components
leads to a conceptual model of cortical function
during simple reaction time tasks.
This model is summarized in figure 2. % Fig
At the start of a trial,
two complementary systems are activated
to produce quick reactions at the time of a stimulus.
The \textit{prepotent response} system
prepares the motor action that the subject will perform
at the end of the delay period.
The \textit{proactive inhibition} system
suppresses that motor action until the delay period
is over.

\subsection{Neural integration as a core mechanism}

An important observation of Narayanan and Laubach
is that the second principal component
is the cumulative sum (integral)
of the first principal component.
This suggests that some neurons in these cortical areas
are representing the integral of another
signal represented in these areas.

Neural integration through recurrently connected
populations of neurons has been widely studied
in the context of working memory.
% ? cite a bunch -- Mawkinson, Roman, Tanji, X-J Wang, look at singh eliasmith biblio

The Singh and Eliasmith model, in particular,
facilitates a neural implementation of the conceptual model
proposed as a result of the PC analysis.
In addition to outlining a method of constructing
multidimensional neural integrators,
it also provides a method of connecting those integrators
to other populations of neuron representing other signals.

\subsection{Coupled integrators in a control system}

With non-integrative and integrative neuronal populations
as the building blocks, we propose a neural circuit
realization of the conceptual model
discussed in Narayanan and Laubach.
The model is summarized in figure 3.

Two populations are hypothesized to reside in motor cortex.
The \textit{Press/Release} population represents
whether the subject is currently executing
a press or release motor action.
The activity of this population directly affects the environment
in the model, which can be thought of
as functioning like Betz cells.
The \textit{Holding} population tracks
whether or not the subject is currently holding down the lever.
It is analogous to the \textit{prepotent response}
from the conceptual model, as in the absence of inhibition,
it will activate the \textit{Release} neurons
in the \textit{Press/Release} population.
The \textit{Holding} population is integrative,
and integrates the
\textit{Press} signal from the \textit{Press/Release}
population. The integration is controlled
by the \textit{Release} signal, meaning that
when the lever is released the \textit{Holding} signal
is reset to some baseline.

Three populations are hypothesized to reside
in medial prefrontal cortex.
The \textit{Waiting} population represents
that the subject is in a delay state,
waiting for the appropriate time to perform a motor action.
It is analogous to the \textit{proactive inhibition}
from the conceptual model, as it works to balance
the prepotent release response being signaled
from the motor cortex's \textit{Holding} population.
The \textit{Cue} population is sensitive
to the auditory stimulus delivered to the subject
when the delay period is over.
It inhibits the \textit{Waiting} population's
proactive inhibition, disinhibiting \textit{Release}
sensitive neurons.
Similarly, the \textit{Timer} population
tracks the relative amount of time that has passed
since the subject pressed down the lever,
starting the trial.
When the time of the stimulus can be predicted,
the \textit{Timer} population speeds
responding by increasingly inhibiting the
\textit{Waiting} population over time.
It can also effect a motor action on its own by
disinhibiting \textit{Release} neurons sufficiently.
The \textit{Timer} population is integrative,
gradually ramping up as it integrates
the \textit{Waiting} signal.

\section{Methods}

\subsection{Simple reaction time task}

In both the empirical and simulated data,
subjects and models performed the same simple reaction time task.
At the start of a trial,
subjects press down a lever.
A timer tracks the time since the lever was pressed down,
and after an experimenter-specified time period,
called the \textit{delay period} or \textit{foreperiod},
an auditory cue signals that the subject must respond
by releasing the lever.
If the lever is released within 600 milliseconds,
the trial is considered a success
and reward is delivered.
If the lever is released before the auditory cue
or after the 600 millisecond response window,
then the trials is considered a failure
and subjects are given a time out in which
lights are extinguished.
Lever presses during this timeout period
lengthen the timeout period.

Empirically, this study is done with rats
in a MedPC box containing a lever,
house lights, and a spout for delivering water reward.
Rats were trained by first
associating reward with lever presses,
and then introducing progressively longer
delay periods until they reached the goal delay
of 1 second.
They were considered to have learned the task
when they performed the task correctly $>60\%$
of the time.

Theoretically, this study was performed
in the Nengo simulation environment.
Subjects were networks of leaky integrate-and-fire
neurons, as described below.
The environment was modeled using a
\texttt{simplenode}, which allow arbitrary code
to interact with a neural model.
In this case, the \texttt{simplenode} implemented
a finite state machine to track task state,
and scalar values to indicate the state of
house lights, lever position, auditory cue, and reward.

\subsection{Empirical summary}

The empirical study (described in more depth in ?) %???cite???)
was done with twelve male Long-Evans rats (3-4 months old; Harlan).
Rats were water restricted, with food always available,
and consumed 10-15 ml of water during each
behavioural session.

Arrays of recording electrodes were implanted
in dmPFC ($n = 8$), motor cortex ($n = 3$), or both ($n = 1$).
Electrodes were 50 $\mu$m stainless steel wires
with 250 $\mu$m between wires.
Arrays were in configurations of
4 $\times$ 4 or 3 $\times$ 3 $\times$ 2.
Microwire arrays were targeted at dmPFC
based on the following coordinates from bregma:
anteroposterior [AP]: $+$3.2mm;
mediolateral [ML]: $\pm$1.4mm;
dorsoventral [DV]: $-$3.6mm at 10$^\circ$ in the frontal plane.
Motor cortex was targeted with
AP: $-$0.5mm; ML: $\pm$2.5-3.5mm;
DV: $-$1.5mm at $-25^\circ$ in the frontal plane.
Histology was performed to confirm electrode locations
(???see fig???).

Ensemble recordings were done using a
Many Neuron Acquisition Program (Plexon, Dallas, TX).
Spike sorting using principal component analysis (PCA)
and waveform shape identified 414 single units
whose spikes were included in this study.

\subsection{Theoretical}

Two models were created to explain neural processing
in the simple reaction time task.
The first is a double-integrator model,
as is described in ?, %???cite singh eliasmith???,
which allows for precise comparisons
between the spiking behaviour of
the model and the empirical subjects
given a controllable behavioural outcome.
The second is an integrated model that
expands the double-integrator model
to include populations that manipulate the environment,
allowing for the comparison between
the model and empirical behaviour.

Both models are made up of heterogeneous populations
of leaky integrate-and-fire neurons
created and connected using the methods
of Eliasmith and Anderson ?. % ???cite???.
Connections between populations of neurons
are determined based on the function computed
through the connection.
Connection weights are determined
by solving a least-squares minimization
of the function representation
given the tuning properties
of the neurons representing the function parameters,
as described in ? %???singheliasmith???
and in detail in ?. %???eliasmith anderson ???.

\subsubsection{Double-integrator model}

In  %???cite???
it is shown that
a population of heterogeneous adaptive leaky integrate-and-fire neurons
representing a two-dimensional signal
can capture the dynamics of
working memory cells in prefrontal cortex.
There are, however, several network configurations
that can result in a population of neurons
representing two integrated signals,
as shown in figure ?. %???copy fig 5 from paper???.

We adopt the coupled one-dimensional integrator model,
despite little evidence that reciprocal projections
are localized to smaller integrative populations.
Instead, it is the PCA of mPFC activity
that suggests many neurons are
only sensitive to one of the principal components.
Additionally, it is more intuitive
to connect each principal component with
a separate neural population.
The important task, then,
is to connect each population
with the appropriate value being encoded.

Figure ? %???put in???
summarizes our mapping.
The two coupled one-dimensional oscillators represent
the subject's existence in a particular behavioural state
and how long the subject has been in that state.
The behavioural state is entered
once the subject presses down the lever,
modeled by the activation
of a population of press-sensitive neurons.
The timing integrator is driven
by the state population, as follows from ?. %???citesingh???.
Both coupled integrators are controllable
through a second dimension, %(???cite eliasmith here? ???),
which allows the integrators to be
reset to an initial point (0.0)
when reward is delivered to the subject.
The two one-dimensional integrators
are then projected to an output population that
tracks both signals.

The press- and reward-sensitive populations
drive the rest of the activity of the network.
These two populations are driven by the experimenter,
based on the behaviour being simulated.
For example, for a sequence of correct trials,
the press-sensitive population
will be driven strongly at the start of a trial,
followed 1.6 seconds later
by strongly driving the reward-sensitive population.
For error trials, the reward population
is not driven between activations of the press population.

\subsubsection{Integrated motor-mPFC model}

In order to expand the double-integrator model
to directly manipulate elements of the environment,
three additional variables must be accounted for:
release, auditory cue, and control.

The PCA in ? %???kumar???
revealed
neurons that are sensitive to both press and release.
For this reason, the press-sensitive population
was increased in dimensionality
such that the population was sensitive
to both press and release,
in the same manner to the multidimensional
methods used in ?. %???singh eliasmith???.

The auditory cue ?
%??? todo : see if it's necessary to be an integrator ???

Control is the signal
that causes release-sensitive neurons to fire,
resulting in the release of the lever.
Determining how to implement this control
is one of the primary goals
of mPFC research in tasks with delays.
Its implementation is in itself
a hypothesis of how the mPFC and motor cortex interact
in order to perform these tasks.

\subsubsection{Implementing release control}

? %???kumar???
use their PCA to propose a conceptual model,
shown in Fig ?. %???include???.
We extend this model
by associating prepotent response and proactive inhibition
with neural populations and
showing how they can be connected
such that the model can respond
within the 600 ms response window.

In the double integrator model,
a population of neurons tracks the current behavioural state.
In the integrated model, two populations
track a similar but subtly different quantity.
One population is assumed to exist
in motor cortex, and tracks whether or not
the subject is currently holding down the lever.
This is an integrative population,
which is driven by press-sensitive neurons.
The other population is assumed to exist
in medial PFC, and tracks whether or not
the subject should be delaying the release response.
It is primarily driven by the motor cortex's
holding population.

The motor cortex's holding population
sends excitatory projections
to the release-sensitive neurons,
effecting the prepotent response from the conceptual model.
The medial PFC's waiting population
send inhibitory projects
to the release-sensitive neurons,
effecting the proactive inhibition from the conceptual model.
It should be noted that this organization
produces a slight delay between these two quantities,
with prepotent response occurring first,
quickly followed by proactive inhibition.
This is consistent with ?. %???figure from JNP2009???.

The medial PFC's waiting population's activity
is further modulated by two task-related populations.
One is a ramping population that tracks
the time since the press was initiated.
The other is a cue-sensitive population
that is driven by the onset of the auditory cue.
These two populations
release the proactive inhibition,
causing the release neurons to be disinhibited,
effecting the release of the lever.

As will be discussed further in ?, %???sec???,
the similarity between the two populations
may explain why the first two principal components
of motor cortex and mPFC are very similar.

\subsubsection{Model manipulations}

The model is manipulated in three ways
in this study.

The model emulates the effects of aging
by decreasing the number of neurons
in each neural population,
by uniformly decreasing the strength
of each synapse in the network,
and by randomly removing a fraction
of all synapses in the network.

Lesion studies are emulated
by targeted ablations of either
neurons or synapses in or between specific
neural populations.

Cooling is emulated by
lowering the distribution
of neural gains associated with the
leaky integrate-and-fire neurons
in the neural populations being cooled.
Time constants affecting transmission speed
between populations are raised
to emulate slowed transmission speed
as a result of the cooling of white matter.

\subsection{Analysis}

Both empirical and theoretical experiments
produce data in the form of spike trains
and event timestamps.
The files produced by NeuroExplorer and Nengo
were read by a customized version of Neo,
a Python package for reading neurophysiology data files.
The subsequent analysis was done
with custom written scripts in Python
using the NumPy and SciPy libraries for numerical analysis
and Matplotlib for plotting.
All analysis scripts are freely available
at ?. % github URL

\subsubsection{Behaviour}

Figure 4 summarizes a single trial
of the reaction time task.
In both the empirical and theoretical studies,
the following relevant event times are recorded:
lever press, lever release, reward delivery,
trigger stimulus onset, and house light extinguishing.
Sequences of event times can be used to
identify the result of each trial,
as shown in Figure ?.
Filtering the set of event timestamps
for these sequences identifies each trial outcome.

Reaction time distributions are also analyzed
for correct trials.
Reaction time is defined as the difference between
trigger stimulus onset and lever release.

\subsubsection{Neural}

Principal component analysis (PCA)
is used to identify the important features
in spike trains around certain events.
The previously mentioned techniques
were used to isolate an event of interest.
For each neuron, spike trains from 1 second prior to
2 seconds after each instance of an event are isolated.
Each perievent spike train is binned in 1 ms bins and is then
convolved with a Gaussian filter ($\sigma = $ 25 ms;
PCs are consistent with many other $\sigma$ values).
The spike trains were slightly extended
in order to eliminated edge artifacts.
These spike density functions
are normalized to z-scores,
and then averaged over all trials to
produce a matrix in which each row
the normalized average response of a neuron
over the perievent epoch.
PCA was performed on that matrix
using singular value decomposition.
The resulting principal components
were also normalized to z-scores.
The amount of variance accounted for by each principal component
is computed as the square of the component's eigenvalue
over the sum of all squared eigenvalues
(i.e., $s_i^2 / \sum s^2$).

In all analyses, neurons not significantly modulated
during the perievent epoch were omitted.
Specifically, we omitted neurons with average firing rates
under 1 Hz and whose firing rates were non-stationary
across trials
(determined by a Wald-Wolfowitz runs test with $|Z| > 4$).

\section{Results}

\subsection{Model behaviour is less variable}

A stereotypical trial for both the model
and experimental subjects is shown
in Figure 5.
The experimental session lasted ? seconds,
over which ? trials were performed.
? trials were identified as correct trials;
? were premature responses,
and ? were delayed responses.
The model session lasted 500 seconds,
over which ? trials were performed.
? trials were identified as correct trials,
? were premature responses
and ? were delayed responses.

Over all trials for all subjects,
experimental subjects performed ? trials
with a ? \% success rate.
The model performed ? trials with a
? \% success rate.

The distribution of reactions times, shown in figure 5,
is Gaussian-like with a long tail greater than the mean
for both model and data.
The mean reaction time for all experimental subjects
was ? $\pm$ ?.
The mean reaction time for all model subjects
was ? $\pm$ ?.

\subsection{Model principal components are similar}

Principal component analysis on the spike trains
produced by the model show the same number
of primary principal components, with similar shapes.

In the neural populations hypothesized
to reside in motor cortex, as shown in Figure 6,
three leading components are identified.
As in the experimental data,
PC1 is rises around the time of the press
and maintains a high value until the release,
at which point it drops.
PC2, which is very similar
to the cumulative sum of PC1,
is modulated at the time of lever press,
and slowly changes during the delay period.
PC3 is strongly modulated at both the
time of lever press and release.
Firing rates in the model match those in
the data, increasing before the press
and after release.

In the neural populations hypothesized
to reside in medial prefrontal cortex,
as shown in figure 7,
two leading components are identified.
The first component is strongly modulated
around the time of lever press,
and slowly decreases during the delay period.
The second component,
which is similar to the cumulative sum of
PC1, increases at the time of lever press,
and remains high during the delay period.

\subsection{Model manipulations}

Aging results, figure 8.

Lesion results, figure 9.

Cooling results, figure 10.

\section{Discussion}

\subsection{Model behavioural variability is low}

The primary difference is in the amount
of inter-trial and intra-trial variability.
Experimental subjects had several blocks of time
in which they performed as many trials as the model,
but those blocks of time were interrupted by
error trials and time in which the animal
was performing other actions such as grooming.
Within the trial, reaction time distribution
for the model and experimental subjects are
similar but with far less variability.
The model's reactions times
are within the experimental subjects'
mean and standard deviation, however,
suggesting that the model
could have the same variability
by simply injecting noise into the system.
This would also result in
the error trials that are missing in the model.

A better basis for comparison for the model
would be human performance in a similar
reaction time task. ?find some % Look up some behavioural data

\subsection{Model principal components match the data}

While it may be unsurprising
due to the model being inspired by the original
principal component analysis,
it is strong validation for the original conceptual model
that the implemented model maintain
the same principal components
while still being able to simulate behaviours.

? List differences and investigate (with final version of model)

\subsection{Model predictions}

The model manipulations assert that blah has the effect of blah.

Prior studies (some unpublished) have suggested that these are indeed the effects of those things.

Cooling has not been studied in an experimental setting,
and so we take the results as a prediction of the model.

\clearpage

\appendix

\section{Figures}

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{f1-pcasummary}
    \caption{Summary of principal components identified
      by Narayanan and Laubach, 2009. Three components
      were identified in motor cortex (top row), accounting
      for over 60\% of variation.
      Two components were identified in prefrontal cortex
      (bottom row), accounting for over 25\% of variation.}
\end{figure}

\clearpage

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{f2-conceptual}
    \caption{The conceptual model proposed by Narayana and
      Laubach, 2009. In the conceptual model, the start of the
      trial activates two systems; one prepares the response
      action (prepotent response) and the other inhibits
      that action until the appropriate time (proactive inhibition).}
\end{figure}

\clearpage

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{f3-fullmodel}
    \caption{The proposed implementation of the conceptual model.
      The motor cortex is made up of two populations of
      leaky integrate-and-fire neurons, one of which interacts
      with the environment by pressing or releasing a lever.
      The medial prefrontal cortex is made up of three
      populations of LIF neurons. See text (section 2) for details.}
\end{figure}

\clearpage

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{f4-behaviour}
    \caption{Possible outcomes for each trial. If the subject
      responds within the 0.6 second response window, the trial
      is counted as correct and reward is delivered. If the
      subject responds prematurely, it is counted as an error
      and house lights are extinguished for a timeout period.
      If the subject responds after the response window,
      an error is also recorded and the lights are extinguished.}
\end{figure}

\clearpage

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{f5-behaviour}
    \caption{Summary of model and experimental behaviour.
      (Left) Seconds 200 - 500 of an experimental (top)
      and model (bottom) session.
      (Right) Histograms of reactions times for
      experimental and model subjects.}
\end{figure}

\clearpage

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{f6-mcpca}
    \caption{Principal components identified in motor cortex
      spike data. (Top) Experiment. (Bottom) Model.}
\end{figure}

\clearpage

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{f7-pfcpca}
    \caption{Principal components identified in medial prefrontal
      cortex spike data. (Top) Experiment. (Bottom) Model.}
\end{figure}

\clearpage

\begin{figure}
  \centering
    % \includegraphics[width=\columnwidth]{f7-pfcpca}
    Placeholder
    \caption{Aging results}
\end{figure}

\clearpage

\begin{figure}
  \centering
    % \includegraphics[width=\columnwidth]{f7-pfcpca}
    Placeholder
    \caption{Lesion results}
\end{figure}

\clearpage

\begin{figure}
  \centering
    % \includegraphics[width=\columnwidth]{f7-pfcpca}
    Placeholder
    \caption{Cooling results}
\end{figure}

\end{document}

### 2013-06-10 paper/jneurosci2013.tex
Second version of the paper, with figures.

@@
-%\documentclass[twocolumn]{article}
 \documentclass[12pt]{article}
@@
 
-\newcommand{\framedscalefigone}[3]{
-  \begin{figure}[ht!]
-    % Requires \usepackage{graphicx}
-    \centering
-    \begin{framed}
-      \begin{minipage}{\textwidth}
-        \centering
-        \includegraphics[width=#2\columnwidth]{#1}
-        \caption{#3}
-        \label{#1}
-      \end{minipage}
-    \end{framed}
-  \end{figure}}
-
-\newcommand{\framedscalefigtwo}[4]{
-  \begin{figure}[ht!]
-    % Requires \usepackage{graphicx}
-    \centering
-    \begin{framed}
-      \begin{minipage}{0.49\textwidth}
-        \includegraphics[width=#3\columnwidth]{#1}
-      \end{minipage}
-      \begin{minipage}{0.49\textwidth}
-        \includegraphics[width=#3\columnwidth]{#2}
-      \end{minipage}
-      \caption{#4}
-      \label{#1}
-    \end{framed}
-  \end{figure}}
-
 \newcommand{\scalefigone}[3]{
@@
 
-\newcommand{\scalefigtwo}[4]{
-  \begin{figure}[ht!]
-    % Requires \usepackage{graphicx}
-    \centering
-    \includegraphics[width=#3\columnwidth]{#1}
-    \includegraphics[width=#3\columnwidth]{#2}
-    \caption{#4}
-    \label{#1}
-  \end{figure}}
-
 \setlength\fboxrule{1pt}
@@
 \begin{center} \begin{Large}
-  \textbf{A spiking neural circuit model of cortical function
-          in simple reaction time tasks}
+  \textbf{A task-state model of medial PFC \\ can perform top-down
+    control}
 \end{Large} \end{center}
@@
 \begin{abstract}
-  todo
+  Many hypotheses on the function of medial prefrontal cortex (mPFC)
+  have been put forth,
+  including action-outcome learning, ???More.
+  We adopt a more general approach and posit that
+  the mPFC tracks task state,
+  which can be usd to accomplish previously hypothesized functions.
+  We use as a model rodents performing a simple reaction time (RT) task.
+  First, we show that mPFC activity
+  tracks task state in the simple RT task,
+  and that neural integration is a core mechanism
+  for accomplishing task state tracking.
+  Second, we show that a working memory model
+  using neural integrators closely matches
+  rodent neural activity.
+  Finally, we propose a neural circuit model
+  that imposes top-down control on a control system
+  as a proof-of-concept that the task-state model
+  can explain the myriad of functions ascribed
+  to the medial PFC.
 \end{abstract}
@@
 
-Animals make complicated motor actions
-with high spatial and temporal resolutions.
-In temporally precise tasks,
-the motor cortex and several afferents coordinate
-to bring about motor actions at appropriate times.
-For tasks with delay periods,
-the medial prefrontal cortex is critically important
-for suppressing motor action during the delay period.
-This effect has been seen in ?.  %???primates humans rodents etc???.
-
-In rodents, the role of the medial prefrontal cortex
-has been investigated through ?.
-%???previous work???.
-Previously, we have used a simple reaction task
-and principal component analysis
-to investigate ?. %???summarize Narayana \& laubach jnp 2009???
-These results point to neural integration
-as playing a key role
-in the mPFC's ability
-to control the timing of motor actions.
-
-In order to investigate the neural circuit
-underlying the production and suppression of motor output,
-we extend Singh \& Eliasmith's model to include
-neural populations that directly affect the environment,
-analogous to the function of Betz cells.
-In doing so, we can relate the model to the experimental data
-in both neural and behavioural contexts.
-
-\section{Model}
-
-The model presented in this study
-is based on a conceptual model of PFC and MC
-function in delay tasks proposed by Narayanan and Laubach.
-Their conceptual model is based on
-principal component analysis of neural data
-in rats performing a simple reaction time task.
-
-Figure 1 summarizes their PC analysis.
-In PFC, two components account for over 35\%
-of variance in population firing.
-The first principal component increases slightly
-at the time of the lever press,
-and then decreases over the delay period.
-The second principal component
-also increases at the time of lever press,
-but remains high until the time of release.
-
-In motor cortex, three components account
-for over 60\% of variance in population firing.
-The first two principal components are
-similar to those found in PFC;
-the first steadily decreases over the delay period,
-and the second remains high during the delay period.
-The third principal component increases
-sharply slightly before lever press
-and lever release.
-
-The identification of these principal components
-leads to a conceptual model of cortical function
-during simple reaction time tasks.
-This model is summarized in figure 2. % Fig
+mPFC function background
+
+mPFC models background
+
+We propose that the function of the mPFC is to
+track task states.
+These task states can be very high-dimensional,
+if, for example, visual features are necessary
+for performing a particular task.
+The other functions that have been ascribed
+to the mPFC are accomplished
+by other areas, but use the task state
+information in the mPFC to perform their function.
+
+As an application of this theory,
+we investigate a simple reaction-time task.
+The remainder of the paper is organized as follows.
+First, we review experimental data from rodents
+doing the reaction-time task.
+The task state space being tracked is determined
+by a dimensionality reduction technique
+(principal component analysis).
+Second, we review a spiking neural model
+of working memory and
+analyze it during the simple reaction-time task.
+Third, we analyze the two-dimensional dynamical system
+that makes up the task state space and
+show how it changes as a result of performing the task.
+Fourth, we create a neural control system
+that can perform the simple reaction-time task
+and show how the mPFC model can exert top-down control
+on that control system.
+Finally, we give some examples of how the model
+can be manipulated in order to predict
+neural and behavioural effects in certain pathologies.
+
+\section{Experimental data}
+
+Narayanan \& Laubach (2009) recorded the medial PFC
+of rodents performing a simple reaction-time task
+(figure ???).
+In this task, the subject presses and holds down the lever
+for a delay (the ``foreperiod'';
+1 second for the data in this study).
+After the foreperiod, an auditory tone (stimulus) is played.
+The subject then has 0.6 seconds (the response window)
+in which to release the lever.
+If the lever is released in that time,
+the subject is rewarded.
+If the lever is released during the foreperiod,
+it is called a premature error,
+and is penalized by a time-out.
+If the lever is not released within the response window,
+it is called a late error,
+and is also penalized by a time-out.
+
+Narayanan \& Laubach used principal component analysis
+in order to determine the important signals
+being represented in the mPFC.
+As shown in figure ???,
+the first two principal components were significant,
+and appear to track
+1) if the subject was performing the task, and
+2) the relative amount of time the subject has been
+performing the task.
+Narayanan \& Laubach also found that the neural activity
+in the mPFC changed significantly depending on
+the outcome of the previous trial;
+the first two principal components
+change significantly after errors,
+and appear to contain a trace that an error occurred.
+
+Interestingly, in the post-correct case,
+the first principal components
+is very high correlated with the integral
+of the second principal components,
+and similarly the second principal component
+with the integral of the first.
+In the post-error case,
+these correlations are significantly weaker.
+This result suggests that
+the mPFC is performing some kind of integration.
+
+\section{Coupled neural integrators}
+
+One model of neural integration is
+Singh \& Eliasmith (2006).
+Their model was created to explain working memory activity
+in PFC during a vibrotactile discrimination task.
+It employs two-dimensional neurons in integrative populations
+in order to show time-varying activity during the task.
+Due to the identification of two strong components
+in the rodent data,
+we hypothesized that the Singh \& Eliasmith model
+could be used to model the simple RT task as well.
+
+There are two different realizations
+of this model, as shown in figure ???.
+In one, a single population of adaptive
+leaky integrate-and-fire neurons tracks both
+integrative dimensions, as well as two control dimensions.
+In the other, two coupled populations
+of adaptive leaky integrate-and-fire neurons
+each track one integrative dimension, as well
+as one control dimension.
+The integrative dimensions are projected
+to another population that does not explicitly
+represent the control dimensions.
+In both realizations of the model,
+input to the integrators is provided
+by motor actions, and are controlled
+by the task outcome.
+The idea here is that a motor action
+(in this case a lever press)
+starts the presistent activations
+of one of the integrative dimensions.
+That dimension provides input to the other integrative dimension.
+Task outcomes (in this case either reward
+or time-out penalties) affect the control dimensions,
+either stopping all integration
+in the reward case or
+reversing the integration in the penalty case.
+
+These two different model configurations
+make an important distinction between
+what is being represented in the brain
+area being analyzed.
+In our case, should the single population model
+match the neural data,
+it predicts that the task information
+as well as the action and outcome information
+is being represented in the mPFC.
+If the coupled population model matches
+the neural data,
+it predicts that only the task information
+is being represented, and action and outcome
+information only affects the mPFC's
+task state.
+
+As can be seen in figure ???,
+the coupled integrator model matches
+the neural data closely, while the single population model
+does not.
+In the single population model,
+there is highly modulated activity
+around the time of reward delivery;
+this modulation is not seen in the data.
+Therefore, it can be concluded that,
+at least for the rodent in a simple RT task,
+outcome information is not explicitly
+represented in the mPFC,
+though the outcome does affect the task
+information reflected in mPFC activity.
+We will analyze the coupled integrator
+realization of the Singh \& Eliasmith
+model in the subsequent sections.
+
+\section{Task state space analysis}
+
+If the role of the mPFC is to track the task-state,
+then we can view mPFC activity as a trajectory
+in the task state space.
+This is done in figure ??? for the experimental data,
+and the simulated data.
+
+In the post-correct case,
+the trajectory through task space is cyclical.
+In other words,
+when the subject is performing the task properly,
+the mPFC travels in a predictable loop.
+In the post-error case,
+the trajectory through task space is less predictable.
+
+The dynamics of this task state space
+can be approximated based on these trajectories,
+and on how the Singh \& Eliasmith model was created.
+These dynamics are captured in the following equation.
+
+\begin{equation*}
+  \begin{bmatrix}
+    \dot{x_1} \\ \dot{x_2}
+  \end{bmatrix} = \begin{bmatrix}
+    -\beta R & 0 \\ \alpha & -\beta R
+  \end{bmatrix} \begin{bmatrix}
+    x_1 \\ x_2
+  \end{bmatrix} + \begin{bmatrix}
+    1 & -1 \\ 0 & 0
+  \end{bmatrix} \begin{bmatrix}
+    A \\ P
+  \end{bmatrix} 
+\end{equation*}
+where $A$ is the action (lever press),
+$R$ is the reward, and $P$ is the penalty.
+Figure ??? shows the result of directly simulating
+the trajectory of this dynamical system.
+
+The Singh \& Eliasmith model uses the methods
+of the Neural Engineering Framework (NEF),
+which can simulate arbitrary dynamical systems.
+The dynamical system defines how the model
+encodes information in the populations of neurons,
+and how those encoded values are transformed
+through connections between populations.
+Because we know exactly how the information
+is encoded and transformed, we can also
+decode that information directly;
+this results in figure ???,
+which looks like a noisy version
+of the mathematical model.
+It is important to note that this decoding process
+was done on the same neural data that
+results in the principal components in figure ???;
+therefore, we can say with some certainty
+that the mathematical model is closely emulating
+the experimental mPFC.
+
+\section{Top-down control of motor actions}
+
+Previous work by Laubach and colleagues (and other groups)
+has implicated mPFC as a top-down controller
+on the motor cortex in delay tasks like simple RT.
+Their work suggests that the mPFC directly inhibits
+the motor cortex.
+In this paper, we propose instead that the mPFC
+can either inhibit, directly excite,
+or modulate other activity being projected
+to the motor cortex depending on
+the part of task-state space the mPFC is in.
+
+In order to demonstrate that this is possible,
+we created a simple stimulus-response control system
+using the methods used in Singh \& Eliasmith, 2006
+(i.e., the NEF; see figure ???).
+Briefly, the system is composed of three populations
+of adaptive leaky integrate-and-fire neurons.
+One population presses and releases the lever,
+one population tracks if the lever is currently being held down,
+and one population is sensitive to the auditory stimulus,
+and stimulates release neurons if the lever
+is currently being held down.
+This system is trivial to build,
+and can perform the simple RT task
+with little difficulty (see figure ???).
+
+Top-down control can be accomplished
+because of the predictability with which
+the mPFC traverses the task state space.
+If the subject continues to perform the task correctly,
+the release should happen at the same part
+of the state space (see figure ???);
+if the subject forces a release at the appropriate
+part of the state space, it can, in fact,
+release the lever faster than it would be able to
+if it waits for the stimulus.
+In fact, if the foreperiod is completely
+predictable, then even devoid of a stimulus,
+the subject could perform this task
+solely through a forced release mediated
+by the mPFC.
+
+Through a similar argument, the mPFC
+can also cause the subject to be more
+cue-sensitive after errors occur.
+When an error occurs, the mPFC is driven to a
+low part of the state space.
+When the initial press occurs, it may not
+be driven to the top part of the state space
+as it should; if that occurs (i.e.,
+if the subject is performing this task
+in the wrong part of the state space),
+the subject can attempt to become more cue-sensitive
+(see figure ???).
+
+This top-down control can be exerted
+on the stimulus-response motor control system
+with the network configuration depicted in
+figure ???. Briefly, top-down control
+is mediated by directly stimulating
+the population that is sensitive to the auditory stimulus.
+When the mPFC is in the force-release zone,
+then these neurons will be excited;
+this can either make the population react quicker to
+the stimulus, or can directly cause a release
+even in the absence of a stimulus.
+When the mPFC is in the cue-sensitive zone,
+these neurons are stimulated in such a way that
+they react quicker and stronger to the stimulus,
+but cannot directly cause a release
+in the absence of a stimulus.
+
+\section{Task-state space dynamics and errors}
+
+Looking at the dynamics of the task state space
+at each part of a simple RT task trial
+(figure ???)
+gives a full picture of what the task-state
+theory of mPFC function predicts is occurring
+during the simple RT task.
+
 At the start of a trial,
-two complementary systems are activated
-to produce quick reactions at the time of a stimulus.
-The \textit{prepotent response} system
-prepares the motor action that the subject will perform
-at the end of the delay period.
-The \textit{proactive inhibition} system
-suppresses that motor action until the delay period
-is over.
-
-\subsection{Neural integration as a core mechanism}
-
-An important observation of Narayanan and Laubach
-is that the second principal component
-is the cumulative sum (integral)
-of the first principal component.
-This suggests that some neurons in these cortical areas
-are representing the integral of another
-signal represented in these areas.
-
-Neural integration through recurrently connected
-populations of neurons has been widely studied
-in the context of working memory.
-% ? cite a bunch -- Mawkinson, Roman, Tanji, X-J Wang, look at singh eliasmith biblio
-
-The Singh and Eliasmith model, in particular,
-facilitates a neural implementation of the conceptual model
-proposed as a result of the PC analysis.
-In addition to outlining a method of constructing
-multidimensional neural integrators,
-it also provides a method of connecting those integrators
-to other populations of neuron representing other signals.
-
-\subsection{Coupled integrators in a control system}
-
-With non-integrative and integrative neuronal populations
-as the building blocks, we propose a neural circuit
-realization of the conceptual model
-discussed in Narayanan and Laubach.
-The model is summarized in figure 3.
-
-Two populations are hypothesized to reside in motor cortex.
-The \textit{Press/Release} population represents
-whether the subject is currently executing
-a press or release motor action.
-The activity of this population directly affects the environment
-in the model, which can be thought of
-as functioning like Betz cells.
-The \textit{Holding} population tracks
-whether or not the subject is currently holding down the lever.
-It is analogous to the \textit{prepotent response}
-from the conceptual model, as in the absence of inhibition,
-it will activate the \textit{Release} neurons
-in the \textit{Press/Release} population.
-The \textit{Holding} population is integrative,
-and integrates the
-\textit{Press} signal from the \textit{Press/Release}
-population. The integration is controlled
-by the \textit{Release} signal, meaning that
-when the lever is released the \textit{Holding} signal
-is reset to some baseline.
-
-Three populations are hypothesized to reside
-in medial prefrontal cortex.
-The \textit{Waiting} population represents
-that the subject is in a delay state,
-waiting for the appropriate time to perform a motor action.
-It is analogous to the \textit{proactive inhibition}
-from the conceptual model, as it works to balance
-the prepotent release response being signaled
-from the motor cortex's \textit{Holding} population.
-The \textit{Cue} population is sensitive
-to the auditory stimulus delivered to the subject
-when the delay period is over.
-It inhibits the \textit{Waiting} population's
-proactive inhibition, disinhibiting \textit{Release}
-sensitive neurons.
-Similarly, the \textit{Timer} population
-tracks the relative amount of time that has passed
-since the subject pressed down the lever,
-starting the trial.
-When the time of the stimulus can be predicted,
-the \textit{Timer} population speeds
-responding by increasingly inhibiting the
-\textit{Waiting} population over time.
-It can also effect a motor action on its own by
-disinhibiting \textit{Release} neurons sufficiently.
-The \textit{Timer} population is integrative,
-gradually ramping up as it integrates
-the \textit{Waiting} signal.
-
-\section{Methods}
+the system is driven upwards towards [1.0, 0.0].
+This sets the initial state for the task.
+The subject may have been doing another task
+before initiating this RT task;
+by resetting to the same initial state,
+the subject's behaviour can proceed predictably.
+As will be seen, however, after errors,
+the system may be sufficiently far from this
+initial state that it cannot quite get to the initial state,
+and instead operates in an unpredictable part of the task space.
+
+In a correct trial (b), the passive dynamics of the system
+push the state to the right
+at a predictable (i.e., learned) rate.
+The release happens when the state of the system
+is sufficiently far to the right (i.e., $x_1$ is high).
+If this release happens properly, reward is delivered,
+which pushes the system back to the origin.
+Then, in the absence of any task-related dynamics,
+the system will stay approximately at rest
+because the passive dynamics of the system are very weak
+at the origin point.
+
+A premature error can be caused by some other factor
+(i.e., a motor system mistake) or by the passive dynamics
+of the system being too strong in the part of the state space
+that the system is driven to once the subject presses the lever.
+In the case of some other factor,
+the system will receive the negative outcome feedback
+before reaching the right side of the state-space,
+and be driven to the bottom.
+In the case of fast integration,
+the system will reach the right side of the state-space,
+force a release, and then receive the negative outcome feedback
+and be driven to the bottom.
+The exact state after receiving the negative feedback
+can be used to figure out the cause of the error,
+though after some time in the lower part
+of the state space, the system will be
+driven to the left,
+and all that can be decoded
+is the notion that an error occurred.
+
+A late error can be caused by either another factor
+not responding to the top-down control properly or in time,
+by right-ward drift happening too slowly,
+or by the system not being driven
+to the appropriate place upon pressing down the lever
+and drifting slowly as a result.
+
+\section{Behavioural results}
+
+While there are a number of tunable parameters,
+(e.g., $\alpha$ and $\beta$ from equation ???),
+there are many more parameters that are randomly selected
+from probability distributions.
+Neuron parameters (neuronal gain, encoding vectors,
+adaptation constants, etc.) are all randomized,
+and the results of those randomizations are used
+to solve for appropriate connection weights.
+As such, there is a large amount of variation
+between instances of this model.
+For that reason,
+we generate several instances of the model
+and run the model for over forty trials
+to get an idea of how
+that particular instance of the model behaves.
+
+Figure ??? summarizes the behavioural results
+from the experimental rodents, the stimulus-response model
+alone, and the stimulus-response model with top-down
+control by a coupled integrator model.
+The experimental rodents ???.
+The stimulus-response model by itself has a relatively slow
+reaction time (median ???, $\pm$ ???),
+but very high accuracy (??? correct, ??? late, 0\% early).
+With top-down control, some models were able to achieve
+extremely fast reaction times (e.g., ???).
+On average, however, the median reaction time was
+??? $\pm$ ???. Note that the variance in reactions times
+was much higher with top-down control,
+and the accuracy was in general lower (??? correct, ??? late,
+??? early).
+
+\section{Model manipulations}
+
+As a spiking neuron model,
+this theory of mPFC function is uniquely able to be
+manipulated; manipulations at the neural level
+can be examined at any point, from the single-cell
+to the behavioural level.
+
+One possible manipulation is to
+ablate a portion of recurrent synapses,
+which could model the effect of brain trauma or aging.
+The effects on the mPFC's ability to track the
+task state space can be seen in figure ???.
+%%% MORE
+
+Since we also know the mathematics behind the mPFC activity,
+we can instead manipulate the mathematical model
+and make detailed predictions about neural activity.
+For example, it is hypothesized that in aging,
+recurrent connections become weaker, and therefore
+the persistent activity that is seen in working memory models
+cannot be maintained for long periods of time.
+This can be modeled by including a small damping term, $d < 0$,
+in equation ???.
+
+\begin{equation*}
+  \begin{bmatrix}
+    \dot{x_1} \\ \dot{x_2}
+  \end{bmatrix} = \begin{bmatrix}
+    -\beta R + d & 0 \\ \alpha & -\beta R + d
+  \end{bmatrix} \begin{bmatrix}
+    x_1 \\ x_2
+  \end{bmatrix} + \begin{bmatrix}
+    1 & -1 \\ 0 & 0
+  \end{bmatrix} \begin{bmatrix}
+    A \\ P
+  \end{bmatrix} 
+\end{equation*}
+
+In addition to seeing the lower-level effects
+of these two manipulations,
+we can also make the same manipulations in the full
+top-down model. As can be seen in figure ???, % STUFF
+
+\subsection{Predictions}
+
+??? collect together predictions
+
+\section{Discussion}
+
+??? Talk about the results
+
+On interesting feature of the model is its stochastic nature.
+While it can be though to represent individual differences,
+it is also a useful notion to consider that the model
+represents one snapshot during a subject's
+learning of the task.
+There is no learning in this model;
+while the connection weights are solved for ``optimally,''
+due to the large amount of noise in the system,
+learning would be very likely to fine-tune
+the mPFC's ability to traverse and
+be sensitive to parts of the task-state space.
+Therefore, the poor performance of, for example,
+top-down model instance ??? can be thought of as
+a subject whose recurrent connection is too strong
+and therefore is a poor performer in the task,
+or it could be thought of as a subject
+who has not yet learned how to properly time
+the interval, and after more trials would,
+with learning, be able to make less premature errors.
+
+Stepping back, the overarching theory of the paper,
+that the mPFC tracks the task state space, ??? more
+
+The theory sets up the possibility for many arbitrary tasks.
+Basically, each task has $N$ important dimensions.
+The passive dynamics are likely a combination
+of integrative and non-integrative dimensions,
+enabling for flows in many directions of the space.
+The mPFC is drive to specific parts of the state space
+based on the sensory environment (e.g.,
+being pushed to the top of the state space when lever pressing).
+The mPFC then learns to be sensitive to salient
+parts of the state space (e.g.,
+the force release zone).
+
+\clearpage
+
+\appendix
+
+\section{Materials and methods}
 
@@
 
-Two models were created to explain neural processing
-in the simple reaction time task.
-The first is a double-integrator model,
-as is described in ?, %???cite singh eliasmith???,
-which allows for precise comparisons
-between the spiking behaviour of
-the model and the empirical subjects
-given a controllable behavioural outcome.
-The second is an integrated model that
-expands the double-integrator model
-to include populations that manipulate the environment,
-allowing for the comparison between
-the model and empirical behaviour.
-
-Both models are made up of heterogeneous populations
-of leaky integrate-and-fire neurons
-created and connected using the methods
-of Eliasmith and Anderson ?. % ???cite???.
-Connections between populations of neurons
-are determined based on the function computed
-through the connection.
-Connection weights are determined
-by solving a least-squares minimization
-of the function representation
-given the tuning properties
-of the neurons representing the function parameters,
-as described in ? %???singheliasmith???
-and in detail in ?. %???eliasmith anderson ???.
-
 \subsubsection{Double-integrator model}
 
-In  %???cite???
-it is shown that
-a population of heterogeneous adaptive leaky integrate-and-fire neurons
-representing a two-dimensional signal
-can capture the dynamics of
-working memory cells in prefrontal cortex.
-There are, however, several network configurations
-that can result in a population of neurons
-representing two integrated signals,
-as shown in figure ?. %???copy fig 5 from paper???.
-
-We adopt the coupled one-dimensional integrator model,
-despite little evidence that reciprocal projections
-are localized to smaller integrative populations.
-Instead, it is the PCA of mPFC activity
-that suggests many neurons are
-only sensitive to one of the principal components.
-Additionally, it is more intuitive
-to connect each principal component with
-a separate neural population.
-The important task, then,
-is to connect each population
-with the appropriate value being encoded.
-
-Figure ? %???put in???
-summarizes our mapping.
-The two coupled one-dimensional oscillators represent
-the subject's existence in a particular behavioural state
-and how long the subject has been in that state.
-The behavioural state is entered
-once the subject presses down the lever,
-modeled by the activation
-of a population of press-sensitive neurons.
-The timing integrator is driven
-by the state population, as follows from ?. %???citesingh???.
-Both coupled integrators are controllable
-through a second dimension, %(???cite eliasmith here? ???),
-which allows the integrators to be
-reset to an initial point (0.0)
-when reward is delivered to the subject.
-The two one-dimensional integrators
-are then projected to an output population that
-tracks both signals.
-
-The press- and reward-sensitive populations
-drive the rest of the activity of the network.
-These two populations are driven by the experimenter,
-based on the behaviour being simulated.
-For example, for a sequence of correct trials,
-the press-sensitive population
-will be driven strongly at the start of a trial,
-followed 1.6 seconds later
-by strongly driving the reward-sensitive population.
-For error trials, the reward population
-is not driven between activations of the press population.
-
-\subsubsection{Integrated motor-mPFC model}
-
-In order to expand the double-integrator model
-to directly manipulate elements of the environment,
-three additional variables must be accounted for:
-release, auditory cue, and control.
-
-The PCA in ? %???kumar???
-revealed
-neurons that are sensitive to both press and release.
-For this reason, the press-sensitive population
-was increased in dimensionality
-such that the population was sensitive
-to both press and release,
-in the same manner to the multidimensional
-methods used in ?. %???singh eliasmith???.
-
-The auditory cue ?
-%??? todo : see if it's necessary to be an integrator ???
-
-Control is the signal
-that causes release-sensitive neurons to fire,
-resulting in the release of the lever.
-Determining how to implement this control
-is one of the primary goals
-of mPFC research in tasks with delays.
-Its implementation is in itself
-a hypothesis of how the mPFC and motor cortex interact
-in order to perform these tasks.
-
-\subsubsection{Implementing release control}
-
-? %???kumar???
-use their PCA to propose a conceptual model,
-shown in Fig ?. %???include???.
-We extend this model
-by associating prepotent response and proactive inhibition
-with neural populations and
-showing how they can be connected
-such that the model can respond
-within the 600 ms response window.
-
-In the double integrator model,
-a population of neurons tracks the current behavioural state.
-In the integrated model, two populations
-track a similar but subtly different quantity.
-One population is assumed to exist
-in motor cortex, and tracks whether or not
-the subject is currently holding down the lever.
-This is an integrative population,
-which is driven by press-sensitive neurons.
-The other population is assumed to exist
-in medial PFC, and tracks whether or not
-the subject should be delaying the release response.
-It is primarily driven by the motor cortex's
-holding population.
-
-The motor cortex's holding population
-sends excitatory projections
-to the release-sensitive neurons,
-effecting the prepotent response from the conceptual model.
-The medial PFC's waiting population
-send inhibitory projects
-to the release-sensitive neurons,
-effecting the proactive inhibition from the conceptual model.
-It should be noted that this organization
-produces a slight delay between these two quantities,
-with prepotent response occurring first,
-quickly followed by proactive inhibition.
-This is consistent with ?. %???figure from JNP2009???.
-
-The medial PFC's waiting population's activity
-is further modulated by two task-related populations.
-One is a ramping population that tracks
-the time since the press was initiated.
-The other is a cue-sensitive population
-that is driven by the onset of the auditory cue.
-These two populations
-release the proactive inhibition,
-causing the release neurons to be disinhibited,
-effecting the release of the lever.
-
-As will be discussed further in ?, %???sec???,
-the similarity between the two populations
-may explain why the first two principal components
-of motor cortex and mPFC are very similar.
-
-\subsubsection{Model manipulations}
-
-The model is manipulated in three ways
-in this study.
-
-The model emulates the effects of aging
-by decreasing the number of neurons
-in each neural population,
-by uniformly decreasing the strength
-of each synapse in the network,
-and by randomly removing a fraction
-of all synapses in the network.
-
-Lesion studies are emulated
-by targeted ablations of either
-neurons or synapses in or between specific
-neural populations.
-
-Cooling is emulated by
-lowering the distribution
-of neural gains associated with the
-leaky integrate-and-fire neurons
-in the neural populations being cooled.
-Time constants affecting transmission speed
-between populations are raised
-to emulate slowed transmission speed
-as a result of the cooling of white matter.
+The Singh \& Eliasmith model (2006).
+This model is created using the methods of the
+Neural Engineering Framework.
+The model uses neural integration in order to explain
+neural data in a working memory task.
+The paper proposes two candidate models
+that can create the desired functional effects
+with different neural dynamics.
+
+In the single population model, a single population of
+adaptive leaky integrate-and-fire neurons represents
+a four-dimensional signal.
+A recurrent connection causes the signals
+being represented in two of the dimensions
+to integrate; the connection
+also causes one of the integrative dimensions
+to drive the second integrative dimension.
+The remaining two dimensions are used
+to control the integrative dimensions,
+returning them to some baseline activity if necessary.
+
+In the coupled population model, two populations of
+adaptive leaky integrate-and-fire neurons represent
+two two-dimensional signals.
+Recurrent connections on each population cause
+one of the represented dimensions to be integrated,
+while the other dimension controls the integration.
+An explicit connection between one integrative population
+drives the integrator in the other population.
+Both of the integrative dimensions are then
+projected to a final two-dimensional population,
+which only represents the integrative signals.
+
+In this paper, we hypothesize that the mPFC
+is a realization of one of these two models.
+The integrators are driven by the actions taken by the subjects
+(e.g., lever pressing),
+and are controlled by the outcomes
+(e.g., reward delivery).
 
@@
 
-Figure 4 summarizes a single trial
-of the reaction time task.
 In both the empirical and theoretical studies,
@@
 
-\section{Results}
-
-\subsection{Model behaviour is less variable}
-
-A stereotypical trial for both the model
-and experimental subjects is shown
-in Figure 5.
-The experimental session lasted ? seconds,
-over which ? trials were performed.
-? trials were identified as correct trials;
-? were premature responses,
-and ? were delayed responses.
-The model session lasted 500 seconds,
-over which ? trials were performed.
-? trials were identified as correct trials,
-? were premature responses
-and ? were delayed responses.
-
-Over all trials for all subjects,
-experimental subjects performed ? trials
-with a ? \% success rate.
-The model performed ? trials with a
-? \% success rate.
-
-The distribution of reactions times, shown in figure 5,
-is Gaussian-like with a long tail greater than the mean
-for both model and data.
-The mean reaction time for all experimental subjects
-was ? $\pm$ ?.
-The mean reaction time for all model subjects
-was ? $\pm$ ?.
-
-\subsection{Model principal components are similar}
-
-Principal component analysis on the spike trains
-produced by the model show the same number
-of primary principal components, with similar shapes.
-
-In the neural populations hypothesized
-to reside in motor cortex, as shown in Figure 6,
-three leading components are identified.
-As in the experimental data,
-PC1 is rises around the time of the press
-and maintains a high value until the release,
-at which point it drops.
-PC2, which is very similar
-to the cumulative sum of PC1,
-is modulated at the time of lever press,
-and slowly changes during the delay period.
-PC3 is strongly modulated at both the
-time of lever press and release.
-Firing rates in the model match those in
-the data, increasing before the press
-and after release.
-
-In the neural populations hypothesized
-to reside in medial prefrontal cortex,
-as shown in figure 7,
-two leading components are identified.
-The first component is strongly modulated
-around the time of lever press,
-and slowly decreases during the delay period.
-The second component,
-which is similar to the cumulative sum of
-PC1, increases at the time of lever press,
-and remains high during the delay period.
-
-\subsection{Model manipulations}
-
-Aging results, figure 8.
-
-Lesion results, figure 9.
-
-Cooling results, figure 10.
-
-\section{Discussion}
-
-\subsection{Model behavioural variability is low}
-
-The primary difference is in the amount
-of inter-trial and intra-trial variability.
-Experimental subjects had several blocks of time
-in which they performed as many trials as the model,
-but those blocks of time were interrupted by
-error trials and time in which the animal
-was performing other actions such as grooming.
-Within the trial, reaction time distribution
-for the model and experimental subjects are
-similar but with far less variability.
-The model's reactions times
-are within the experimental subjects'
-mean and standard deviation, however,
-suggesting that the model
-could have the same variability
-by simply injecting noise into the system.
-This would also result in
-the error trials that are missing in the model.
-
-A better basis for comparison for the model
-would be human performance in a similar
-reaction time task. ?find some % Look up some behavioural data
-
-\subsection{Model principal components match the data}
-
-While it may be unsurprising
-due to the model being inspired by the original
-principal component analysis,
-it is strong validation for the original conceptual model
-that the implemented model maintain
-the same principal components
-while still being able to simulate behaviours.
-
-? List differences and investigate (with final version of model)
-
-\subsection{Model predictions}
-
-The model manipulations assert that blah has the effect of blah.
-
-Prior studies (some unpublished) have suggested that these are indeed the effects of those things.
-
-Cooling has not been studied in an experimental setting,
-and so we take the results as a prediction of the model.
-
 \clearpage
 
-\appendix
-
 \section{Figures}
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f1-pcasummary}
-    \caption{Summary of principal components identified
-      by Narayanan and Laubach, 2009. Three components
-      were identified in motor cortex (top row), accounting
-      for over 60\% of variation.
-      Two components were identified in prefrontal cortex
-      (bottom row), accounting for over 25\% of variation.}
-\end{figure}
+\scalefigone{f1.png}{1.0}{~}
+
+A summary of previous experimental findings. Data from JNP 2009.
+
+A: The simple reaction time task used in the experiment.
+The subject presses and holds down the lever
+for a delay (also called the foreperiod);
+1 second is used in this study.
+After the foreperiod, an auditory tone (the stimulus) is played.
+The subject then has 0.6 seconds (called the response window)
+in which to release the lever.
+If the lever is released in that time,
+the subject is rewarded with juice.
+If the lever is released during the foreperiod,
+it is called a premature error,
+and is penalized by a 5 second time-out. % NOTE: HOW LONG IS TIMEOUT?
+If the lever is not released within the response window,
+it is called a late error,
+and is also penalized by a 5 second time-out.
+
+B: When lever release events occurred.
+Each trial is aligned such that the press event
+occurs at time $t=0$.
+All of the trials shown are correct trials.
+Black dots represent correct trials
+that were preceded by correct trials.
+Red dots represent correct trials
+that were preceded by error trails
+(either premature or late).
+
+C: Peri-event spike rasters and histograms
+showing different neural activity
+based on the previous outcome.
+In all six plots, black represents activity
+when the last trial was correct, and red
+represents activity when the last trial was an error.
+The shaded grey areas note where
+the neural activity is significantly different
+between the two conditions (post-correct and post-error).
+In the top row, there is elevated neural activity
+before the press is initiated, following an error.
+In the middle row, there is elevated neural activity
+before the press is initiated (both)
+and during the foreperiod (right)
+following correct trials.
+In the bottom row, there is elevated neural activity
+during the whole trial (left)
+and after the trial (right)
+following error trials.
+
+D: Linear regression with a GLM was performed
+on the neural activity.
+Before the press is initiated,
+18.7\% of neurons encoded the previous outcome,
+compared to less than 10\% that
+encoded the current reaction time. % WHAT DOES INTERACTION MEAN HERE?
+
+E: Z-statistics showing which neurons were significantly
+modulated by the previous outcome, and when.
+(Top) Sorted by the absolute value of the Z-statistic.
+The bottom 18.7\% of the rows have high average Z-scores
+($p > 0.05$?). % WHAT SAYS THOSE ARE SIGNIFICANT?
+(Bottom) Sorted by the peak of the Z-statistic.
+The line is close to linear, suggesting that the
+previous outcome information is being encoded over time
+by the entire population. % DOES IT SUGGEST THAT?
 
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f2-conceptual}
-    \caption{The conceptual model proposed by Narayana and
-      Laubach, 2009. In the conceptual model, the start of the
-      trial activates two systems; one prepares the response
-      action (prepotent response) and the other inhibits
-      that action until the appropriate time (proactive inhibition).}
-\end{figure}
+\scalefigone{f2.png}{1.0}{~}
+
+A summary of the principal component analysis done
+in JNP 2009;
+the motivation for investigating Singh \& Eliasmith 2006.
+
+A: A graphical summary of principal component analysis.
+Nomarlized peri-event neural data,
+in the form of the z-scores of instantaneous firing rates,
+is organized in a matrix, with each row
+being the z-scored firing rates
+of a single neuron on a single trial.
+Singular value decomposition is performed on the matrix,
+resulting in a matrix such that
+the number of columns (time bins) is the same.
+The rows are now ordered such that the first row
+contains largest eigenvector,
+which represents the value when the original
+data is projected onto the axes of highest variance.
+
+B: The fraction of variance explained
+by each singular value
+when PCA is performed on the post-correct neural data
+(top), and the post-error data (bottom).
+In both cases, the two eigenvectors with the highest
+singular values account for nearly 50\% of variance.
+The next two eigenvectors also account for an amount
+of variance higher than would be expected when
+we interpolate from the smaller
+singular values.
+
+C: A summary of the loadings of the two top
+eigenvector on each neuron.
+If a neuron were encoding both the first and second
+eigenvectors to the same degree,
+we would expect a horizontal line.
+The varied lines criss-crossing each other
+suggests that each individual neuron
+has different sensitivities
+to the first and second principal components.
+
+D: The top three principal components
+for the post-correct trials (black)
+and the post-error trials (red).
+
+In the post-correct case,
+the first principal component is generally high
+during the foreperiod of the task, with bumps
+before and after the foreperiod, and is low elsewhere.
+The second principal component appear to ramp up
+before the task, and down during the task.
+The third principal component is similar
+to the first.
+
+In the post-error case,
+the first principal component is generally high
+before the press event, mostly constant during
+the foreperiod, and low after the trial.
+The second principal component is generally high
+during the foreperiod of the task,
+and closely resembles the first component
+in the post-correct case.
+The third principal component ramps down
+during the task state, and closely resembles
+the second component in the post-correct case.
+
+E: (Top) Plotting the cumulative sum (integral)
+of the first principal component closely matches
+the second principal component, and vice-versa
+($R^2 = 0.904$ and $R^2 = 0.939$ respectively),
+in the post-correct case.
+(Bottom) In the post-error case,
+the first two principal components are no longer
+cumulative sums of each other
+($R^2 = 0.639$ and $R^2 = 0.676$).
+This points to neural integration
+as a potential mechanism explaining mPFC activity.
+
+F: Normalized spike-density functions
+for all of the neurons analyzed in the post-correct case,
+organized by the loading
+(left) on the first principal component, and
+(right) on the second principal component.
+These clearly show neurons whose activity
+matches the calculated principal components.
 
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f3-fullmodel}
-    \caption{The proposed implementation of the conceptual model.
-      The motor cortex is made up of two populations of
-      leaky integrate-and-fire neurons, one of which interacts
-      with the environment by pressing or releasing a lever.
-      The medial prefrontal cortex is made up of three
-      populations of LIF neurons. See text (section 2) for details.}
-\end{figure}
+\scalefigone{f3_blocks.pdf}{1.0}{~}
+
+The Singh \& Eliasmith model (2006).
+This model is created using the methods of the
+Neural Engineering Framework.
+The model uses neural integration in order to explain
+neural data in a working memory task.
+The paper proposes two candidate models
+that can create the desired functional effects
+with different neural dynamics.
+
+(Left) A single-population model.
+In this model, a single population of
+adaptive leaky integrate-and-fire neurons represents
+a four-dimensional signal.
+A recurrent connection causes the signals
+being represented in two of the dimensions
+to integrate; the connection
+also causes one of the integrative dimensions
+to drive the second integrative dimension.
+The remaining two dimensions are used
+to control the integrative dimensions,
+returning them to some baseline activity if necessary.
+
+(Right) A coupled-population model.
+In this model, two populations of
+adaptive leaky integrate-and-fire neurons represent
+two two-dimensional signals.
+Recurrent connections on each population cause
+one of the represented dimensions to be integrated,
+while the other dimension controls the integration.
+An explicit connection between one integrative population
+drives the integrator in the other population.
+Both of the integrative dimensions are then
+projected to a final two-dimensional population,
+which only represents the integrative signals.
+
+In this paper, we hypothesize that the mPFC
+is a realization of one of these two models.
+The integrators are driven by the actions taken by the subjects
+(e.g., lever pressing),
+and are controlled by the outcomes
+(e.g., reward delivery).
 
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f4-behaviour}
-    \caption{Possible outcomes for each trial. If the subject
-      responds within the 0.6 second response window, the trial
-      is counted as correct and reward is delivered. If the
-      subject responds prematurely, it is counted as an error
-      and house lights are extinguished for a timeout period.
-      If the subject responds after the response window,
-      an error is also recorded and the lights are extinguished.}
-\end{figure}
+\scalefigone{f3_pca.pdf}{1.0}{~}
+
+A comparison between the two model realizations
+from Singh \& Eliasmith in this task.
+
+A: Raw spike rasters generated by the two models.
+Modulation around the foreperiod can clearly be seen.
+Differences between the post-correct and post-error
+conditions can also be determined,
+though the difference is more visually discernible
+in the coupled model.
+
+B: Principal component analysis
+on the simulated data (right) shows similar trends
+as the experimental data (left).
+% MORE DETAIL
+
+C: Normalized spike-density functions,
+organized by loadings on PCs (as in the experimental data)
+show neurons that clearly follow the PCs.
+Neurons that do not follow the PCs
+can also be visually identified.
 
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f5-behaviour}
-    \caption{Summary of model and experimental behaviour.
-      (Left) Seconds 200 - 500 of an experimental (top)
-      and model (bottom) session.
-      (Right) Histograms of reactions times for
-      experimental and model subjects.}
-\end{figure}
+\scalefigone{f4_traj.pdf}{1.0}{~}
+
+Plotting the state space (phase space) of the data
+allows us to reason about what the principal components
+may be representing in general.
+
+A: The experimental data's trajectory
+through the phase space formed by the first two
+principal components.
+
+B: The simulated data's trajectory
+through the phase space formed by the first two
+principal components.
+
+C: A mathematical model of the trajectory
+through state space.
+This model is defined by a two-dimensional
+dynamical system that is affected
+by the actions taken by some outside system,
+and the outcomes of those actions.
+
+D: When the mathematical model is implemented
+in a spiking neural model using the NEF,
+we can look at the decoded activity in the
+two integrative dimensions,
+instead of relying on the PCs
 
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f6-mcpca}
-    \caption{Principal components identified in motor cortex
-      spike data. (Top) Experiment. (Bottom) Model.}
-\end{figure}
+\scalefigone{f5_tdcontrol.pdf}{1.0}{~}
+
+Previous work by Laubach et al. (and other groups)
+has implicated mPFC as a top-down controller
+on the motor cortex in delay tasks like simple RT.
+Their work suggests that the mPFC directly inhibits
+the motor cortex; this paper's hypothesis
+is more general, and proposes that the mPFC
+can either inhibit, directly excite,
+or modulate other activity being projected
+to the motor cortex
+depending on the state of the mPFC.
+
+A: A schematic of how top-down control
+could be exerted on a simple stimulus-response
+motor control system.
+The mPFC state is tracked in this two-dimensional system.
+In the green area, which is entered
+by the black trajectory
+(exemplifying correct trials),
+a release is forced to occur.
+In the red area, which is entered
+by the red trajectory
+(exemplifying error trials),
+the subject is not performing the task well,
+and enters a cue-sensitive state,
+in which gain is applied
+to the auditory cue stimulus input.
+
+B: The simple stimulus-response motor control system
+on which top-down control is being exerted.
+This system is shown to perform the simple RT task
+with high accuracy,
+but with a relatively slow reaction time
+(see Fig ???).
+This is due to the system waiting until the cue
+to start the process of initiating the release.
+
+Briefly, the stimulus response system is made up
+of three populations of adaptive leaky integrate-and-fire neurons
+that interact with the environment
+(explicitly modeled in the Nengo simulation environment).
+At the start of a trial, the press-sensitive neurons
+in the \textit{Press/Release} population become active.
+This drives the \textit{Holding} integrator,
+which will maintain a relatively high amount of activity.
+This activity is projected to the \textit{Trigger} neurons,
+which are then primed to cause a release.
+When the trigger stimulus is delivered by the environment,
+it further activates the \textit{Trigger} neurons,
+which require both inputs in order to activate
+the release-sensitive neurons
+in the \textit{Press/Release} population,
+causing the release of the lever
+modeled by the environment.
+
+C: The combined stimulus-response motor control system,
+with the Singh \& Eliasmith coupled-integrators
+providing top-down control.
+Top-down control is exerted through the \textit{Trigger}
+population, allowing it to explicitly force a release
+by providing input in both dimensions,
+or sensitizing and desensitizing the population
+to the trigger stimulus
+by providing input only in the same dimension
+that the \textit{Holding} population projects to.
+Details on the coupled-integrator model
+can be found in figure ???.
 
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f7-pfcpca}
-    \caption{Principal components identified in medial prefrontal
-      cortex spike data. (Top) Experiment. (Bottom) Model.}
-\end{figure}
+\scalefigone{f6_dynamics.pdf}{1.0}{~}
+
+Breaking down the mathematical model
+into what happens in the three cases gives
+an intuitive explanation for how this dynamical system operates.
+
+A: The start of all trials.
+In all cases, no matter what the initial state,
+the system is drive upwards,
+close to [1.0, 0.0]. This sets the initial state
+for the task.
+
+B: In a correct trial, the passive dynamics of the system
+push the state to the right
+at a predictable (i.e., learned) rate.
+The release happens when the state of the system
+is sufficiently far to the right (i.e., $x$ is high).
+If this release happens properly, reward is delivered,
+which pushes the system back to the origin.
+Then, in the absence of any task-related dynamics,
+the system will stay approximately at rest
+because the passive dynamics of the system are very small
+at the origin point.
+
+C: A premature error can be caused by some other factor
+(i.e., a motor system mistake) or by the passive dynamics
+of the system being too strong in the part of the state space
+that the system is driven to once the subject presses the lever.
+In the case of some other factor,
+the system will receive the negative outcome feedback
+before reaching the right side of the state-space,
+and be driven to the bottom.
+In the case of fast integration,
+the system will reach the right side of the state-space,
+force a release, and then receive the negative outcome feedback
+and be driven to the bottom.
+The exact state after receiving the negative feedback
+can be used to figure out the cause of the error,
+though after some time in the lower part
+of the state space, the system will be
+driven to the left,
+and all that can be decoded
+is the notion that an error occurred.
+
+D: A late error can be caused by either another factor
+not responding to the top-down control properly or in time,
+by right-ward drift happening too slowly,
+or by the system not being driven
+to the appropriate place upon pressing down the lever
+and drifting slowly as a result.
 
@@
 
-\begin{figure}
-  \centering
-    % \includegraphics[width=\columnwidth]{f7-pfcpca}
-    Placeholder
-    \caption{Aging results}
-\end{figure}
+\scalefigone{f7_behaviour.pdf}{1.0}{~}
+
+Summary of behaviour in experimental subjects,
+simulated subjects with only the stimulus-response system,
+and simulated subjects with the stimulus-response system
+under top-down control.
+
+A: Experimental results.
+% GIVE TRENDS AND STUFF
+
+B: For the stimulus-response system,
+% GIVE TRENDS AND STUFF
+
+C: For the system with top-down control,
+% GIVE TRENDS AND STUFF
 
@@
 
-\begin{figure}
-  \centering
-    % \includegraphics[width=\columnwidth]{f7-pfcpca}
-    Placeholder
-    \caption{Lesion results}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=\columnwidth]{f7-pfcpca}
-    Placeholder
-    \caption{Cooling results}
-\end{figure}
+\scalefigone{f8_manipulations.pdf}{1.0}{~}
+
+The model can be manipulated in several ways
+in order to investigate the effects
+of various pathologies.
+
+A: By including a damping term in the dynamical system,
+we impair the system's ability to maintain its state.
+This can be visualized as an attractive force
+pulling the system towards the origin.
+This makes it more difficult or impossible
+to confidently exert top-down control,
+depending on the amount of damping.
+This effect can be modeled by degrading
+the recurrent connection weights by a small amount,
+causing the integrative populations to drift
+back to the origin.
+The amount of weight degradation affects
+how quickly the population drifts back to the origin.
+This model realization of this damping
+closely resembles what happens in
+normal aging: recurrent NDMA connections
+lose synaptic strength.
+The model predicts that the new trajectory
+in the state space would more closely resemble
+those in the damped mathematical model.
+
+B: The simulated model can
+instead be manipulated by randomly deleting
+recurrent synaptic connections.
+When this is done,
+the model takes the trajectories shown.
+This predicts what may occur in the mPFC
+in the case of targeted lesions,
+or general trauma such as a concussion.
+
+C: The behavioural effects of these two
+model manipulations when they are done
+on the full model with top-down control.
+% GIVE SOME EXPLANATIONS
 

### 2013-06-10 paper/jneurosci2013.tex
Third version of the paper, with figures.

--- 2013-06-10 paper/jneurosci2013.tex
@@
-%\documentclass[twocolumn]{article}
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 % Change "article" to "report" to get rid of page number on title page
@@
 \usepackage{soul}
+\usepackage[round]{natbib}
 \usepackage[usenames,dvipsnames]{color}
@@
 \usepackage{inconsolata}
-\usepackage{cite}
+\usepackage{hyperref}
+
+\hypersetup{
+  colorlinks = true, %Colours links instead of ugly boxes
+  urlcolor = blue, %Colour for external hyperlinks
+  linkcolor = blue, %Colour of internal links
+  citecolor = red %Colour of citations
+}
 
@@
 \pagestyle{empty}
-
-\newcommand{\framedscalefigone}[3]{
-  \begin{figure}[ht!]
-    % Requires \usepackage{graphicx}
-    \centering
-    \begin{framed}
-      \begin{minipage}{\textwidth}
-        \centering
-        \includegraphics[width=#2\columnwidth]{#1}
-        \caption{#3}
-        \label{#1}
-      \end{minipage}
-    \end{framed}
-  \end{figure}}
-
-\newcommand{\framedscalefigtwo}[4]{
-  \begin{figure}[ht!]
-    % Requires \usepackage{graphicx}
-    \centering
-    \begin{framed}
-      \begin{minipage}{0.49\textwidth}
-        \includegraphics[width=#3\columnwidth]{#1}
-      \end{minipage}
-      \begin{minipage}{0.49\textwidth}
-        \includegraphics[width=#3\columnwidth]{#2}
-      \end{minipage}
-      \caption{#4}
-      \label{#1}
-    \end{framed}
-  \end{figure}}
 
@@
 % Make title
-\title{}
+\title{Adaptive control of action timing with a spiking neural integrator
+  model of anterior cingulate cortex}
 \date{}
-\author{}
+\author{Trevor Bekolay, Chris Eliasmith, Mark Laubach}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{center} \begin{Large}
-  \textbf{A spiking neural circuit model of cortical function
-          in simple reaction time tasks}
-\end{Large} \end{center}
-
-\begin{center}
-  Trevor Bekolay, Benjamine Liu, Nandakumar Narayanan, Chris
-  Eliasmith, Mark Laubach
-\end{center}
+\maketitle
 
 \begin{abstract}
-  todo
+  Animals can learn to perform a simple reaction time task
+  with a predictable foreperiod
+  by predicting the time of the cue,
+  achieving reaction times faster
+  than through cue-responding alone.
+  The outcome of the previous trial
+  is one factor in determining
+  whether an animal will attempt to time
+  the foreperiod.
+  The anterior cingulate cortex (ACC) has been implicated
+  in enabling control of action timing
+  depending on the outcome of the previous trial.
+  Analysis of ACC activity suggests that
+  neural integration is a key mechanism
+  for adaptive control in precisely timed tasks.
+  We show that a spiking neural circuit
+  consisting of coupled neural integrators
+  captures the neural dynamics
+  of the experimentally recorded ACC.
+  We use this coupled integrator model
+  to adaptively control
+  a spiking neural circuit that follows
+  a cue-response strategy,
+  and show that it performs the simple reaction time task
+  with similar reaction times as experimental subjects.
+  We then alter the model
+  in a way consistent with
+  how the aged brain degrades,
+  and predict that aged subjects
+  will adopt a cue-response strategy.
 \end{abstract}
@@
 
-Animals make complicated motor actions
-with high spatial and temporal resolutions.
-In temporally precise tasks,
-the motor cortex and several afferents coordinate
-to bring about motor actions at appropriate times.
-For tasks with delay periods,
-the medial prefrontal cortex is critically important
-for suppressing motor action during the delay period.
-This effect has been seen in ?.  %???primates humans rodents etc???.
-
-In rodents, the role of the medial prefrontal cortex
-has been investigated through ?.
-%???previous work???.
-Previously, we have used a simple reaction task
-and principal component analysis
-to investigate ?. %???summarize Narayana \& laubach jnp 2009???
-These results point to neural integration
-as playing a key role
-in the mPFC's ability
-to control the timing of motor actions.
-
-In order to investigate the neural circuit
-underlying the production and suppression of motor output,
-we extend Singh \& Eliasmith's model to include
-neural populations that directly affect the environment,
-analogous to the function of Betz cells.
-In doing so, we can relate the model to the experimental data
-in both neural and behavioural contexts.
-
-\section{Model}
-
-The model presented in this study
-is based on a conceptual model of PFC and MC
-function in delay tasks proposed by Narayanan and Laubach.
-Their conceptual model is based on
-principal component analysis of neural data
-in rats performing a simple reaction time task.
-
-Figure 1 summarizes their PC analysis.
-In PFC, two components account for over 35\%
-of variance in population firing.
-The first principal component increases slightly
-at the time of the lever press,
-and then decreases over the delay period.
-The second principal component
-also increases at the time of lever press,
-but remains high until the time of release.
-
-In motor cortex, three components account
-for over 60\% of variance in population firing.
-The first two principal components are
-similar to those found in PFC;
-the first steadily decreases over the delay period,
-and the second remains high during the delay period.
-The third principal component increases
-sharply slightly before lever press
-and lever release.
-
-The identification of these principal components
-leads to a conceptual model of cortical function
-during simple reaction time tasks.
-This model is summarized in figure 2. % Fig
-At the start of a trial,
-two complementary systems are activated
-to produce quick reactions at the time of a stimulus.
-The \textit{prepotent response} system
-prepares the motor action that the subject will perform
-at the end of the delay period.
-The \textit{proactive inhibition} system
-suppresses that motor action until the delay period
-is over.
-
-\subsection{Neural integration as a core mechanism}
-
-An important observation of Narayanan and Laubach
-is that the second principal component
-is the cumulative sum (integral)
-of the first principal component.
-This suggests that some neurons in these cortical areas
-are representing the integral of another
-signal represented in these areas.
-
-Neural integration through recurrently connected
-populations of neurons has been widely studied
-in the context of working memory.
-% ? cite a bunch -- Mawkinson, Roman, Tanji, X-J Wang, look at singh eliasmith biblio
-
-The Singh and Eliasmith model, in particular,
-facilitates a neural implementation of the conceptual model
-proposed as a result of the PC analysis.
-In addition to outlining a method of constructing
-multidimensional neural integrators,
-it also provides a method of connecting those integrators
-to other populations of neuron representing other signals.
-
-\subsection{Coupled integrators in a control system}
-
-With non-integrative and integrative neuronal populations
-as the building blocks, we propose a neural circuit
-realization of the conceptual model
-discussed in Narayanan and Laubach.
-The model is summarized in figure 3.
-
-Two populations are hypothesized to reside in motor cortex.
-The \textit{Press/Release} population represents
-whether the subject is currently executing
-a press or release motor action.
-The activity of this population directly affects the environment
-in the model, which can be thought of
-as functioning like Betz cells.
-The \textit{Holding} population tracks
-whether or not the subject is currently holding down the lever.
-It is analogous to the \textit{prepotent response}
-from the conceptual model, as in the absence of inhibition,
-it will activate the \textit{Release} neurons
-in the \textit{Press/Release} population.
-The \textit{Holding} population is integrative,
-and integrates the
-\textit{Press} signal from the \textit{Press/Release}
-population. The integration is controlled
-by the \textit{Release} signal, meaning that
-when the lever is released the \textit{Holding} signal
-is reset to some baseline.
-
-Three populations are hypothesized to reside
-in medial prefrontal cortex.
-The \textit{Waiting} population represents
-that the subject is in a delay state,
-waiting for the appropriate time to perform a motor action.
-It is analogous to the \textit{proactive inhibition}
-from the conceptual model, as it works to balance
-the prepotent release response being signaled
-from the motor cortex's \textit{Holding} population.
-The \textit{Cue} population is sensitive
-to the auditory stimulus delivered to the subject
-when the delay period is over.
-It inhibits the \textit{Waiting} population's
-proactive inhibition, disinhibiting \textit{Release}
-sensitive neurons.
-Similarly, the \textit{Timer} population
-tracks the relative amount of time that has passed
-since the subject pressed down the lever,
-starting the trial.
-When the time of the stimulus can be predicted,
-the \textit{Timer} population speeds
-responding by increasingly inhibiting the
-\textit{Waiting} population over time.
-It can also effect a motor action on its own by
-disinhibiting \textit{Release} neurons sufficiently.
-The \textit{Timer} population is integrative,
-gradually ramping up as it integrates
-the \textit{Waiting} signal.
+The anterior cingulate cortex has been implicated
+in a wide variety of functions,
+including conflict monitoring
+\citep{Aarts2009, Botvinick2004, VanVeen2004},
+error monitoring \citep{Brown2005, Carter1998, Fecteau2003,
+  Modirrousta2008, Rushworth2004},
+anticipatory control \citep{Aarts2008, Koyama2001},
+arousal of the sympathetic nervous system
+\citep{Critchley2003, Luu2003},
+and control of action timing \citep{Muir1996, Mulert2003,
+  Naito2013, Narayanan2009, Risterucci2003}.
+Many have proposed integrative or general models
+that would allow the ACC to be involved
+in several of these functions,
+depending on the behavioral context
+\citep{Botvinick2004, Luu2003}.
+However, while there have been
+attempts at modeling the ACC at the level
+of artificial neural networks
+\citep{Botvinick2004, Brown2005},
+there are currently no mathematical
+accounts of ACC function that would
+explain the myriad of behaviors
+modulated by the ACC.
+Similarly, no existing models
+have been implemented with spiking neurons,
+and are therefore not directly comparable
+to neural activity.
+
+In this paper, we propose that a
+two-dimensional dynamical system
+can capture the activity of the ACC.
+We test the model in a simple reaction time task
+with a predictable foreperiod and
+show that the mathematical model
+enables precise timing of actions,
+and changes behavior depending
+on the outcome of the last trial.
+We implement the dynamical system
+in a spiking neural network
+based on \citet{Singh2006}
+and show that simulated neural activity
+closely matches recorded ACC activity.
+We show that the model
+can be combined with
+a motor control model,
+resulting in behavioral changes
+analogous to previous behavioral theories
+of ACC function.
+Finally, we damage the model in a way analogous
+to aging, and predict that
+aged subjects will adopt a more conservative
+strategy due to a reduction of the
+ability of the ACC to control action timing.
+
+Our mathematical account of ACC function
+is applied directly to
+error monitoring and
+the control of action timing.
+However, we assert that the dynamical systems
+approach is general and can be used
+to explain other theories
+of high-level ACC function.
+
+\section{Simple reaction time task} \label{sec:simple-rt}
+
+The simple reaction time task
+serves as a minimal task in which
+to examine the control of precisely timed actions.
+In this task, the subject presses and holds down a lever
+for a fixed amount of time (the ``foreperiod'';
+1 second for this study).
+After the foreperiod, a cue is presented
+(e.g., an auditory tone).
+The subject then has 0.6 seconds (the response window)
+in which to release the lever.
+If the lever is released in that time,
+the trial is a success, and the subject is rewarded.
+If the lever is released during the foreperiod,
+the trial is classified as a ``premature error,''
+and is penalized by a time-out.
+If the lever is not released within the response window,
+the trail is classified as a ``late error,''
+and is also penalized by a time-out.
+Figure~\ref{fig1} presents a schematic view of the task.
+
+\section{Analysis of experimental data}
+
+\citet{Narayanan2009} recorded
+the anterior cingulate cortex (ACC) and motor cortex
+of rodents performing this simple reaction-time task.
+They used principal component analysis
+on normalized spike densities
+in order to determine
+what is being represented in these brain areas
+during this task.
+As shown in Figure~\ref{fig2},
+the first two principal components
+found from analyzing the ACC data are large,
+and are hypothesized to track
+1) if the subject was performing the task, and
+2) the relative amount of time the subject has been
+performing the task.
+\citeauthor{Narayanan2009} also found that the neural activity
+in the ACC changed significantly depending on
+the outcome of the previous trial;
+the first two principal components
+change significantly after errors,
+and appear to contain information that an error occurred.
+This information would be necessary
+in order to adapt behavior based on the last trial.
+
+Interestingly, in the post-correct case,
+the first principal component
+is highly correlated with the integral
+of the second principal component,
+and vice versa.
+In the post-error case,
+these correlations are significantly weaker.
+This result suggests that
+the ACC is performing some kind of integration..
+
+\section{Double integrator model}
+
+One model of neural integration
+was proposed in the context of working memory
+by \citet{Singh2006}.
+Their model, called the \textit{double integrator} model,
+uses multidimensional neurons
+in integrative populations
+to exhibit time-varying activity
+during a vibrotactile discrimination task.
+It relies on multiple coupled integrators,
+whose activity resembles
+the two strong principal components
+in the ACC data.
+The model employs
+the Neural Engineering Framework (NEF; \citealp{Eliasmith2003})
+to implement a spiking neural network
+that captures the dynamics of the
+integrative dynamical system
+(see section \ref{sec:nef}).
+We hypothesize that the \citeauthor{Singh2006} model
+will exhibit the same neural dynamics
+as the experimental ACC in the simple RT task.
+
+\subsection{Dynamics}
+
+The dynamics of the double integrator model
+are captured in the following dynamics equations.
+\begin{align*}
+  \dot{x}_1 &= u \\
+  \dot{x}_2 &= \alpha x_1,
+\end{align*}
+where $u$ represents an input signal
+and $\alpha$ is the strength of
+the connection between the two integrators.
+This results in a two-dimensional system
+in which $x_1$ integrates its input,
+maintaining the accumulated value over time.
+$x_2$ integrates the value of $x_1$.
+\citet{Singh2006} have proposed that this system
+enables time tracking;
+because $x_2$ integrates $x_1$,
+the amount of time elapsed since
+$x_1$ changes can be determined
+based on how much $x_2$ has changed.
+
+In the simple reaction time task,
+the input, $u$, is the activation of neurons
+that effect the lever press.
+The system is also modified by signals
+representing the outcome of a particular trial;
+on a correct trial, reward delivery ($R > 0$)
+resets the system to a starting state.
+On an error trail, a second input signal ($E < 0$)
+drives $x_1$ to a low state.
+When implemented in a spiking network,
+this causes $x_2$ to saturate at a low value,
+effectively breaking the integration,
+as is seen in Figure~\ref{fig3} in post-error trails.
+This results in the following dynamics.
+\begin{align} \label{dint}
+  \dot{x}_1 &= u + E - R x_1 \nonumber \\
+  \dot{x}_2 &= \alpha x_1 - R x_2.
+\end{align}
+See Figure~\ref{fig4} for a graphical
+depiction of the double integrator model.
+
+In addition to the depicted system,
+we inject a slowly oscillating signal into $x_1$.
+This is motivated by the observation
+that system \eqref{dint} is stationary
+during the intertrial interval;
+$x_1$ is either near 0 on rewarded trials
+or some low value on error trials.
+$x_2$ does not change on rewarded trials,
+and saturates at some low value on error trials.
+Figure~\ref{fig3} shows some activity during
+the intertrial interval;
+we model this additional activity
+by a simple oscillation, $O = \sin(2 \pi f t + \phi)$,
+where the frequency $f$ is low ($\le 0.4$Hz).
+The final dynamics are
+\begin{align} \label{dint-full}
+  \dot{x}_1 &= u + E + O - R x_1 \nonumber \\
+  \dot{x}_2 &= \alpha x_1 - R x_2.
+\end{align}
+
+Note that while there is experimental evidence for
+a slow oscillations like $O$ in various brain regions
+\citep{Buzsaki2004}, this injected oscillation
+does not affect the performance of the network,
+and is intended to model ACC activity
+not being otherwise modelled in the current study.
+This activity could reflect
+other functions that the ACC is performing,
+or could reflect the overall pace
+of the simple RT task.
+
+\section{Adaptive control circuit}
+
+\citet{Narayanan2009} hypothesized
+that the ACC activity analyzed in the present study
+could be used to inhibit motor responses
+until the time of the cue.
+In order to verify that hypothesis,
+we constructed a network that
+performs the simple RT task
+using the simple non-adaptive strategy
+of responding to cues.
+We then combined that network
+with the double integrator network
+such that the final model
+is able to flexibly switch between
+an aggressive timing strategy
+and a conservative cue-response strategy
+depending on the outcome of the previous trial.
+
+\subsection{Cue-responding circuit}
+
+The cue-responding circuit is composed
+of three populations
+of adaptive leaky integrate-and-fire neurons
+(see Figure~\ref{fig4}).
+The population labelled \textit{Trigger}
+is sensitive to the cue.
+The population labelled \textit{Holding}
+tracks whether the lever is currently being held down.
+The population labelled \textit{Press/Release}
+causes lever presses and releases
+through activation of different subsets
+of its neurons.
+The actual lever press and release
+is delayed by $\sim$250ms,
+the approximate amount of time
+it takes for muscles to respond
+to the activation of command neurons
+in primary motor cortex.  % ???travis
+
+The \textit{Holding} neurons project
+to the \textit{Trigger} population
+to provide the behavioral context
+(i.e., \textit{can} I release the lever?)
+The \textit{Trigger} neurons send projections
+to the \textit{Press/Release} neurons
+upon a cue that occurs in the correct context
+(i.e., \textit{should} I release the lever?)
+
+The context-driven design of this circuit
+provides an alternative theory to
+the prepotent inhibition model
+proposed by \citet{Narayanan2009}.
+They suggest that a motor command
+is activated at the same time
+as an inhibitory waiting signal,
+and it is the removal of that inhibition
+that produces the motor command.
+This model proposes that
+an excitatory context signal is activated,
+and the pairing of that context signal
+with the cue produces the motor command.
+This leads to an alternative hypothesis
+for the function of ACC activity:
+the ACC tracks the behavioral context,
+and is able to inhibit or excite motor cortex,
+but does not always necessarily do so.
+
+\subsection{Adaptive control circuit}
+
+A weakness of the contextual cue-responding circuit
+is that there is a fixed lower-bound
+on the reaction time
+based on the time taken to detect the cue
+and for the delay between activation
+of motor cortex and muscle activity.
+Since the goal is to time the action
+as close as possible to the time of the cue,
+the cue-responding strategy may be too slow,
+and so a predictive approach must be employed.
+
+The double integrator model described
+previously can implement such a strategy.
+The state of the dynamical system being tracked by
+the double integrator (equation \eqref{dint-full})
+is approximately the same when the cue occurs.
+By connecting the output of the double integrator
+to the \textit{Trigger} population,
+the double integrator can cause a lever release
+when the state of the double integrator
+is slightly before the predicted
+time of the cue.
+
+However, if the previous trial was an error,
+the double integrator will not
+be following the trajectory
+that predicts the time of the cue,
+and therefore will not cause a lever release.
+The circuit then naturally
+uses the conservative
+cue-responding strategy.
+Switching between these two strategies
+depending on the outcome
+of the previous trial
+allows for adaptive control of action timing.
+
+\section{Emulating aging effects}
+
+The spiking neural implementation
+of the double integrator model can be used
+to emulate the effects of several pathologies.
+As a proof of concept,
+we emulate the effects of NMDA receptor degradation
+in aging \citep{Magnusson1998}.
+
+Feedforward connections
+in the adaptive control model (Figure~\ref{fig4})
+use a time constant of 10 ms,
+which is consistent with the time constant
+of the postsynaptic current
+induced by uptake of neurotransmitter
+by AMPA receptors \citep{Spruston1995}.
+Recurrent connections use a time constant of 50 ms,
+which is consistent with the PSC
+of NMDA receptors \citep{Sah1990}.
+To emulate the effects of aging on NMDA receptors,
+we decrease the synaptic strengths
+on these recurrent connections.
+
+We hypothesize that this will have the same effect
+as adding a small damping term, $D < 0$,
+in equation \eqref{dint-full}, resulting in
+\begin{align} \label{dint-damped}
+  \dot{x}_1 &= u + E + O - R x_1 - D x_1 \nonumber \\
+  \dot{x}_2 &= \alpha x_1 - R x_2 - D x_2.
+\end{align}
+This hypothesis can be tested directly
+in the double integrator model,
+and its effects on behavior tested
+in the adaptive control circuit.
+
+\section{Theoretical results}
+
+\subsection{Model dynamics during the simple RT task}
+
+The dynamical system described by equation \eqref{dint-full}
+exhibits the desired behavior during simulation
+of the simple RT task.
+Figure~\ref{fig5} provides a breakdown of
+the trajectory and vector field
+throughout a correct, premature, and error trial.
+Note that the oscillation is omitted
+in this figure.
+The final position of
+the trajectory becomes the starting position
+of the next trial, and therefore
+provides information about
+the outcome of the last trial.
+It is clear from these trajectories
+that the time of the cue
+can be predicted,
+as the ACC is only in the predictive state
+when the cue occurs on correct trials.
+It is also clear that the
+outcome of the previous trials
+is easily decoded based
+on the state after the previous trial.
+Note that premature trials
+and late trials cannot be distinguished
+on the next trial;
+this is consistent with the results of
+\citet{Narayanan2009}.
+
+\subsection{The aged model loses adaptivity}
+
+Figure~\ref{fig6} shows trajectories for
+correct, premature, and late trials
+for the aged model (i.e., the model with a damping term,
+equation \eqref{dint-damped}).
+As the damping term increases, the final position
+of the trajectories on error trials
+becomes closer to the final position of correct trials,
+indicating that increased damping
+makes distinguishing between correct and error trials
+progressive more difficult.
+Additionally, the distance (in state space) between
+the state at which the cue occurs
+and other states visited in the trajectory
+decreases with increased damping,
+which results in more difficulty
+predicting the time of the cue
+following correct trials as well.
+Timing the foreperiod is therefore
+a poor strategy with significant aging
+(i.e., with $D \le 0.05$).
+
+\section{Simulation results}
+
+\subsection{Spiking networks follow predicted trajectories}
+
+Figure~\ref{fig7} shows that
+the decoded values of a spiking neural network
+with the same dynamics as equation \eqref{dint-full}
+also take a similar trajectory through
+state space in correct, premature, and late trials.
+Despite the introduction of the oscillation
+and the additional noise present
+in spiking neural systems,
+behaviorally relevant states
+are significantly separated
+in the decoded state space of the spiking model.
+Therefore, the cue can be accurately predicted,
+and the outcome of the previous trial
+has a significant effect
+on the trajectory taken during a trial.
+
+\subsection{Decreasing recurrent weights is analogous to damping}
+
+Figure~\ref{fig8} shows that
+the decoded values of a spiking neural network
+in which recurrent connections are degraded
+by a small amount
+take similar trajectories through state space
+as the damped model (Figure~\ref{fig6}).
+This supports the hypothesis
+that a small damping term
+has a similar effect as
+the degraded NMDA receptors seen in aging
+in the double integrator model.
+
+\subsection{Double integrator model has the same principal components as experimental data}
+
+In order to perform an identical
+principal component analysis
+on the simulated data as was done
+on the experimental data,
+we randomly sampled 174 neurons
+with firing rates $> 1$ Hz from
+the spiking double integrator model.
+The spike trains of these neurons
+were convolved with a Gaussian filter ($\sigma = $ 25 ms)
+and then normalized to Z-scores.
+The results are shown in Figure~\ref{fig9}.
+The simulated principal components
+are very similar to the experimental principal components;
+$R^2 = 0.96$ and $R^2 = 0.895$ for
+the first two components in the post-correct case,
+$R^2 = 0.979$ and $R^2 = 0.877$ for
+the first two components in the post-premature case,
+and $R^2 = 0.967$ and $R^2 = 0.862$ for
+the first two components in the post-late case.
+
+\subsection{Adaptive control model has similar performance as experimental subjects}
+
+The parameters of the adaptive leaky integrate-and-fire neurons
+used in all models are randomly assigned
+from biologically determined distributions.
+While the differences in neural parameters
+has little effect on the principal components
+of the double integrator model,
+they can have a significant effect
+on performance and reaction times
+in the simple RT task.
+Similarly, individual experimental subjects
+have varying performance levels and reaction times
+in the task.
+For these reasons, we analyze
+each experimental subject and instance of the model separately.
+
+Figure~\ref{fig10} shows the performance
+of the experimental subjects,
+the model using only the cue-responding strategy,
+and the full adaptive model.
+The length of the simulations were
+varied to approximately match the number of trials
+as the experimental subjects.
+The mean of the median reaction times
+is 272$\pm$48ms for the experimental subjects,
+368$\pm$34ms for the cue-responding models,
+and 260$\pm$100ms for the adaptive control model.
+The mean performance (\% correct trials) is
+70.8\% for experimental subjects,
+96.9\% for cue-responding models,
+and 84.0\% for adaptive control models.
+Therefore, the adaptive control model
+has similar reaction times as the experimental subjects.
+The cue-responding model is significantly slower,
+but makes fewer errors than the adaptive control model.
+
+\subsection{Aged model is slower and less variable}
+
+When the aged double integrator model
+is used in the adaptive control network,
+the mean performance is 89.5\%,
+and the mean median reaction time
+is 375$\pm$47ms (see Figure~\ref{fig10} for individual results).
+These results are closer
+to the results of the cue-responding strategy,
+suggesting that the model emulating
+some effects of aging
+uses the conservative cue-responding strategy
+over the time prediction strategy
+more often than the non-aged adaptive control network.
+
+\section{Discussion}
+
+The results indicate that the double integrator model
+is allowing for adaptive control of action timing.
+The ability to predict the time of the cue
+produces faster reaction times at the cost
+of additional premature errors.
+The model that enables this adaptive control
+has very similar neural dynamics
+as the experimental ACC.
+
+One important difference between
+the experimentally recorded and simulated ACC
+is in the amount of variance
+explained by the principal components
+and the lack of noise in the
+spike density function (see Figure~\ref{fig9}).
+The relative cleanness of the simulated data
+is expected for several reasons.
+Primarily, we expect that the ACC
+is sensitive to many other factors
+than are represented in the simulation.
+The simulated data is more analogous
+to the 174 neurons in the entire ACC
+that are modulated the most during the simple RT task,
+rather than the 174 neurons that happened
+to be recorded during the experiments.
+This could be simulated
+by either injecting
+noise into the simulation,
+or by representing additional
+randomly varying dimensions.
+Additionally, the adaptive
+leaky integrate-and-fire neuron model used in this study
+has a more regular spiking pattern than
+biological neurons,
+so using a more detailed neural model
+would provide a closer match
+between experiment and simulation.
+
+An interesting feature of the model
+is that there is significant randomness
+involved in model generation,
+and no learning occurs once the model is generated.
+While each instance of the model
+can be thought of an individual
+from the population of subjects,
+one can also think of each model instance as representing
+a snapshot during a subject's
+learning of the task.
+Therefore, the poor performance of, for example,
+adaptive control model instance 10 (see Figure~\ref{fig10})
+can be explained at least two ways.
+It may be a subject
+whose recurrent connections are too strong,
+and therefore is a poor performer in the task,
+or it may be a subject
+who has not yet learned how to properly time
+the interval, and after more trials would,
+with learning, be able to make less premature errors.
+
+The results from the aged model
+form an important prediction of
+how aged subjects will perform in the simple RT task.
+It is notable, however, that the aged model
+does not make additional premature errors;
+in fact, premature errors are completely eliminated.
+This is due to the targeted manner in which
+aging is emulated (i.e., NMDA receptor degradation).
+If the model is more generally damaged,
+for example to emulate an ACC inactivation study,
+we would predict that the effects
+would be different than in the aging example.
+In the case of overall ACC damage,
+we hypothesize that the effect is more similar
+to decreasing the ``resolution'' of the state space.
+That is, it becomes harder to differentiate
+similar states, though the system
+still takes the same approximate trajectory
+through the state space.
+In this case, it becomes difficult
+to predict the time of the cue,
+and therefore additional premature errors would occur.
+
+While we have tested our mathematical model
+of ACC function in the context of
+control of action timing in a simple RT task,
+the tracking of behavioral variables
+in a dynamical system could be used
+to perform many of the other high-level functions
+that the ACC is theorized to perform.
+For example, in the commonly studied Stroop task,
+the state space being tracked by
+the ACC could be the difference between
+the response suggested by word-reading and the
+response suggested by the color;
+the context that colors should be read
+would be reflected in the passive dynamics
+of the ACC, which would push the ACC
+to increasingly drift towards a state
+that compensates for the word-reading response.
+
+In conclusion, we have shown that neural integration
+is a key mechanism in the ACC's
+ability to control the timing of actions.
+Analysis of the activity of a neural integrator
+model closely matches that of the experimentally recorded ACC.
+That model is sufficient to allow
+a simple cue-responding network
+to react more quickly at the cost
+of additional errors,
+which matches the performance of experimental subjects.
+When the model is manipulated
+in a similar way as the aged brain degrades,
+performance degrades and forms a prediction
+of how aged experimental subjects may perform.
+
+\bibliographystyle{plainnat}
+\bibliography{jneurosci2013}
+
+\clearpage
+
+\appendix
+
+\section*{Supplementary material}
 
@@
 
-In both the empirical and simulated data,
-subjects and models performed the same simple reaction time task.
-At the start of a trial,
-subjects press down a lever.
-A timer tracks the time since the lever was pressed down,
-and after an experimenter-specified time period,
-called the \textit{delay period} or \textit{foreperiod},
-an auditory cue signals that the subject must respond
-by releasing the lever.
-If the lever is released within 600 milliseconds,
-the trial is considered a success
-and reward is delivered.
-If the lever is released before the auditory cue
-or after the 600 millisecond response window,
-then the trials is considered a failure
-and subjects are given a time out in which
-lights are extinguished.
-Lever presses during this timeout period
-lengthen the timeout period.
-
-Empirically, this study is done with rats
+The simple reaction time task
+(section \ref{sec:simple-rt}; Figure~\ref{fig1})
+is implemented both experimentally
+and in a simulated environment.
+
+Experimentally, Long-Evans rats perform the task
 in a MedPC box containing a lever,
@@
 Theoretically, this study was performed
-in the Nengo simulation environment.
-Subjects were networks of leaky integrate-and-fire
-neurons, as described below.
+in the Nengo simulation environment.\footnote{Nengo is open source
+  software available at \url{http://www.nengo.ca}.}
 The environment was modeled using a
-\texttt{simplenode}, which allow arbitrary code
-to interact with a neural model.
-In this case, the \texttt{simplenode} implemented
 a finite state machine to track task state,
-and scalar values to indicate the state of
+and scalar values between -1 and 1
+to indicate the state of
 house lights, lever position, auditory cue, and reward.
 
-\subsection{Empirical summary}
-
-The empirical study (described in more depth in ?) %???cite???)
-was done with twelve male Long-Evans rats (3-4 months old; Harlan).
-Rats were water restricted, with food always available,
-and consumed 10-15 ml of water during each
-behavioural session.
-
-Arrays of recording electrodes were implanted
-in dmPFC ($n = 8$), motor cortex ($n = 3$), or both ($n = 1$).
-Electrodes were 50 $\mu$m stainless steel wires
-with 250 $\mu$m between wires.
-Arrays were in configurations of
-4 $\times$ 4 or 3 $\times$ 3 $\times$ 2.
-Microwire arrays were targeted at dmPFC
-based on the following coordinates from bregma:
-anteroposterior [AP]: $+$3.2mm;
-mediolateral [ML]: $\pm$1.4mm;
-dorsoventral [DV]: $-$3.6mm at 10$^\circ$ in the frontal plane.
-Motor cortex was targeted with
-AP: $-$0.5mm; ML: $\pm$2.5-3.5mm;
-DV: $-$1.5mm at $-25^\circ$ in the frontal plane.
-Histology was performed to confirm electrode locations
-(???see fig???).
-
-Ensemble recordings were done using a
-Many Neuron Acquisition Program (Plexon, Dallas, TX).
-Spike sorting using principal component analysis (PCA)
-and waveform shape identified 414 single units
-whose spikes were included in this study.
-
-\subsection{Theoretical}
-
-Two models were created to explain neural processing
-in the simple reaction time task.
-The first is a double-integrator model,
-as is described in ?, %???cite singh eliasmith???,
-which allows for precise comparisons
-between the spiking behaviour of
-the model and the empirical subjects
-given a controllable behavioural outcome.
-The second is an integrated model that
-expands the double-integrator model
-to include populations that manipulate the environment,
-allowing for the comparison between
-the model and empirical behaviour.
-
-Both models are made up of heterogeneous populations
-of leaky integrate-and-fire neurons
-created and connected using the methods
-of Eliasmith and Anderson ?. % ???cite???.
-Connections between populations of neurons
-are determined based on the function computed
-through the connection.
-Connection weights are determined
-by solving a least-squares minimization
-of the function representation
-given the tuning properties
-of the neurons representing the function parameters,
-as described in ? %???singheliasmith???
-and in detail in ?. %???eliasmith anderson ???.
+\subsection{Experimental data}
+
+The experimental data analyzed in this study
+is the same data presented in
+\citet{Narayanan2009}.
+The data consists of 174 neurons
+recorded from anterior cingulate cortex with microwire arrays
+in twelve male Long-Evans rats.
+For more details on the data collection,
+refer to the original paper.
+
+\subsection{Simulation details}
+
+All simulations were run
+in the Nengo simulation environment
+through Python scripts.
+Those scripts are available
+at \url{https://github.com/tbekolay}.
+
+\subsubsection{Neural Engineering Framework} \label{sec:nef}
+
+Nengo simplifies the construction
+of spiking neural networks
+using the Neural Engineering Framework (NEF; \citealp{Eliasmith2003}).
+The NEF makes the assumption
+that populations of spiking neurons
+represent real-valued vectors,
+and use a least-squares optimal method
+to transform those representations
+through the connections between
+populations of neurons.
+A population can also be recurrently connected
+in order to produce dynamics,
+such as the neural integrators
+employed in this study.
+
+Briefly, the representation scheme
+is similar to population coding,
+as proposed by \citet{Georgopoulos1986}
+but extended to $n$-dimensional vector spaces.
+Each neuron in a population is sensitive
+to a particular direction,
+called the neuron's \textit{encoder}.
+The activity of a neuron can be expressed as
+\begin{equation}
+	a = G[\alpha \mathbf{e} \cdot \mathbf{x} + J_{bias}],
+\end{equation}
+where $G[\cdot]$ is the nonlinear neural activation function,
+$\alpha$ is a scaling factor (gain) associated with the neuron,
+$\mathbf{e}$ is the neuron's encoder,
+$\mathbf{x}$ is the vector to be encoded, and
+$J_{bias}$ is the background current of the cell
+when $\mathbf{x} = 0$.
+The currently encoded value, $\mathbf{\hat{x}}$,
+can be estimated linearly.
+\begin{equation}
+  \mathbf{\hat{x}}(t) = \sum_i \mathbf{d}_i a_i(t),
+\end{equation}
+where $\mathbf{d}_i$ is the decoder,
+and $a_i$ is activity of neuron $i$.
+
+Neural activity is interpreted as a
+filtered spike train.
+\begin{equation}
+  a_i(t) = \sum_s h(t - t_s) = \sum_s e^{-(t - t_s) / \tau_{PSC}},
+\end{equation}
+where $h(\cdot)$ is the exponential filter
+applied to each spike,
+and $s$ is the set of all spikes occurring
+before the current time $t$.
+
+The decoders are found through a least-squares minimization
+of the difference between the decoded estimate
+and the actual encoded vector.
+\begin{equation}
+  \mathbf{d} = \Upsilon^{-1} \Gamma \hspace{1.8em}
+  \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
+  \Upsilon_j = \int a_j \mathbf{x} dx.
+\end{equation}
+
+Connection weights computing a function
+between two populations connected in a feedforward manner
+$f(\mathbf{x})$ can be determined by
+solving for a set of decoding weights
+for that function,
+\begin{equation}
+  \mathbf{d}^{f(\mathbf{x})} = \Upsilon^{-1} \Gamma \hspace{1.8em}
+  \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
+  \Upsilon_j = \int a_j f(\mathbf{x}) dx,
+\end{equation}
+and then computing the following weight matrix, $\omega$;
+\begin{equation}
+	\omega_{ij} = \alpha_j \mathbf{e}_j L \mathbf{d}^{f(\mathbf{x})}_i,
+\end{equation}
+where $i$ indexes the input population,
+$j$ indexes the output population,
+and $L$ is a linear operator.
+
+Dynamical systems of the form
+$\dot{\mathbf{x}} = A(\mathbf{x}) + B(\mathbf{u})$
+can be created by connecting a population
+to itself with a weight matrix
+that computes the function $f(\mathbf{x}) = \mathbf{x}$,
+setting $L = A$ on the recurrent connection,
+and receiving input $\mathbf{u}$
+from other neural populations
+and setting $L = B$ on those connections.
 
@@
 
-In  %???cite???
-it is shown that
-a population of heterogeneous adaptive leaky integrate-and-fire neurons
-representing a two-dimensional signal
-can capture the dynamics of
-working memory cells in prefrontal cortex.
-There are, however, several network configurations
-that can result in a population of neurons
-representing two integrated signals,
-as shown in figure ?. %???copy fig 5 from paper???.
-
-We adopt the coupled one-dimensional integrator model,
-despite little evidence that reciprocal projections
-are localized to smaller integrative populations.
-Instead, it is the PCA of mPFC activity
-that suggests many neurons are
-only sensitive to one of the principal components.
-Additionally, it is more intuitive
-to connect each principal component with
-a separate neural population.
-The important task, then,
-is to connect each population
-with the appropriate value being encoded.
-
-Figure ? %???put in???
-summarizes our mapping.
-The two coupled one-dimensional oscillators represent
-the subject's existence in a particular behavioural state
-and how long the subject has been in that state.
-The behavioural state is entered
-once the subject presses down the lever,
-modeled by the activation
-of a population of press-sensitive neurons.
-The timing integrator is driven
-by the state population, as follows from ?. %???citesingh???.
-Both coupled integrators are controllable
-through a second dimension, %(???cite eliasmith here? ???),
-which allows the integrators to be
-reset to an initial point (0.0)
-when reward is delivered to the subject.
-The two one-dimensional integrators
-are then projected to an output population that
-tracks both signals.
-
-The press- and reward-sensitive populations
-drive the rest of the activity of the network.
-These two populations are driven by the experimenter,
-based on the behaviour being simulated.
-For example, for a sequence of correct trials,
-the press-sensitive population
-will be driven strongly at the start of a trial,
-followed 1.6 seconds later
-by strongly driving the reward-sensitive population.
-For error trials, the reward population
-is not driven between activations of the press population.
-
-\subsubsection{Integrated motor-mPFC model}
-
-In order to expand the double-integrator model
-to directly manipulate elements of the environment,
-three additional variables must be accounted for:
-release, auditory cue, and control.
-
-The PCA in ? %???kumar???
-revealed
-neurons that are sensitive to both press and release.
-For this reason, the press-sensitive population
-was increased in dimensionality
-such that the population was sensitive
-to both press and release,
-in the same manner to the multidimensional
-methods used in ?. %???singh eliasmith???.
-
-The auditory cue ?
-%??? todo : see if it's necessary to be an integrator ???
-
-Control is the signal
-that causes release-sensitive neurons to fire,
-resulting in the release of the lever.
-Determining how to implement this control
-is one of the primary goals
-of mPFC research in tasks with delays.
-Its implementation is in itself
-a hypothesis of how the mPFC and motor cortex interact
-in order to perform these tasks.
-
-\subsubsection{Implementing release control}
-
-? %???kumar???
-use their PCA to propose a conceptual model,
-shown in Fig ?. %???include???.
-We extend this model
-by associating prepotent response and proactive inhibition
-with neural populations and
-showing how they can be connected
-such that the model can respond
-within the 600 ms response window.
-
-In the double integrator model,
-a population of neurons tracks the current behavioural state.
-In the integrated model, two populations
-track a similar but subtly different quantity.
-One population is assumed to exist
-in motor cortex, and tracks whether or not
-the subject is currently holding down the lever.
-This is an integrative population,
-which is driven by press-sensitive neurons.
-The other population is assumed to exist
-in medial PFC, and tracks whether or not
-the subject should be delaying the release response.
-It is primarily driven by the motor cortex's
-holding population.
-
-The motor cortex's holding population
-sends excitatory projections
-to the release-sensitive neurons,
-effecting the prepotent response from the conceptual model.
-The medial PFC's waiting population
-send inhibitory projects
-to the release-sensitive neurons,
-effecting the proactive inhibition from the conceptual model.
-It should be noted that this organization
-produces a slight delay between these two quantities,
-with prepotent response occurring first,
-quickly followed by proactive inhibition.
-This is consistent with ?. %???figure from JNP2009???.
-
-The medial PFC's waiting population's activity
-is further modulated by two task-related populations.
-One is a ramping population that tracks
-the time since the press was initiated.
-The other is a cue-sensitive population
-that is driven by the onset of the auditory cue.
-These two populations
-release the proactive inhibition,
-causing the release neurons to be disinhibited,
-effecting the release of the lever.
-
-As will be discussed further in ?, %???sec???,
-the similarity between the two populations
-may explain why the first two principal components
-of motor cortex and mPFC are very similar.
-
-\subsubsection{Model manipulations}
-
-The model is manipulated in three ways
-in this study.
-
-The model emulates the effects of aging
-by decreasing the number of neurons
-in each neural population,
-by uniformly decreasing the strength
-of each synapse in the network,
-and by randomly removing a fraction
-of all synapses in the network.
-
-Lesion studies are emulated
-by targeted ablations of either
-neurons or synapses in or between specific
-neural populations.
-
-Cooling is emulated by
-lowering the distribution
-of neural gains associated with the
-leaky integrate-and-fire neurons
-in the neural populations being cooled.
-Time constants affecting transmission speed
-between populations are raised
-to emulate slowed transmission speed
-as a result of the cooling of white matter.
+In addition to the description in the main text,
+it should be noted that \citeauthor{Singh2006}
+discuss four possible
+realizations of their double integrator model,
+all of which produce the same dynamics
+in the final output population.
+The realizations differ in whether
+the signals are internally or externally generated,
+and in whether a single multidimensional integrator
+or multiple one-dimensional integrators are used.
+
+For ease of implementation,
+we used multiple one-dimensional integrators
+because the resulting network is more stable than
+a network with a single multidimensional integrator
+\citep{Singh2006}.
+No microstimulation was done
+during the behavioral experiments
+so it is difficult to determine
+if the experimental ACC
+can also be thought of as
+a system of coupled integrators;
+preliminary simulations using
+a single multidimensional integrator
+produced PCs that matched no better
+than the coupled integrator model,
+and often worse.
+
+The principal component analysis
+shows that activity is modulated
+at the time of the lever press
+and when the outcome is determined;
+it is likely that these signals
+arrive from the motor cortex
+and other cortical areas,
+but we cannot rule out the possibility
+that these signals are internally generated
+from the presented data alone.
 
@@
 
-Both empirical and theoretical experiments
-produce data in the form of spike trains
+Both experimental and theoretical experiments
+produced data in the form of spike trains
 and event timestamps.
-The files produced by NeuroExplorer and Nengo
-were read by a customized version of Neo,
-a Python package for reading neurophysiology data files.
+These data files were read
+by a customized version of Python-Neo.\footnote{Open source software
+  available at \url{http://neuralensemble.org/neo/}.}
 The subsequent analysis was done
-with custom written scripts in Python
-using the NumPy and SciPy libraries for numerical analysis
-and Matplotlib for plotting.
-All analysis scripts are freely available
-at ?. % github URL
-
-\subsubsection{Behaviour}
-
-Figure 4 summarizes a single trial
-of the reaction time task.
-In both the empirical and theoretical studies,
-the following relevant event times are recorded:
-lever press, lever release, reward delivery,
-trigger stimulus onset, and house light extinguishing.
-Sequences of event times can be used to
-identify the result of each trial,
-as shown in Figure ?.
-Filtering the set of event timestamps
-for these sequences identifies each trial outcome.
-
-Reaction time distributions are also analyzed
-for correct trials.
-Reaction time is defined as the difference between
-trigger stimulus onset and lever release.
-
-\subsubsection{Neural}
+with custom Python scripts.
+All analysis scripts are available
+at \url{https://github.com/tbekolay}.
+
+\subsubsection{Neural analysis}
 
@@
 in spike trains around certain events.
-The previously mentioned techniques
-were used to isolate an event of interest.
-For each neuron, spike trains from 1 second prior to
-2 seconds after each instance of an event are isolated.
+For each neuron with an average firing rate over 1 Hz,
+we isolate the spike trains from 4 seconds prior to and
+4 seconds after each press event on a correct trial.
 Each perievent spike train is binned in 1 ms bins and is then
@@
 These spike density functions
-are normalized to z-scores,
+are normalized to Z-scores,
 and then averaged over all trials to
-produce a matrix in which each row
+produce a matrix in which each row is
 the normalized average response of a neuron
@@
 The resulting principal components
-were also normalized to z-scores.
-The amount of variance accounted for by each principal component
+were also normalized to Z-scores.
+The amount of variance accounted for
+by each principal component
 is computed as the square of the component's eigenvalue
 over the sum of all squared eigenvalues
-(i.e., $s_i^2 / \sum s^2$).
-
-In all analyses, neurons not significantly modulated
-during the perievent epoch were omitted.
-Specifically, we omitted neurons with average firing rates
-under 1 Hz and whose firing rates were non-stationary
-across trials
-(determined by a Wald-Wolfowitz runs test with $|Z| > 4$).
+($s_i^2 / \sum s^2$).
+
+\subsubsection{Behavioral analysis}
+
+In both the experimental and theoretical studies,
+the following relevant event times are recorded:
+lever press, lever release, reward delivery,
+trigger stimulus onset, and house light extinguishing.
+Sequences of event times were used to
+identify the result of each trial.
+as shown in Figure~\ref{fig1}.
+Filtering the set of event timestamps
+for these sequences identifies each trial outcome.
+Reaction times are calculated
+only for correct trials,
+and are defined as the difference between
+trigger stimulus onset and lever release.
 
@@
 
-\subsection{Model behaviour is less variable}
-
-A stereotypical trial for both the model
-and experimental subjects is shown
-in Figure 5.
-The experimental session lasted ? seconds,
-over which ? trials were performed.
-? trials were identified as correct trials;
-? were premature responses,
-and ? were delayed responses.
-The model session lasted 500 seconds,
-over which ? trials were performed.
-? trials were identified as correct trials,
-? were premature responses
-and ? were delayed responses.
-
-Over all trials for all subjects,
-experimental subjects performed ? trials
-with a ? \% success rate.
-The model performed ? trials with a
-? \% success rate.
-
-The distribution of reactions times, shown in figure 5,
-is Gaussian-like with a long tail greater than the mean
-for both model and data.
-The mean reaction time for all experimental subjects
-was ? $\pm$ ?.
-The mean reaction time for all model subjects
-was ? $\pm$ ?.
-
-\subsection{Model principal components are similar}
-
-Principal component analysis on the spike trains
-produced by the model show the same number
-of primary principal components, with similar shapes.
-
-In the neural populations hypothesized
-to reside in motor cortex, as shown in Figure 6,
-three leading components are identified.
-As in the experimental data,
-PC1 is rises around the time of the press
-and maintains a high value until the release,
-at which point it drops.
-PC2, which is very similar
-to the cumulative sum of PC1,
-is modulated at the time of lever press,
-and slowly changes during the delay period.
-PC3 is strongly modulated at both the
-time of lever press and release.
-Firing rates in the model match those in
-the data, increasing before the press
-and after release.
-
-In the neural populations hypothesized
-to reside in medial prefrontal cortex,
-as shown in figure 7,
-two leading components are identified.
-The first component is strongly modulated
-around the time of lever press,
-and slowly decreases during the delay period.
-The second component,
-which is similar to the cumulative sum of
-PC1, increases at the time of lever press,
-and remains high during the delay period.
-
-\subsection{Model manipulations}
-
-Aging results, figure 8.
-
-Lesion results, figure 9.
-
-Cooling results, figure 10.
-
-\section{Discussion}
-
-\subsection{Model behavioural variability is low}
-
-The primary difference is in the amount
-of inter-trial and intra-trial variability.
-Experimental subjects had several blocks of time
-in which they performed as many trials as the model,
-but those blocks of time were interrupted by
-error trials and time in which the animal
-was performing other actions such as grooming.
-Within the trial, reaction time distribution
-for the model and experimental subjects are
-similar but with far less variability.
-The model's reactions times
-are within the experimental subjects'
-mean and standard deviation, however,
-suggesting that the model
-could have the same variability
-by simply injecting noise into the system.
-This would also result in
-the error trials that are missing in the model.
-
-A better basis for comparison for the model
-would be human performance in a similar
-reaction time task. ?find some % Look up some behavioural data
-
-\subsection{Model principal components match the data}
-
-While it may be unsurprising
-due to the model being inspired by the original
-principal component analysis,
-it is strong validation for the original conceptual model
-that the implemented model maintain
-the same principal components
-while still being able to simulate behaviours.
-
-? List differences and investigate (with final version of model)
-
-\subsection{Model predictions}
-
-The model manipulations assert that blah has the effect of blah.
-
-Prior studies (some unpublished) have suggested that these are indeed the effects of those things.
-
-Cooling has not been studied in an experimental setting,
-and so we take the results as a prediction of the model.
+\subsection{Adaptive control causes premature releases}
+
+Figure~\ref{fig11} shows the decoded values
+of neural populations during a trial in which
+the double integrator model
+causes the control model to respond faster than
+the cue-responding strategy.
+Figure~\ref{fig12} shows the decoded values
+during a trial in which
+the double integrator causes a premature response.
+Late responses, as seen in Figure~\ref{fig12},
+are identical with and without the double integrator model.
 
@@
 
-\appendix
-
 \section{Figures}
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f1-pcasummary}
-    \caption{Summary of principal components identified
-      by Narayanan and Laubach, 2009. Three components
-      were identified in motor cortex (top row), accounting
-      for over 60\% of variation.
-      Two components were identified in prefrontal cortex
-      (bottom row), accounting for over 25\% of variation.}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f2-conceptual}
-    \caption{The conceptual model proposed by Narayana and
-      Laubach, 2009. In the conceptual model, the start of the
-      trial activates two systems; one prepares the response
-      action (prepotent response) and the other inhibits
-      that action until the appropriate time (proactive inhibition).}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f3-fullmodel}
-    \caption{The proposed implementation of the conceptual model.
-      The motor cortex is made up of two populations of
-      leaky integrate-and-fire neurons, one of which interacts
-      with the environment by pressing or releasing a lever.
-      The medial prefrontal cortex is made up of three
-      populations of LIF neurons. See text (section 2) for details.}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f4-behaviour}
-    \caption{Possible outcomes for each trial. If the subject
-      responds within the 0.6 second response window, the trial
-      is counted as correct and reward is delivered. If the
-      subject responds prematurely, it is counted as an error
-      and house lights are extinguished for a timeout period.
-      If the subject responds after the response window,
-      an error is also recorded and the lights are extinguished.}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f5-behaviour}
-    \caption{Summary of model and experimental behaviour.
-      (Left) Seconds 200 - 500 of an experimental (top)
-      and model (bottom) session.
-      (Right) Histograms of reactions times for
-      experimental and model subjects.}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f6-mcpca}
-    \caption{Principal components identified in motor cortex
-      spike data. (Top) Experiment. (Bottom) Model.}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{f7-pfcpca}
-    \caption{Principal components identified in medial prefrontal
-      cortex spike data. (Top) Experiment. (Bottom) Model.}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=\columnwidth]{f7-pfcpca}
-    Placeholder
-    \caption{Aging results}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=\columnwidth]{f7-pfcpca}
-    Placeholder
-    \caption{Lesion results}
-\end{figure}
-
-\clearpage
-
-\begin{figure}
-  \centering
-    % \includegraphics[width=\columnwidth]{f7-pfcpca}
-    Placeholder
-    \caption{Cooling results}
-\end{figure}
+\scalefigone{fig1}{0.5}{
+  The simple reaction time task used in the experiment.
+  Trials are classified as correct, premature, or late
+  depending on the time of lever release.
+}
+
+\scalefigone{fig2}{1.0}{
+  Experimental motivation for the adaptive control model. Data from JNP 2009.
+  (A) The simple reaction time task.
+  (B) When lever release events occurred.
+  Each trial is aligned such that the press event
+  occurs at time $t=0$.
+  All of the trials shown are correct trials.
+  Black dots represent correct trials
+  that were preceded by correct trials.
+  Red dots represent correct trials
+  that were preceded by error trails
+  (either premature or late).
+  (C) Peri-event spike rasters and histograms
+  showing different neural activity
+  based on the previous outcome.
+  In all six plots, black represents activity
+  when the last trial was correct, and red
+  represents activity when the last trial was an error.
+  The shaded grey areas note where
+  the neural activity is significantly different
+  between the two conditions (post-correct and post-error).
+  In the top row, there is elevated neural activity
+  before the press is initiated, following an error.
+  In the middle row, there is elevated neural activity
+  before the press is initiated (both)
+  and during the foreperiod (right)
+  following correct trials.
+  In the bottom row, there is elevated neural activity
+  during the whole trial (left)
+  and after the trial (right)
+  following error trials.
+  (D) Linear regression with a GLM was performed
+  on the neural activity.
+  Before the press is initiated,
+  18.7\% of neurons encoded the previous outcome,
+  compared to less than 10\% that
+  encoded the current reaction time
+  and the interaction between the two.
+  (E) Z-statistics showing which neurons were significantly
+  modulated by the previous outcome, and when.
+  (Top) Sorted by the absolute value of the Z-statistic.
+  The bottom 18.7\% of the rows have high average Z-scores
+  ($p < 0.05$).
+  (Bottom) Sorted by the peak of the Z-statistic.
+  The line is close to linear, suggesting that the
+  previous outcome information is being encoded over time
+  by the entire population.
+}
+
+\scalefigone{fig3}{1.0}{
+  A summary of the principal component analysis done
+  in JNP 2009;
+  the motivation for investigating Singh \& Eliasmith 2006.
+  (A) A graphical summary of principal component analysis.
+  Nomarlized perievent neural data,
+  in the form of the Z-scores of instantaneous firing rates,
+  is organized in a matrix, with each row
+  being the z-scored firing rates
+  of a single neuron on a single trial.
+  Singular value decomposition is performed on the matrix,
+  resulting in a matrix such that
+  the number of columns (time bins) is the same.
+  The rows are now ordered such that the first row
+  contains largest eigenvector,
+  which represents the value when the original
+  data is projected onto the axes of highest variance.
+  (B) The fraction of variance explained
+  by each singular value
+  when PCA is performed on the post-correct neural data
+  (top), and the post-error data (bottom).
+  In both cases, the two eigenvectors with the highest
+  singular values account for nearly 50\% of variance.
+  The next two eigenvectors also account for an amount
+  of variance higher than would be expected when
+  we linearly interpolate from the smaller
+  singular values.
+  (C) A summary of the loadings of the two top
+  eigenvector on each neuron.
+  If a neuron were encoding both the first and second
+  eigenvectors to the same degree,
+  we would expect a horizontal line.
+  The varied lines criss-crossing each other
+  suggests that each individual neuron
+  has different sensitivities
+  to the first and second principal components.
+  (D) The top three principal components
+  for the post-correct trials (black)
+  and the post-error trials (red).
+  (E) (Top) Plotting the cumulative sum (integral)
+  of the first principal component closely matches
+  the second principal component, and vice-versa
+  ($R^2 = 0.904$ and $R^2 = 0.939$ respectively),
+  in the post-correct case.
+  (Bottom) In the post-error case,
+  the first two principal components are no longer
+  cumulative sums of each other
+  ($R^2 = 0.639$ and $R^2 = 0.676$).
+  This points to neural integration
+  as a potential mechanism explaining ACC activity.
+  (F) Normalized spike-density functions
+  for all of the neurons analyzed in the post-correct case,
+  organized by the loading
+  (left) on the first principal component, and
+  (right) on the second principal component.
+}
+
+\scalefigone{fig4}{0.8}{
+  Block diagram of the adaptive control model.
+  (Left) The adaptive control model contains
+  the double integrator model,
+  from Singh \& Eliasmith.
+  It has been modified by accepting
+  both press information ($u$)
+  and error information ($E$)
+  to drive $x_1$, and
+  reward information ($R$)
+  to control the integration
+  of both $x_1$ and $x_2$.
+  The output of the double integrator,
+  colored red, is the only component
+  that interacts with the cue-responding
+  circuit, by activating neurons
+  in the \textit{Trigger} population.
+  (Middle) The cue-responding circuit
+  that is able to perform
+  the simple reaction time task
+  by responding to the Cue
+  provided by the environment
+  (see text for details).
+  (Right) Values that are not modeled
+  with spiking neural populations,
+  and are therefore considered
+  signals that are provided by the environment.
+  The lever accept the decoded
+  output of the \textit{Press/Release} population,
+  lowering after \textit{Press} sensitive
+  neurons are activated (following some delay),
+  and raising after \textit{Release} sensitive
+  neurons are activated (following the same delay).
+}
+
+\scalefigone{fig5}{1.0}{
+  A detailed breakdown of the dynamics and
+  trajectory the model takes through state space
+  for (A) correct, (B) premature, and (C) late trials
+  following a correct trial.
+  The injected oscillation is omitted
+  in these plots;
+  its time dependence makes it misleading
+  to draw vector fields for a particular time.
+  (A) In a correct trial, the pressing of the lever
+  sends the double integrator model into
+  a high part of the state space.
+  The connection between the two integrators
+  causes the passive dynamics (panel 2)
+  to have a shift towards the right.
+  The cue is delivered while drifting
+  to the right, and the release occurs
+  near the top-right portion of the state space.
+  After the successful release,
+  the reward turns off both integrators,
+  causing the system to return to the origin point,
+  where it remains until the next trial.
+  (B) In a premature trial, the lever is released
+  prior to the onset of the cue (panel 2).
+  Premature release causes the subject
+  to detect an error, which drives
+  the ACC to a low point in the state space.
+  Between trials (panel 4),
+  the passive dynamics of the state space
+  drive the ACC to the bottom-left portion
+  of the space, where it will be
+  at the start of the next trial.
+  (C) In a late trial, the cue occurs,
+  but the lever is not released (panel 2).
+  Late release causes the subject
+  to detect an error, which drives the ACC
+  to a low point in the state space.
+  Between trials, the passive dynamics
+  of the state space drive the ACC
+  to the same bottom-left portion of the space
+  that the premature error results in.
+  Therefore, late and premature errors
+  cannot be distinguished on the next trial
+  in this model.
+}
+
+\scalefigone{fig6}{1.0}{
+  The effects of damping on the mathematical
+  model for correct (left), premature (middle),
+  and late (right) trials.
+  Note the different end points (marked with arrows).
+  Four damping values were simulated;
+  $D = [0, 0.025, 0.05, 0.075]$.
+  As can be seen in the middle and right panels,
+  with $D \ge 0.05$,
+  correct and error trials can no longer
+  be differentiated on the next trial
+  because the ACC is in the same state
+  regardless of the outcome of the previous trial.
+}
+
+\scalefigone{fig7}{1.0}{
+  Decoded values of the double integrator model.
+  They take the same trajectory through state
+  space as the mathematical model (see Figure 5).
+  Similar to the mathematical model,
+  the ACC represents a very different state
+  after correct (left) and error (middle, right)
+  trials.
+}
+
+\scalefigone{fig8}{1.0}{
+  The effect of aging through degrading recurrent connection weights
+  on the spiking neural model for correct (left), premature (center),
+  and late (right) trials.
+  Note the different end points (marked with arrows).
+  Weights were degraded by first solving for an appropriate
+  weight matrix, and then multiplying them by four values:
+  $[1.0, 0.99, 0.98, 0.97]$.
+  As can be seen in the middle and right panels,
+  by multiplying weights by $\le 0.98$,
+  correct and error trials can no longer
+  be differentiated on the next trial
+  because the ACC is in the same state
+  regardless of the outcome of the previous trial.
+}
+
+\scalefigone{fig9}{1.0}{
+  Results of principal component analysis
+  on the data generated by the double integrator model
+  for correct presses (A) post-correct trails,
+  (B) post-premature trials, and (C) post-late trials.
+  The top panels compare the top principal components
+  of the model to the experimental ACC.
+  In all cases, the PCs are significantly similar
+  ($R^2 > 0.86$).
+  The middle panels show normalized
+  peri-event spike density functions
+  for 174 randomly sampled neurons from the
+  simulated double integrator model.
+  The bottom panels show normalized
+  peri-event spike density functions
+  for all 174 neurons recorded
+  during the experimental study.
+}
+
+\scalefigone{fig10}{1.0}{
+  Behavioural performance of the
+  experiment, cue-responding, adaptive control,
+  and aged adaptive control models.
+  The top panels summarize the number of
+  correct, premature, and late trials
+  for each experimental or simulated subject.
+  The number of total trials
+  for the simulations was approximately
+  matched to the experimental subjects.
+  The bottom panels summarize the reaction times
+  on correct trials for each subject.
+  The dotted black line represents
+  the mean of all subjects' median reaction times.
+}
+
+\scalefigone{fig11}{1.0}{
+  Decoded values from a representative correct trial
+  for the simple network (A) and the adaptive network (B).
+  (A) In the simple network, the trigger neurons
+  activate only once the cue is delivered.
+  Trigger neurons activate the release neurons,
+  which initiate the release of the lever.
+  There is a delay ($\sim 300ms$)
+  between the activation of release neurons
+  and the actual release of the lever.
+  (B) In the adaptive network, when
+  in the appropriate state, the cue
+  can be predicted by the double integrator model,
+  causing the activation of trigger---and
+  therefore release---neurons before
+  the time of the cue. This results
+  in a faster reaction to the cue.
+}
+
+\scalefigone{fig12}{1.0}{
+  Decoded values from a representative late trial for the simple
+  network (A), and a premature trial for
+  the adaptive network (B).
+  (A) Late release occurs when
+  the release neurons are not sufficiently
+  active in order to effect the lever release.
+  Alternatively, late release can occur
+  if the release signal is not faithfully
+  transmitted from the release neurons
+  (modeling activity in motor cortex)
+  to the intermediate populations
+  (modeling activity between motor cortex and muscles)
+  before interfacing with the simulated environment.
+  A late release in the adaptive model
+  would look identical, with little activity
+  in the double integrator model that predicts cues.
+  (B) Premature release occurs when
+  the double integrator model incorrectly predicts
+  the time of the cue.
+  This occurs when integration is too fast,
+  and the trigger neurons are activated
+  such that a lever release is effected
+  before the time of the cue.
+  Note that the non-adaptive cue-responding
+  circuit cannot make premature releases,
+  because releases are critically dependent
+  on cue information.
+}
 

### 2013-06-10 paper/coverletter.tex
Fourth version of the paper, with figures.
- This is the version submitted to JNeuroscience

% LaTeX resume using res.cls
\documentclass{letter}
\usepackage[T1]{fontenc}
\usepackage{charter}
\usepackage{inconsolata}
\usepackage{graphicx}

\begin{document}

Dear Editor,

Please accept for consideration our paper entitled
``A spiking neural integrator model of the adaptive
control of action by the anterior cingulate cortex.''

This paper presents a mathematical hypothesis
to explain the role of the anterior cingulate cortex (ACC)
in the control of action.
This hypothesis is tested by implementing
it in a spiking neural network model,
and it is validated by showing it qualitatively
and quantitatively matches both neural and
behavioral data recorded from rodents.
We believe that the proposed model is the most
detailed account of ACC function
to date. We suggest that our framework can
be extended to explain a variety of
ACC functions we have not directly explored.

This study is the result of a close collaboration
between the Centre for Theoretical Neuroscience
at the University of Waterloo
and the John B. Pierce laboratory in New Haven.
We believe that this combination of
experimental and theoretical expertise
is one of the strengths of this paper.

Sincerely, \\
Trevor Bekolay, Mark Laubach, Chris Eliasmith

\end{document}

### 2013-06-10 paper/jneurosci2013.tex
Fourth version of the paper, with figures.
- This is the version submitted to JNeuroscience

@@
 \documentclass[11pt]{article}
-% Change "article" to "report" to get rid of page number on title page
 \usepackage{amsmath,amsfonts,amsthm,amssymb}
@@
 \usepackage{chngpage}
+\usepackage{caption}
 \usepackage{soul}
-\usepackage[round]{natbib}
+\usepackage{namedplus}
 \usepackage[usenames,dvipsnames]{color}
@@
 
-\pagestyle{empty}
+\pagestyle{plain}
 
@@
 \setlength\fboxrule{1pt}
+\linespread{1.5}
 
@@
 % Make title
-\title{Adaptive control of action timing with a spiking neural integrator
-  model of anterior cingulate cortex}
-\date{}
-\author{Trevor Bekolay, Chris Eliasmith, Mark Laubach}
+\title{A spiking neural integrator model of the adaptive control of
+  action by the anterior cingulate cortex}
+\date{Abbreviated title: A spiking model of ACC adaptive control}
+\author{%
+  Trevor Bekolay \\
+  Centre for Theoretical Neuroscience \\
+  University of Waterloo \\ ~ \\
+  Mark Laubach \\
+  The John B. Pierce Lab \\
+  New Haven, CT \\ ~ \\
+  Chris Eliasmith \\
+  Centre for Theoretical Neuroscience \\
+  University of Waterloo}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\textbf{Corresponding author:} Chris Eliasmith.
+Email: \url{celiasmith@uwaterloo.ca}.
+Centre for Theoretical Neuroscience, University of Waterloo.
+200 University Avenue West, Waterloo, ON, N2L 3G1.
+
+\textbf{Pages:} 24. \textbf{Figures:} 8. \textbf{Tables, multimedia, 3D models:} 0
+
+\textbf{Word counts:} Abstract: 164, Introduction: 445, Discussion: 479
+
+\textbf{Conflict of Interest}
+The authors declare no competing financial interests.
+
+\textbf{Acknowledgements}
+
+Financial support: NSF 1121147 to ML;
+NSERC Discovery, CRC, CFI, and OIT to CE;
+and NSERC CGS-D to TB.
+We thank Benjamine Liu for technical contributions
+to the modeling effort and
+Nandakumar Narayanan who recorded the neurons
+that were analyzed by TB and ML.
+
+\clearpage
+
 \begin{abstract}
-  Animals can learn to perform a simple reaction time task
-  with a predictable foreperiod
-  by predicting the time of the cue,
-  achieving reaction times faster
-  than through cue-responding alone.
-  The outcome of the previous trial
-  is one factor in determining
+  Reaction times can be improved by preparing movements
+  in advance and by learning the expected timing
+  of action-imperative stimuli.
+  Success or failure on the previous trial
+  can be an important factor for determining
   whether an animal will attempt to time
-  the foreperiod.
+  the stimulus or simply wait for it to occur
+  before initiating action.
   The anterior cingulate cortex (ACC) has been implicated
-  in enabling control of action timing
+  in enabling the top-down control of action
   depending on the outcome of the previous trial.
@@
   for adaptive control in precisely timed tasks.
-  We show that a spiking neural circuit
+  We show through simulation
+  that a spiking neural circuit
   consisting of coupled neural integrators
@@
   of the experimentally recorded ACC.
-  We use this coupled integrator model
-  to adaptively control
-  a spiking neural circuit that follows
-  a cue-response strategy,
-  and show that it performs the simple reaction time task
+  We expand on this coupled integrator model
+  to construct a spiking neural circuit that
+  performs a reaction time task by following
+  either a timing or cue-response strategy,
+  and show that it performs a simple reaction time task
   with similar reaction times as experimental subjects.
-  We then alter the model
-  in a way consistent with
-  how the aged brain degrades,
-  and predict that aged subjects
-  will adopt a cue-response strategy.
 \end{abstract}
 
+\linespread{1.5}
+
 \section{Introduction}
 
-The anterior cingulate cortex has been implicated
+The anterior cingulate cortex (ACC) has been implicated
 in a wide variety of functions,
 including conflict monitoring
-\citep{Aarts2009, Botvinick2004, VanVeen2004},
-error monitoring \citep{Brown2005, Carter1998, Fecteau2003,
-  Modirrousta2008, Rushworth2004},
-anticipatory control \citep{Aarts2008, Koyama2001},
+\cite{Sheth2012,Aarts2009,Botvinick2004,VanVeen2004},
+performance monitoring
+\cite{Brown2005,Carter1998,Fecteau2003,Modirrousta2008,Rushworth2004,Histed2009,Hyman2013,Horst2012},
+anticipatory control \cite{Aarts2008,Koyama2001},
 arousal of the sympathetic nervous system
-\citep{Critchley2003, Luu2003},
-and control of action timing \citep{Muir1996, Mulert2003,
-  Naito2013, Narayanan2009, Risterucci2003}.
-Many have proposed integrative or general models
+\cite{Critchley2003,Luu2003},
+and control of action timing
+\cite{Muir1996,Mulert2003,Naito2013,Narayanan2009,Risterucci2003}.
+Conceptual models have been proposed
 that would allow the ACC to be involved
 in several of these functions,
-depending on the behavioral context
-\citep{Botvinick2004, Luu2003}.
+depending on behavioral context
+(e.g., \citenoparens{Botvinick2004,Luu2003}).
 However, while there have been
-attempts at modeling the ACC at the level
-of artificial neural networks
-\citep{Botvinick2004, Brown2005},
+attempts at modeling the ACC
+using artificial neural networks
+\cite{Botvinick2004,Brown2005},
 there are currently no mathematical
@@
 modulated by the ACC.
-Similarly, no existing models
+As well, no existing models
 have been implemented with spiking neurons,
-and are therefore not directly comparable
-to neural activity.
+and therefore no models can be analyzed
+with the same methods used
+to analyze single neuron recordings.
 
@@
 two-dimensional dynamical system
-can capture the activity of the ACC.
+can account for ACC activity
+related to error monitoring and motor action timing.
 We test the model in a simple reaction time task
@@
 on the outcome of the last trial.
-We implement the dynamical system
-in a spiking neural network
-based on \citet{Singh2006}
-and show that simulated neural activity
-closely matches recorded ACC activity.
-We show that the model
-can be combined with
-a motor control model,
-resulting in behavioral changes
-analogous to previous behavioral theories
-of ACC function.
-Finally, we damage the model in a way analogous
-to aging, and predict that
-aged subjects will adopt a more conservative
-strategy due to a reduction of the
-ability of the ACC to control action timing.
-
-Our mathematical account of ACC function
-is applied directly to
-error monitoring and
+The dynamical system is based
+on a multi-dimensional attractor
+network proposed by \citetext{Singh2006},
+and is implemented in a spiking neural network
+using the Neural Engineering Framework
+(NEF; \citenoparens{Eliasmith2003}).
+We show, using identical analyses
+to those employed on the experimental data,
+that simulated neural activity from this circuit
+closely matches recorded ACC activity
+and that the model can be combined with
+a simple motor control model
+to produce behavior consistent
+with behavioral studies of simple reaction time tasks.
+
+Our model also provides an alternative hypothesis
+of how the ACC accomplishes the control of action timing.
+\citetext{Narayanan2009} hypothesized
+that ACC activity inhibits motor responses
+until the time of the cue.
+Our model instead proposes that
+ACC activity encodes the duration of the action
+and that inhibition is carried out elsewhere.
+
+In this paper we apply our
+mathematical account of ACC function
+to explain error monitoring and
 the control of action timing.
-However, we assert that the dynamical systems
-approach is general and can be used
+However, we suggest that the dynamical modelling
+method we employ is general and can be extended
 to explain other theories
@@
 
-\section{Simple reaction time task} \label{sec:simple-rt}
-
-The simple reaction time task
-serves as a minimal task in which
-to examine the control of precisely timed actions.
-In this task, the subject presses and holds down a lever
+\section{Methods}
+
+\subsection{The reaction time task} \label{sec:simple-rt}
+
+Twelve male Long-Evans rats
+were trained to press down on a lever
 for a fixed amount of time (the ``foreperiod'';
-1 second for this study).
+1 second for this study)
+using standard methods \cite{Laubach2000}.
 After the foreperiod, a cue is presented
-(e.g., an auditory tone).
+(an auditory tone).
 The subject then has 0.6 seconds (the response window)
@@
 If the lever is released in that time,
-the trial is a success, and the subject is rewarded.
+the trial is a success, and the subject is rewarded
+with water.
 If the lever is released during the foreperiod,
 the trial is classified as a ``premature error,''
-and is penalized by a time-out.
+and is penalized by a time-out with the house lights extinguished.
 If the lever is not released within the response window,
 the trail is classified as a ``late error,''
-and is also penalized by a time-out.
+and is also penalized by a time-out with the house lights extinguished.
 Figure~\ref{fig1} presents a schematic view of the task.
-
-\section{Analysis of experimental data}
-
-\citet{Narayanan2009} recorded
-the anterior cingulate cortex (ACC) and motor cortex
-of rodents performing this simple reaction-time task.
-They used principal component analysis
-on normalized spike densities
-in order to determine
-what is being represented in these brain areas
-during this task.
+Subjects were considered to have learned the task
+when they performed in correctly $> 60 \%$
+of the time.
+
+The same reaction time task was simulated
+in the Nengo simulation environment
+using a finite state machine to track
+task state, and scalar values between -1 and 1
+to indicate the state of house lights,
+lever position, auditory cue, and reward.
+The Nengo scripts simulating this environment
+are available at
+\url{https://github.com/tbekolay/jneurosci2013}.
+
+\subsection{Data analysis}
+
+\subsubsection{Neural data}
+
+The experimental data analyzed in this study is the same
+as that in \citetext{Narayanan2009}, in which
+the authors recorded the anterior cingulate cortex (ACC) and motor cortex
+of rodents performing this simple reaction time task.
+(The ACC recordings were made in the prelimbic cortex,
+which is a component of the ACC.
+The area was referred to as the dorsomedial PFC
+in the original paper.)
+\citeauthor{Narayanan2009} used principal component analysis
+(PCA) to characterize common firing patterns
+in the ACC during the task.
+
+In the PCA, only neurons with an average
+firing rate over 1 Hz are considered.
+Spike trains from 4 seconds prior to and 4 seconds after
+each press event on a correct trial are isolated.
+Each perievent spike train is binned in 1 ms bins and is then
+convolved with a Gaussian filter ($\sigma = $ 25 ms;
+PCs are consistent with many other $\sigma$ values).
+These spike density functions
+are normalized to Z-scores,
+and then averaged over all trials to
+produce a matrix in which each row is
+the normalized average response of a neuron
+over the perievent epoch.
+Singular value decomposition
+is performed on that matrix
+to isolate the principal components,
+which are then normalized to Z-scores.
+The amount of variance accounted for
+by each principal component
+is computed as the square of the component's eigenvalue
+over the sum of all squared eigenvalues
+($s_i^2 / \sum s^2$).
+
 As shown in Figure~\ref{fig2},
 the first two principal components
-found from analyzing the ACC data are large,
-and are hypothesized to track
-1) if the subject was performing the task, and
-2) the relative amount of time the subject has been
-performing the task.
+of the experimental data
+are modulated around the task events
+and appear to track
+whether the subject has pressed the lever and
+the relative amount of time
+that the lever has been pressed.
 \citeauthor{Narayanan2009} also found that the neural activity
@@
 This result suggests that
-the ACC is performing some kind of integration..
-
-\section{Double integrator model}
-
-One model of neural integration
+the ACC is performing some kind of integration,
+and that integration depends
+on prior behavioral outcomes.
+
+\subsubsection{Behavioral data}
+
+In both the experimental and theoretical studies,
+the following relevant event times are recorded:
+lever press, lever release, reward delivery,
+trigger stimulus onset, and house light extinguishing.
+Sequences of event times were used to
+identify the result of each trial.
+Filtering the set of event timestamps
+for these sequences identifies each trial outcome.
+Reaction times are calculated
+only for correct trials,
+and are defined as the difference between
+trigger stimulus onset and lever release.
+
+\subsection{Double integrator model}
+
+The double integrator model
 was proposed in the context of working memory
-by \citet{Singh2006}.
-Their model, called the \textit{double integrator} model,
-uses multidimensional neurons
+by \citetext{Singh2006}.
+Their model uses multidimensional neurons
 in integrative populations
@@
 during a vibrotactile discrimination task.
-It relies on multiple coupled integrators,
+It relies on two coupled integrators,
 whose activity resembles
@@
 The model employs
-the Neural Engineering Framework (NEF; \citealp{Eliasmith2003})
+the Neural Engineering Framework (NEF; \citenoparens{Eliasmith2003})
 to implement a spiking neural network
 that captures the dynamics of the
-integrative dynamical system
-(see section \ref{sec:nef}).
-We hypothesize that the \citeauthor{Singh2006} model
+integrative dynamical system.
+We hypothesize that a model
+similar to that proposed by \citeauthor{Singh2006}
 will exhibit the same neural dynamics
@@
 
-\subsection{Dynamics}
+\subsubsection{Dynamics} \label{sec:dynamics}
 
 The dynamics of the double integrator model
-are captured in the following dynamics equations.
+are captured in the following dynamics equations:
 \begin{align*}
@@
 in which $x_1$ integrates its input,
-maintaining the accumulated value over time.
+maintaining the accumulated value over time, and
 $x_2$ integrates the value of $x_1$.
-\citet{Singh2006} have proposed that this system
-enables time tracking;
-because $x_2$ integrates $x_1$,
+\citetext{Singh2006} have proposed that this system
+enables time tracking.
+Because $x_2$ integrates $x_1$,
 the amount of time elapsed since
@@
 
-In the simple reaction time task,
-the input, $u$, is the activation of neurons
-that effect the lever press.
+In the simple RT task,
+we want to keep track of how much time
+has elapsed since the subject
+has pressed the lever.
+While the lever press event could be
+used as the input to the system,
+since we aim to embed this system
+in a spiking neural model,
+we instead use the activity of the motor neurons
+that effect the lever press as
+the input, $u$.
 The system is also modified by signals
@@
 effectively breaking the integration,
-as is seen in Figure~\ref{fig3} in post-error trails.
-This results in the following dynamics.
+as is seen in Figure~\ref{fig2} in post-error trails.
+This results in the following overall dynamics.
 \begin{align} \label{dint}
@@
 \end{align}
-See Figure~\ref{fig4} for a graphical
+See Figure~\ref{fig3} for a graphical
 depiction of the double integrator model.
 
-In addition to the depicted system,
+The dynamics of Equation \eqref{dint}
+are predictable for certain values
+of $\alpha$, $E$, and $R$.
+$\alpha$ defines how quickly
+the system drifts during the task;
+we set it such that $x_2 \approx 1$
+at the time of release,
+which occurs when $\alpha = 0.04$.
+If $R$ is sufficiently big,
+then when reward is delivered
+the system will be driven to the origin.
+This occurs when $\int_0^{t_R} R\,dt \ge 1$.
+We set the length of reward delivery,
+$t_R$, to 1 second to reflect the pace
+of the experimental subjects,
+and we set the reward magnitude,
+$R$, to 0.15 so that $\int_0^{t_R} R\,dt \ge 1$.
+Larger values of $R$ would be functionally identical,
+as $x_1 = x_2 = 0$ when $R \ge 0.15$ for $t_R$ seconds.
+If $E$ is sufficiently low,
+then when error is detected,
+the system will be driven to
+the lowest $x_1$ value that can
+be represented by the population
+of spiking neurons.
+This exploits the saturating nature
+of spiking neurons in order to
+obtain stability in this
+mathematically unstable system.
+If $\int_0^{t_E} E\,dt \le -1$,
+then the system will saturate
+at a low value for $x_1$.
+We set $t_E = 2 s$ and $E = -0.06$.
+
+To summarize, there are five parameters
+in this dynamical system:
+$t_R$ and $t_E$ are set based on the pace
+of experimental subjects;
+$R$ and $E$ have little behavioral
+consequence if $R \ge 0.15$ and
+$E \le -0.06$ and are set
+to match the experimental principal components;
+$\alpha$ is a free parameter
+that critically affects the model's success
+and is set such that $x_2 \approx 1$
+at the time of release.
+
+In addition to the above system,
 we inject a slowly oscillating signal into $x_1$.
-This is motivated by the observation
-that system \eqref{dint} is stationary
-during the intertrial interval;
-$x_1$ is either near 0 on rewarded trials
-or some low value on error trials.
-$x_2$ does not change on rewarded trials,
-and saturates at some low value on error trials.
-Figure~\ref{fig3} shows some activity during
+Having made the parameter choices
+described above,
+the system is stable during the intertrial
+interval in both post-correct and post-error
+cases.
+However, the experimental data shown in
+Figure~\ref{fig2} exhibits some activity during
 the intertrial interval;
-we model this additional activity
-by a simple oscillation, $O = \sin(2 \pi f t + \phi)$,
+we capture this additional activity
+by introducing a simple oscillation,
+$O = \sin(2 \pi f t + \phi)$,
 where the frequency $f$ is low ($\le 0.4$Hz).
+While there is experimental evidence for
+a slow oscillations like $O$ in various brain regions
+\cite{Buzsaki2004}, this injected oscillation
+does not affect the behavioral performance of the network,
+but instead models ACC activity
+not being explicitly modelled in
+the remainder of the circuit.
+This activity could reflect
+other functions that the ACC is performing,
+or could reflect the overall pace
+of the simple RT task.
 The final dynamics are
@@
 
-Note that while there is experimental evidence for
-a slow oscillations like $O$ in various brain regions
-\citep{Buzsaki2004}, this injected oscillation
-does not affect the performance of the network,
-and is intended to model ACC activity
-not being otherwise modelled in the current study.
-This activity could reflect
-other functions that the ACC is performing,
-or could reflect the overall pace
-of the simple RT task.
-
-\section{Adaptive control circuit}
-
-\citet{Narayanan2009} hypothesized
-that the ACC activity analyzed in the present study
-could be used to inhibit motor responses
+\subsubsection{Neural Engineering Framework} \label{sec:nef}
+
+Equation~\eqref{dint-full} is implemented
+in a spiking neural network using
+the principles of the Neural Engineering Framework
+(NEF; \citenoparens{Eliasmith2003}).
+The NEF enables the simulation of dynamical systems
+like Equation~\eqref{dint-full} by
+making the assumption that populations of spiking neurons
+represent real-valued vectors,
+and using a least-squares optimal method
+to transform those representations
+through the connections between
+populations of neurons.
+A population can have recurrent connections
+in order to produce dynamics,
+such as the neural integrators
+employed in this study.
+
+The NEF's representation scheme
+is similar to population coding,
+as proposed by \citetext{Georgopoulos1986},
+but extended to $n$-dimensional vector spaces.
+Each neuron in a population is sensitive
+to a particular direction,
+called the neuron's \textit{encoder}.
+The activity of a neuron can be expressed as
+\begin{equation}
+  a = G[\alpha \mathbf{e} \cdot \mathbf{x} + J_{bias}],
+\end{equation}
+where $G[\cdot]$ is a nonlinear neural activation function,
+$\alpha$ is a scaling factor (gain) associated with the neuron,
+$\mathbf{e}$ is the neuron's encoder,
+$\mathbf{x}$ is the vector to be encoded, and
+$J_{bias}$ is the background current of the cell
+when $\mathbf{x} = 0$.
+The currently encoded value, $\mathbf{\hat{x}}$,
+can be estimated linearly:
+\begin{equation}
+  \mathbf{\hat{x}}(t) = \sum_i \mathbf{d}_i a_i(t),
+\end{equation}
+where $\mathbf{d}_i$ is the neuron's decoder,
+and $a_i$ is activity of neuron $i$.
+
+Neural activity is interpreted as a
+filtered spike train.
+\begin{equation}
+  a_i(t) = \sum_s h(t - t_s) = \sum_s e^{-(t - t_s) / \tau_{PSC}},
+\end{equation}
+where $h(\cdot)$ is the exponential filter
+applied to each spike,
+and $s$ is the set of all spikes occurring
+before the current time $t$.
+
+Decoders are found through a least-squares minimization
+of the difference between the decoded estimate
+and the actual encoded vector.
+\begin{equation}
+  \mathbf{d} = \Upsilon^{-1} \Gamma \hspace{1.8em}
+  \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
+  \Upsilon_j = \int a_j \mathbf{x} dx.
+\end{equation}
+
+Connection weights to linearly transform these encoded vectors
+can then be computed as
+\begin{equation}
+  \omega_{ij} = \alpha_j \mathbf{e}_j L \mathbf{d}_i,
+\end{equation}
+where $i$ indexes the input population,
+$j$ indexes the output population,
+and $L$ is a linear operator.
+
+% We only do linear transformations I think?
+
+% Connection weights computing a function
+% between two populations connected in a feedforward manner
+% $f(\mathbf{x})$ can be determined by
+% solving for a set of decoding weights
+% for that function,
+% \begin{equation}
+%   \mathbf{d}^{f(\mathbf{x})} = \Upsilon^{-1} \Gamma \hspace{1.8em}
+%   \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
+%   \Upsilon_j = \int a_j f(\mathbf{x}) dx,
+% \end{equation}
+% and then computing the following weight matrix, $\omega$;
+% \begin{equation}
+%   \omega_{ij} = \alpha_j \mathbf{e}_j L \mathbf{d}^{f(\mathbf{x})}_i,
+% \end{equation}
+% where $i$ indexes the input population,
+% $j$ indexes the output population,
+% and $L$ is a linear operator.
+
+Dynamical systems of the form
+$\dot{\mathbf{x}} = A(\mathbf{x}) + B(\mathbf{u})$
+can be created by connecting a population
+to itself with the linear operator
+$L = A$ on the recurrent connection,
+and receiving input $\mathbf{u}$
+from other populations and
+setting $L = B$ on those connections.
+
+These principles can be used for any neuron model,
+$G[\cdot]$. In this study, we use the
+adaptive leaky integrate-and-fire (ALIF) model
+to match \citetext{Singh2006}.
+For each instance of each model,
+the parameters of each ALIF neuron
+is randomly selected from within
+a biologically constrained range.
+The encoders, $e_i$, are randomly selected
+directions in the encoded vector space.
+All other parameters in the simulation
+are computed from these parameters,
+or are described in section \ref{sec:dynamics}.
+
+\subsection{Adaptive control circuit}
+
+\citetext{Narayanan2009} hypothesized
+that ACC activity could be used
+to inhibit motor responses
 until the time of the cue.
-In order to verify that hypothesis,
+An alternative view is that neuronal activity
+serves to encode the duration of the action
+and that inhibition is carried out elsewhere.
+To examine these hypotheses,
 we constructed a network that
 performs the simple RT task
-using the simple non-adaptive strategy
+using the non-adaptive strategy
 of responding to cues.
@@
 with the double integrator network
-such that the final model
-is able to flexibly switch between
+to give an overall ``adaptive control circuit''
+that is able to switch between
 an aggressive timing strategy
@@
 
-\subsection{Cue-responding circuit}
+\subsubsection{Cue-responding circuit} \label{sec:cue-responding}
 
@@
 of adaptive leaky integrate-and-fire neurons
-(see Figure~\ref{fig4}).
+(see Figure~\ref{fig3}, grey box only).
 The population labelled \textit{Trigger}
@@
 The actual lever press and release
-is delayed by $\sim$250ms,
+is delayed by $\sim$200ms,
 the approximate amount of time
 it takes for muscles to respond
-to the activation of command neurons
-in primary motor cortex.  % ???travis
+to sensory input \cite{Kawato1999}.
 
@@
 the prepotent inhibition model
-proposed by \citet{Narayanan2009}.
+proposed by \citetext{Narayanan2009}.
 They suggest that a motor command
@@
 
-\subsection{Adaptive control circuit}
+\subsubsection{Adaptive control circuit}
 
@@
 of motor cortex and muscle activity.
-Since the goal is to time the action
+Since the goal is to produce the action
 as close as possible to the time of the cue,
@@
 The double integrator model described
-previously can implement such a strategy.
+previously can be combined with the
+cue-responding circuit to implement such a strategy
+(see Figure~\ref{fig3}).
 The state of the dynamical system being tracked by
-the double integrator (equation \eqref{dint-full})
-is approximately the same when the cue occurs.
+the double integrator (Equation [\ref{dint-full}])
+is approximately the same each trial when the cue occurs.
 By connecting the output of the double integrator
@@
 allows for adaptive control of action timing.
-
-\section{Emulating aging effects}
-
-The spiking neural implementation
-of the double integrator model can be used
-to emulate the effects of several pathologies.
-As a proof of concept,
-we emulate the effects of NMDA receptor degradation
-in aging \citep{Magnusson1998}.
-
-Feedforward connections
-in the adaptive control model (Figure~\ref{fig4})
-use a time constant of 10 ms,
-which is consistent with the time constant
-of the postsynaptic current
-induced by uptake of neurotransmitter
-by AMPA receptors \citep{Spruston1995}.
-Recurrent connections use a time constant of 50 ms,
-which is consistent with the PSC
-of NMDA receptors \citep{Sah1990}.
-To emulate the effects of aging on NMDA receptors,
-we decrease the synaptic strengths
-on these recurrent connections.
-
-We hypothesize that this will have the same effect
-as adding a small damping term, $D < 0$,
-in equation \eqref{dint-full}, resulting in
-\begin{align} \label{dint-damped}
-  \dot{x}_1 &= u + E + O - R x_1 - D x_1 \nonumber \\
-  \dot{x}_2 &= \alpha x_1 - R x_2 - D x_2.
-\end{align}
-This hypothesis can be tested directly
-in the double integrator model,
-and its effects on behavior tested
-in the adaptive control circuit.
+A possible extension to this model
+would incorporate synaptic plasticity
+to change the recurrent connections such that
+the cue prediction would be better
+on subsequent trials.
 
@@
 
-The dynamical system described by equation \eqref{dint-full}
-exhibits the desired behavior during simulation
-of the simple RT task.
-Figure~\ref{fig5} provides a breakdown of
-the trajectory and vector field
-throughout a correct, premature, and error trial.
-Note that the oscillation is omitted
-in this figure.
+Here we describe the results of solving
+the dynamical system in equation \eqref{dint} directly
+(i.e., not using neurons).
+This provides a general characterization
+of the expected low-dimensional dynamics
+of the high-dimensional neural simulation.
+Figure~\ref{fig4} demonstrates that
+this dynamical system exhibits
+the behavior described in section \ref{sec:dynamics} during
+correct, premature, and late trials.
+Note that this figure depicts equation \eqref{dint},
+which does not have the oscillatory term $O$.
 The final position of
@@
 the outcome of the last trial.
+
 It is clear from these trajectories
@@
 this is consistent with the results of
-\citet{Narayanan2009}.
-
-\subsection{The aged model loses adaptivity}
-
-Figure~\ref{fig6} shows trajectories for
-correct, premature, and late trials
-for the aged model (i.e., the model with a damping term,
-equation \eqref{dint-damped}).
-As the damping term increases, the final position
-of the trajectories on error trials
-becomes closer to the final position of correct trials,
-indicating that increased damping
-makes distinguishing between correct and error trials
-progressive more difficult.
-Additionally, the distance (in state space) between
-the state at which the cue occurs
-and other states visited in the trajectory
-decreases with increased damping,
-which results in more difficulty
-predicting the time of the cue
-following correct trials as well.
-Timing the foreperiod is therefore
-a poor strategy with significant aging
-(i.e., with $D \le 0.05$).
+\citetext{Narayanan2009}.
+
+These theoretical results suggest that
+the underlying, low-dimensional dynamics
+of the postulated circuit
+should be appropriate
+for capturing the qualitative features of ACC activity.
+We now examine
+both the qualitative and quantitative fit
+of the high-dimensional spiking network
+to the recorded empirical data.
 
@@
 
-Figure~\ref{fig7} shows that
+Figure~\ref{fig5} shows that
 the decoded values of a spiking neural network
@@
 in the decoded state space of the spiking model.
-Therefore, the cue can be accurately predicted,
+Therefore, the cue can be predicted,
 and the outcome of the previous trial
@@
 
-\subsection{Decreasing recurrent weights is analogous to damping}
-
-Figure~\ref{fig8} shows that
-the decoded values of a spiking neural network
-in which recurrent connections are degraded
-by a small amount
-take similar trajectories through state space
-as the damped model (Figure~\ref{fig6}).
-This supports the hypothesis
-that a small damping term
-has a similar effect as
-the degraded NMDA receptors seen in aging
-in the double integrator model.
-
-\subsection{Double integrator model has the same principal components as experimental data}
+\subsection{The double integrator model has the same principal components as the experimental data}
 
@@
 on the experimental data,
-we randomly sampled 174 neurons
+we adopted the methods of the original paper
+and randomly sampled 174 neurons
 with firing rates $> 1$ Hz from
@@
 The spike trains of these neurons
-were convolved with a Gaussian filter ($\sigma = $ 25 ms)
+were convolved with the same Gaussian filter
+as used experimentally
+($\sigma = $ 25 ms)
 and then normalized to Z-scores.
-The results are shown in Figure~\ref{fig9}.
+The results are shown in Figure~\ref{fig6}.
 The simulated principal components
-are very similar to the experimental principal components;
+are very similar to the experimental principal components
+\cite{Narayanan2009}:
 $R^2 = 0.96$ and $R^2 = 0.895$ for
@@
 the first two components in the post-late case.
-
-\subsection{Adaptive control model has similar performance as experimental subjects}
-
-The parameters of the adaptive leaky integrate-and-fire neurons
-used in all models are randomly assigned
-from biologically determined distributions.
-While the differences in neural parameters
-has little effect on the principal components
-of the double integrator model,
-they can have a significant effect
-on performance and reaction times
-in the simple RT task.
-Similarly, individual experimental subjects
-have varying performance levels and reaction times
-in the task.
-For these reasons, we analyze
-each experimental subject and instance of the model separately.
-
-Figure~\ref{fig10} shows the performance
+$R$ is a Pearson correlation coefficient in all cases.
+
+\subsection{The adaptive control model explains observed performance}
+
+Figure~\ref{fig7} shows the performance
 of the experimental subjects,
@@
 The length of the simulations were
-varied to approximately match the number of trials
-as the experimental subjects.
+varied to match the number of trials
+performed by subjects.
 The mean of the median reaction times
@@
 and 84.0\% for adaptive control models.
-Therefore, the adaptive control model
-has similar reaction times as the experimental subjects.
+Therefore, only the adaptive control model
+has statistically indistinguishable reaction times
+compared to the experimental subjects.
 The cue-responding model is significantly slower,
@@
 
-\subsection{Aged model is slower and less variable}
-
-When the aged double integrator model
-is used in the adaptive control network,
-the mean performance is 89.5\%,
-and the mean median reaction time
-is 375$\pm$47ms (see Figure~\ref{fig10} for individual results).
-These results are closer
-to the results of the cue-responding strategy,
-suggesting that the model emulating
-some effects of aging
-uses the conservative cue-responding strategy
-over the time prediction strategy
-more often than the non-aged adaptive control network.
+\subsection{Cue prediction causes premature releases}
+
+Figure~\ref{fig8} shows the decoded values
+of neural populations during representative trials
+of the cue-responding and adaptive control circuits.
+In panel C, the double integrator model
+causes the control model to respond faster than
+the cue-responding strategy.
+Premature release never occurs with
+the cue-responding circuit alone.
+While this cannot be directly compared
+with the experimental data,
+as we cannot determine
+whether the experimental subject
+is attempting to
+predict the time of the cue,
+it suggests that more premature
+releases will occur when the time of the cue
+can be predicted;
+if the foreperiod randomly varied, for example,
+our model predicts that fewer premature
+releases would occur.
 
@@
 
-The results indicate that the double integrator model
-is allowing for adaptive control of action timing.
+Our results indicate that the double integrator model
+allows for adaptive control of action timing.
 The ability to predict the time of the cue
@@
 explained by the principal components
-and the lack of noise in the
-spike density function (see Figure~\ref{fig9}).
+and the lack of noise in
+spike densities (see Figure~\ref{fig6}).
 The relative cleanness of the simulated data
@@
 to be recorded during the experiments.
-This could be simulated
+Increased variability could be simulated
 by either injecting
@@
 randomly varying dimensions.
-Additionally, the adaptive
-leaky integrate-and-fire neuron model used in this study
-has a more regular spiking pattern than
-biological neurons,
-so using a more detailed neural model
-would provide a closer match
-between experiment and simulation.
 
@@
 is that there is significant randomness
-involved in model generation,
-and no learning occurs once the model is generated.
-While each instance of the model
-can be thought of an individual
-from the population of subjects,
-one can also think of each model instance as representing
-a snapshot during a subject's
-learning of the task.
-Therefore, the poor performance of, for example,
-adaptive control model instance 10 (see Figure~\ref{fig10})
-can be explained at least two ways.
-It may be a subject
-whose recurrent connections are too strong,
-and therefore is a poor performer in the task,
-or it may be a subject
-who has not yet learned how to properly time
-the interval, and after more trials would,
-with learning, be able to make less premature errors.
-
-The results from the aged model
-form an important prediction of
-how aged subjects will perform in the simple RT task.
-It is notable, however, that the aged model
-does not make additional premature errors;
-in fact, premature errors are completely eliminated.
-This is due to the targeted manner in which
-aging is emulated (i.e., NMDA receptor degradation).
-If the model is more generally damaged,
-for example to emulate an ACC inactivation study,
-we would predict that the effects
-would be different than in the aging example.
-In the case of overall ACC damage,
-we hypothesize that the effect is more similar
-to decreasing the ``resolution'' of the state space.
-That is, it becomes harder to differentiate
-similar states, though the system
-still takes the same approximate trajectory
-through the state space.
-In this case, it becomes difficult
-to predict the time of the cue,
-and therefore additional premature errors would occur.
+involved in model generation (see section~\ref{sec:nef}).
+This results in significant variability
+in the performance of simulated subjects,
+despite all but one performing the task well
+(see Figure~\ref{fig7}).
+Also interesting is that the model is
+not plastic; the weights determined
+before the simulation do not change.
+The ability to keep track of the state
+of the system over time through recurrent connections
+is what enables the model to adapt.
+This points to a need for models to
+differentiate between adaptability
+due to recurrent activity
+and adaptability due to
+synaptic weight changes.
 
@@
 response suggested by the color;
-the context that colors should be read
+the context in which colors should be read
 would be reflected in the passive dynamics
@@
 that compensates for the word-reading response.
+Similar mappings could be proposed
+for a wide variety of ACC-monitored tasks.
 
 In conclusion, we have shown that neural integration
-is a key mechanism in the ACC's
+is likely a key mechanism in the ACC's
 ability to control the timing of actions.
@@
 which matches the performance of experimental subjects.
-When the model is manipulated
-in a similar way as the aged brain degrades,
-performance degrades and forms a prediction
-of how aged experimental subjects may perform.
-
-\bibliographystyle{plainnat}
+
+\bibliographystyle{namedplus}
 \bibliography{jneurosci2013}
@@
 
-\appendix
-
-\section*{Supplementary material}
-
-\section{Methods}
-
-\subsection{Simple reaction time task}
-
-The simple reaction time task
-(section \ref{sec:simple-rt}; Figure~\ref{fig1})
-is implemented both experimentally
-and in a simulated environment.
-
-Experimentally, Long-Evans rats perform the task
-in a MedPC box containing a lever,
-house lights, and a spout for delivering water reward.
-Rats were trained by first
-associating reward with lever presses,
-and then introducing progressively longer
-delay periods until they reached the goal delay
-of 1 second.
-They were considered to have learned the task
-when they performed the task correctly $>60\%$
-of the time.
-
-Theoretically, this study was performed
-in the Nengo simulation environment.\footnote{Nengo is open source
-  software available at \url{http://www.nengo.ca}.}
-The environment was modeled using a
-a finite state machine to track task state,
-and scalar values between -1 and 1
-to indicate the state of
-house lights, lever position, auditory cue, and reward.
-
-\subsection{Experimental data}
-
-The experimental data analyzed in this study
-is the same data presented in
-\citet{Narayanan2009}.
-The data consists of 174 neurons
-recorded from anterior cingulate cortex with microwire arrays
-in twelve male Long-Evans rats.
-For more details on the data collection,
-refer to the original paper.
-
-\subsection{Simulation details}
-
-All simulations were run
-in the Nengo simulation environment
-through Python scripts.
-Those scripts are available
-at \url{https://github.com/tbekolay}.
-
-\subsubsection{Neural Engineering Framework} \label{sec:nef}
-
-Nengo simplifies the construction
-of spiking neural networks
-using the Neural Engineering Framework (NEF; \citealp{Eliasmith2003}).
-The NEF makes the assumption
-that populations of spiking neurons
-represent real-valued vectors,
-and use a least-squares optimal method
-to transform those representations
-through the connections between
-populations of neurons.
-A population can also be recurrently connected
-in order to produce dynamics,
-such as the neural integrators
-employed in this study.
-
-Briefly, the representation scheme
-is similar to population coding,
-as proposed by \citet{Georgopoulos1986}
-but extended to $n$-dimensional vector spaces.
-Each neuron in a population is sensitive
-to a particular direction,
-called the neuron's \textit{encoder}.
-The activity of a neuron can be expressed as
-\begin{equation}
-	a = G[\alpha \mathbf{e} \cdot \mathbf{x} + J_{bias}],
-\end{equation}
-where $G[\cdot]$ is the nonlinear neural activation function,
-$\alpha$ is a scaling factor (gain) associated with the neuron,
-$\mathbf{e}$ is the neuron's encoder,
-$\mathbf{x}$ is the vector to be encoded, and
-$J_{bias}$ is the background current of the cell
-when $\mathbf{x} = 0$.
-The currently encoded value, $\mathbf{\hat{x}}$,
-can be estimated linearly.
-\begin{equation}
-  \mathbf{\hat{x}}(t) = \sum_i \mathbf{d}_i a_i(t),
-\end{equation}
-where $\mathbf{d}_i$ is the decoder,
-and $a_i$ is activity of neuron $i$.
-
-Neural activity is interpreted as a
-filtered spike train.
-\begin{equation}
-  a_i(t) = \sum_s h(t - t_s) = \sum_s e^{-(t - t_s) / \tau_{PSC}},
-\end{equation}
-where $h(\cdot)$ is the exponential filter
-applied to each spike,
-and $s$ is the set of all spikes occurring
-before the current time $t$.
-
-The decoders are found through a least-squares minimization
-of the difference between the decoded estimate
-and the actual encoded vector.
-\begin{equation}
-  \mathbf{d} = \Upsilon^{-1} \Gamma \hspace{1.8em}
-  \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
-  \Upsilon_j = \int a_j \mathbf{x} dx.
-\end{equation}
-
-Connection weights computing a function
-between two populations connected in a feedforward manner
-$f(\mathbf{x})$ can be determined by
-solving for a set of decoding weights
-for that function,
-\begin{equation}
-  \mathbf{d}^{f(\mathbf{x})} = \Upsilon^{-1} \Gamma \hspace{1.8em}
-  \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
-  \Upsilon_j = \int a_j f(\mathbf{x}) dx,
-\end{equation}
-and then computing the following weight matrix, $\omega$;
-\begin{equation}
-	\omega_{ij} = \alpha_j \mathbf{e}_j L \mathbf{d}^{f(\mathbf{x})}_i,
-\end{equation}
-where $i$ indexes the input population,
-$j$ indexes the output population,
-and $L$ is a linear operator.
-
-Dynamical systems of the form
-$\dot{\mathbf{x}} = A(\mathbf{x}) + B(\mathbf{u})$
-can be created by connecting a population
-to itself with a weight matrix
-that computes the function $f(\mathbf{x}) = \mathbf{x}$,
-setting $L = A$ on the recurrent connection,
-and receiving input $\mathbf{u}$
-from other neural populations
-and setting $L = B$ on those connections.
-
-\subsubsection{Double-integrator model}
-
-In addition to the description in the main text,
-it should be noted that \citeauthor{Singh2006}
-discuss four possible
-realizations of their double integrator model,
-all of which produce the same dynamics
-in the final output population.
-The realizations differ in whether
-the signals are internally or externally generated,
-and in whether a single multidimensional integrator
-or multiple one-dimensional integrators are used.
-
-For ease of implementation,
-we used multiple one-dimensional integrators
-because the resulting network is more stable than
-a network with a single multidimensional integrator
-\citep{Singh2006}.
-No microstimulation was done
-during the behavioral experiments
-so it is difficult to determine
-if the experimental ACC
-can also be thought of as
-a system of coupled integrators;
-preliminary simulations using
-a single multidimensional integrator
-produced PCs that matched no better
-than the coupled integrator model,
-and often worse.
-
-The principal component analysis
-shows that activity is modulated
-at the time of the lever press
-and when the outcome is determined;
-it is likely that these signals
-arrive from the motor cortex
-and other cortical areas,
-but we cannot rule out the possibility
-that these signals are internally generated
-from the presented data alone.
-
-\subsection{Analysis}
-
-Both experimental and theoretical experiments
-produced data in the form of spike trains
-and event timestamps.
-These data files were read
-by a customized version of Python-Neo.\footnote{Open source software
-  available at \url{http://neuralensemble.org/neo/}.}
-The subsequent analysis was done
-with custom Python scripts.
-All analysis scripts are available
-at \url{https://github.com/tbekolay}.
-
-\subsubsection{Neural analysis}
-
-Principal component analysis (PCA)
-is used to identify the important features
-in spike trains around certain events.
-For each neuron with an average firing rate over 1 Hz,
-we isolate the spike trains from 4 seconds prior to and
-4 seconds after each press event on a correct trial.
-Each perievent spike train is binned in 1 ms bins and is then
-convolved with a Gaussian filter ($\sigma = $ 25 ms;
-PCs are consistent with many other $\sigma$ values).
-The spike trains were slightly extended
-in order to eliminated edge artifacts.
-These spike density functions
-are normalized to Z-scores,
-and then averaged over all trials to
-produce a matrix in which each row is
-the normalized average response of a neuron
-over the perievent epoch.
-PCA was performed on that matrix
-using singular value decomposition.
-The resulting principal components
-were also normalized to Z-scores.
-The amount of variance accounted for
-by each principal component
-is computed as the square of the component's eigenvalue
-over the sum of all squared eigenvalues
-($s_i^2 / \sum s^2$).
-
-\subsubsection{Behavioral analysis}
-
-In both the experimental and theoretical studies,
-the following relevant event times are recorded:
-lever press, lever release, reward delivery,
-trigger stimulus onset, and house light extinguishing.
-Sequences of event times were used to
-identify the result of each trial.
-as shown in Figure~\ref{fig1}.
-Filtering the set of event timestamps
-for these sequences identifies each trial outcome.
-Reaction times are calculated
-only for correct trials,
-and are defined as the difference between
-trigger stimulus onset and lever release.
-
-\section{Results}
-
-\subsection{Adaptive control causes premature releases}
-
-Figure~\ref{fig11} shows the decoded values
-of neural populations during a trial in which
-the double integrator model
-causes the control model to respond faster than
-the cue-responding strategy.
-Figure~\ref{fig12} shows the decoded values
-during a trial in which
-the double integrator causes a premature response.
-Late responses, as seen in Figure~\ref{fig12},
-are identical with and without the double integrator model.
-
-\clearpage
-
-\section{Figures}
-
-\scalefigone{fig1}{0.5}{
+\section{Legends}
+
+\paragraph{Figure 1}
+
 The simple reaction time task used in the experiment.
@@
 depending on the time of lever release.
-}
-
-\scalefigone{fig2}{1.0}{
-  Experimental motivation for the adaptive control model. Data from JNP 2009.
-  (A) The simple reaction time task.
-  (B) When lever release events occurred.
-  Each trial is aligned such that the press event
-  occurs at time $t=0$.
-  All of the trials shown are correct trials.
-  Black dots represent correct trials
-  that were preceded by correct trials.
-  Red dots represent correct trials
-  that were preceded by error trails
-  (either premature or late).
-  (C) Peri-event spike rasters and histograms
-  showing different neural activity
-  based on the previous outcome.
-  In all six plots, black represents activity
-  when the last trial was correct, and red
-  represents activity when the last trial was an error.
-  The shaded grey areas note where
-  the neural activity is significantly different
-  between the two conditions (post-correct and post-error).
-  In the top row, there is elevated neural activity
-  before the press is initiated, following an error.
-  In the middle row, there is elevated neural activity
-  before the press is initiated (both)
-  and during the foreperiod (right)
-  following correct trials.
-  In the bottom row, there is elevated neural activity
-  during the whole trial (left)
-  and after the trial (right)
-  following error trials.
-  (D) Linear regression with a GLM was performed
-  on the neural activity.
-  Before the press is initiated,
-  18.7\% of neurons encoded the previous outcome,
-  compared to less than 10\% that
-  encoded the current reaction time
-  and the interaction between the two.
-  (E) Z-statistics showing which neurons were significantly
-  modulated by the previous outcome, and when.
-  (Top) Sorted by the absolute value of the Z-statistic.
-  The bottom 18.7\% of the rows have high average Z-scores
-  ($p < 0.05$).
-  (Bottom) Sorted by the peak of the Z-statistic.
-  The line is close to linear, suggesting that the
-  previous outcome information is being encoded over time
-  by the entire population.
-}
-
-\scalefigone{fig3}{1.0}{
+
+\paragraph{Figure 2}
+
 A summary of the principal component analysis done
-  in JNP 2009;
-  the motivation for investigating Singh \& Eliasmith 2006.
+in \citetext{Narayanan2009};
+the motivation for investigating \citetext{Singh2006}.
 (A) A graphical summary of principal component analysis.
@@
 is organized in a matrix, with each row
-  being the z-scored firing rates
+being the Z-scored firing rates
 of a single neuron on a single trial.
@@
 The varied lines criss-crossing each other
-  suggests that each individual neuron
+suggest that each individual neuron
 has different sensitivities
@@
 the second principal component, and vice-versa
-  ($R^2 = 0.904$ and $R^2 = 0.939$ respectively),
+(Pearson $R^2 = 0.904$ and $R^2 = 0.939$ respectively),
 in the post-correct case.
@@
 cumulative sums of each other
-  ($R^2 = 0.639$ and $R^2 = 0.676$).
+(Pearson $R^2 = 0.639$ and $R^2 = 0.676$).
 This points to neural integration
@@
 (right) on the second principal component.
-}
-
-\scalefigone{fig4}{0.8}{
+
+\paragraph{Figure 3}
+
 Block diagram of the adaptive control model.
@@
 the double integrator model,
-  from Singh \& Eliasmith.
+from \citetext{Singh2006}.
 It has been modified by accepting
@@
 provided by the environment
-  (see text for details).
+(see section \ref{sec:cue-responding}).
 (Right) Values that are not modeled
@@
 signals that are provided by the environment.
-  The lever accept the decoded
+The lever accepts the decoded
 output of the \textit{Press/Release} population,
@@
 neurons are activated (following the same delay).
-}
-
-\scalefigone{fig5}{1.0}{
+
+\paragraph{Figure 4}
+
 A detailed breakdown of the dynamics and
@@
 (A) In a correct trial, the pressing of the lever
-  sends the double integrator model into
-  a high part of the state space.
+sends the double integrator model's state into
+the upper part of the state space.
 The connection between the two integrators
 causes the passive dynamics (panel 2)
-  to have a shift towards the right.
+to push the state towards the right.
 The cue is delivered while drifting
 to the right, and the release occurs
-  near the top-right portion of the state space.
-  After the successful release,
+in the upper-right portion of the state space.
+After successful release,
 the reward turns off both integrators,
@@
 prior to the onset of the cue (panel 2).
-  Premature release causes the subject
-  to detect an error, which drives
-  the ACC to a low point in the state space.
+Premature release causes
+the house lights to extinguish,
+which drives the system to
+a point in the lower part of the state space.
 Between trials (panel 4),
 the passive dynamics of the state space
-  drive the ACC to the bottom-left portion
+drive the system to the lower-left portion
 of the space, where it will be
@@
 but the lever is not released (panel 2).
-  Late release causes the subject
-  to detect an error, which drives the ACC
-  to a low point in the state space.
-  Between trials, the passive dynamics
-  of the state space drive the ACC
-  to the same bottom-left portion of the space
-  that the premature error results in.
+Late release causes
+the house lights to extinguish,
+which drives the system
+to a point in the lower part of the state space.
+Between trials (panel 4), the passive dynamics
+of the state space drive the system
+to the same lower-left portion of the space
+that premature errors result in.
 Therefore, late and premature errors
-  cannot be distinguished on the next trial
-  in this model.
-}
-
-\scalefigone{fig6}{1.0}{
-  The effects of damping on the mathematical
-  model for correct (left), premature (middle),
-  and late (right) trials.
-  Note the different end points (marked with arrows).
-  Four damping values were simulated;
-  $D = [0, 0.025, 0.05, 0.075]$.
-  As can be seen in the middle and right panels,
-  with $D \ge 0.05$,
-  correct and error trials can no longer
-  be differentiated on the next trial
-  because the ACC is in the same state
-  regardless of the outcome of the previous trial.
-}
-
-\scalefigone{fig7}{1.0}{
-  Decoded values of the double integrator model.
-  They take the same trajectory through state
-  space as the mathematical model (see Figure 5).
+cannot be distinguished on the next trial.
+
+\paragraph{Figure 5}
+
+Decoded values of the double integrator model
+simulated in spiking neurons.
+The spiking neural model takes
+very similar trajectories through state
+space as the mathematical model (see Figure~\ref{fig4}).
 Similar to the mathematical model,
-  the ACC represents a very different state
+the spiking model represents a very different state
 after correct (left) and error (middle, right)
 trials.
-}
-
-\scalefigone{fig8}{1.0}{
-  The effect of aging through degrading recurrent connection weights
-  on the spiking neural model for correct (left), premature (center),
-  and late (right) trials.
-  Note the different end points (marked with arrows).
-  Weights were degraded by first solving for an appropriate
-  weight matrix, and then multiplying them by four values:
-  $[1.0, 0.99, 0.98, 0.97]$.
-  As can be seen in the middle and right panels,
-  by multiplying weights by $\le 0.98$,
-  correct and error trials can no longer
-  be differentiated on the next trial
-  because the ACC is in the same state
-  regardless of the outcome of the previous trial.
-}
-
-\scalefigone{fig9}{1.0}{
+
+\paragraph{Figure 6}
+
 Results of principal component analysis
 on the data generated by the double integrator model
-  for correct presses (A) post-correct trails,
+for correct presses in (A) post-correct trails,
 (B) post-premature trials, and (C) post-late trials.
-  The top panels compare the top principal components
-  of the model to the experimental ACC.
+The top row of panels compares
+the top two principal components
+of the model to those of the experimental ACC data.
 In all cases, the PCs are significantly similar
-  ($R^2 > 0.86$).
-  The middle panels show normalized
-  peri-event spike density functions
+(Pearson $R^2 > 0.86$).
+The middle row of panels shows normalized
+peri-event spike densities
 for 174 randomly sampled neurons from the
 simulated double integrator model.
-  The bottom panels show normalized
-  peri-event spike density functions
+The bottom row of panels shows normalized
+peri-event spike densities
 for all 174 neurons recorded
 during the experimental study.
-}
-
-\scalefigone{fig10}{1.0}{
+The patterns of sensitivity
+compared in the middle and bottom rows
+are similar, although the recorded neuron responses
+are more variable over time.
+
+\paragraph{Figure 7}
+
 Behavioural performance of the
-  experiment, cue-responding, adaptive control,
-  and aged adaptive control models.
-  The top panels summarize the number of
+experimental subjects and
+the cue-responding and adaptive control models.
+The top row of panels summarizes the number of
 correct, premature, and late trials
@@
 matched to the experimental subjects.
-  The bottom panels summarize the reaction times
+The bottom row of panels summarizes the reaction times
 on correct trials for each subject.
@@
 the mean of all subjects' median reaction times.
-}
-
-\scalefigone{fig11}{1.0}{
-  Decoded values from a representative correct trial
-  for the simple network (A) and the adaptive network (B).
-  (A) In the simple network, the trigger neurons
+
+\paragraph{Figure 8}
+
+Decoded values from representative trials.
+The labels above denote events;
+`P' is a lever press event,
+`C' is a cue onset event,
+`R' is a lever release event,
+`E' is an error delivery event,
+and `Rw' is a reward delivery event.
+Names of the decoded signals correspond
+to the names of populations in Figure~\ref{fig3}.
+(A) A correct trial using the cue-responding circuit.
+In the cue-responding network, the trigger neurons
 activate only once the cue is delivered.
@@
 which initiate the release of the lever.
-  There is a delay ($\sim 300ms$)
+There is a delay ($\sim 250ms$)
 between the activation of release neurons
 and the actual release of the lever.
-  (B) In the adaptive network, when
+(B) A correct trial using the adaptive control circuit.
+In the adaptive network, when
 in the appropriate state, the cue
@@
 in a faster reaction to the cue.
-}
-
-\scalefigone{fig12}{1.0}{
-  Decoded values from a representative late trial for the simple
-  network (A), and a premature trial for
-  the adaptive network (B).
-  (A) Late release occurs when
+(C) A premature trial using the adaptive control circuit.
+Premature release occurs when
+the double integrator model incorrectly predicts
+the time of the cue.
+This occurs when integration is too fast,
+and the trigger neurons are activated
+such that a lever release is effected
+before the time of the cue.
+Note that the non-adaptive cue-responding
+circuit cannot make premature releases,
+because releases are critically dependent
+on cue information.
+(D) A late trial using the cue-responding circuit.
+Late release occurs when
 the release neurons are not sufficiently
@@
 A late release in the adaptive model
-  would look identical, with little activity
+looks identical, with little activity
 in the double integrator model that predicts cues.
-  (B) Premature release occurs when
-  the double integrator model incorrectly predicts
-  the time of the cue.
-  This occurs when integration is too fast,
-  and the trigger neurons are activated
-  such that a lever release is effected
-  before the time of the cue.
-  Note that the non-adaptive cue-responding
-  circuit cannot make premature releases,
-  because releases are critically dependent
-  on cue information.
-}
+
+\section{Figures}
+
+\scalefigone{fig1}{0.5}{}
+
+\scalefigone{fig2}{1.0}{}
+
+\scalefigone{fig3}{0.8}{}
+
+\scalefigone{fig4}{1.0}{}
+
+\scalefigone{fig5}{1.0}{}
+
+\scalefigone{fig6}{1.0}{}
+
+\scalefigone{fig7}{1.0}{}
+
+\scalefigone{fig8}{1.0}{}
 

### 2013-10-23 run.py
Some work on analyzing the DS with PyDSTool

@@ -105,24 +105,29 @@ if __name__ == '__main__':
     rootdir = os.path.dirname(os.path.realpath(__file__))
     os.chdir(rootdir)
 
-    print ("Options: download_data keep convert plot combine "
+    print ("Options: download_data clean convert plot combine "
            "(dint|ctrl|all) paper")
 
-    # If we're rerunning, we'll delete existing figures
-    if 'keep' not in sys.argv:
+    if 'download_data' in sys.argv:
+        download_data()
+
+    if 'clean' in sys.argv:
         if os.path.exists('plots'):
             for f in os.listdir('plots'):
                 if f.endswith('svg'):
                     os.unlink(os.path.join('plots', f))
 
     # Use IPython parallel stuff
-    rc = parallel.Client()
+    try:
+        rc = parallel.Client()
+    except IOError:
+        print ("IPython cluster not running. Please run:"
+               "\nipcluster start --n=7")
+        sys.exit()
+
     lview = rc.load_balanced_view()
     lview.block = False
     r = []
-
-    if 'download_data' in sys.argv:
-        download_data()
 
     # Convert all CSVs to HDF5
     if 'convert' in sys.argv:

### 2013-10-23 scripts/NEFControl.py

@@ -91,7 +91,7 @@ class AdaptiveControl(DoubleIntegrator):
 
     @staticmethod
     def releasezone(x):
-        return (1. / (1 + math.exp(-100 * (x[1] - 0.8)))) * 1.2
+        return (1. / (1 + math.exp(-9 * (x[1] - 0.9)))) * 1.5
 
     def log_nodes(self, log):
         if self.control == 'adaptive':

### 2013-11-14 paper/rebuttal.tex
Revised version submitted Nov 29
- Responded to JNeurosci reviews
- Now using doit to analyze, plot etc
- Many other improvements; see the diff

\documentclass[11pt,paper=letter]{scrartcl}
\usepackage[english]{babel}
\usepackage{amssymb,amsmath}
\usepackage{namedplus}
\usepackage[font=sffamily]{quoting}
\usepackage{helvet}
\usepackage[backgroundcolor=yellow,linecolor=yellow]{todonotes}
\usepackage{hyperref}
\hypersetup{
  colorlinks = true, %Colours links instead of ugly boxes
  urlcolor = blue, %Colour for external hyperlinks
  linkcolor = blue, %Colour of internal links
  citecolor = blue %Colour of citations
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{12pt}

\begin{document}

\vspace*{1.5\baselineskip}
\begin{flushright}
  \today
\end{flushright}

Dear Editor,

Please find a revised version of
``A spiking neural integrator model of the adaptive
control of action by the medial prefrontal cortex''
attached for your consideration for publication in the
Journal of Neuroscience.

We very much appreciate the reviews
of the original submission.
Considerable effort has been made to address
all issues raised in the reviews.
In order to address these issues,
we have done additional simulations
and revised a significant portion of the text.
Three new figures and several pages
of text have been added
to clarify issues that were unclear
in the original submission,
and to present the results
of the additional simulations.

One notable change in the manuscript,
reflected in the updated title,
is our use of the term ``medial prefrontal cortex,''
or mPFC, instead of ``anterior cingulate cortex,''
or ACC, to characterize the brain area
that was the focus of our study.
This change was made in response to
Reviewer 1, who raised concerns
about functional homologies across species.
Anatomists working on the ACC
consider the prelimbic cortex
(area 32) as a cingulate region
(e.g., studies from the Barbas lab;
\citenoparens{Medalla2009}).
However, there is an unfortunate controversy
about the rat prefrontal cortex,
due in part to the Uylings review
that the reviewer mentioned.
We would like to avoid debate
about this issue in our paper,
and chose to use a more generic term
that is commonly used in rodent studies
to describe the cortical region
in the revised manuscript.

We hope that you find the manuscript
is now suitable for publication
and look forward to hearing from you soon.

Sincerely, \vspace{1em}

Trevor Bekolay, Mark Laubach, Chris Eliasmith

\clearpage
\setlength{\parindent}{1cm}
\setlength{\parskip}{3pt}

\section*{Reviewer 1}

We would like to thank the reviewer for his or her
careful reading of our manuscript
and thoughtful feedback on conceptual
and technical issues in the study.
We have addressed all points raised
in the review,
and we believe that the paper
is very much improved thanks
to Reviewer 1's comments.

\begin{quoting}
  The computational model, named ``double integrator model'' is a
  dynamical system composed of two interacting populations of spiking
  neurons. It is programmed in the Nengo simulation environment, as
  part of the Neural Engineering Framework (NEF; Eliasmith and
  Anderson, 2003) which has the potential to constitute a useful tool
  for neuroscientists to try to simulate, reproduce and better
  understand activities of large cortical networks.

  The description of the model should be strongly clarified and
  extended to enable the reader to precisely understand what is being
  simulated and to distinguish what is part of the general Nengo
  framework and what is specific to the proposed computational model.
\end{quoting}

We appreciate the comments on Nengo and the NEF.
The ``double integrator model'' is originally
from \citetext{Singh2006}.
We have adapted it for the present study
by giving it more inputs, with the most
important being the event that terminates
integration (reward feedback).
Additionally, we have made
a feedforward control network,
and used the double integrator
to control its behavior adaptively.
These distinctions have been made more clear
due to reorganization of the methods section,
and by simplifying several sections of the model.
For example, the \textit{Holding} and \textit{Trigger}
populations only existed in the neural implementation.
We have removed those and ensured that
all neural models are a direct implementation
of their corresponding dynamical systems.
All of the equations governing
the dynamical systems are included
in the revised manuscript.

\begin{quoting}
  More analyses of the results should be provided (dependence to
  parameters, robustness) to better highlight under which conditions
  the model can or cannot reproduce the results - instead of simply
  showing a version of the model that works well.
\end{quoting}

We have now included several numerical simulations
of the dynamical system (Figures 4 and 5)
to clarify the goals of the model,
and under which conditions it meets these goals.
These simulations show how the system changes
as a result of changes to the three free parameters
in the double integrator network equations.
The many parameters affecting the neural implementation
are selected partly at random;
the model performs well and
matches the experimental data
for many different sets of randomized parameters.
We have explicitly listed the distributions
from which we choose neural parameters
in order to make the model easier to reproduce.

\begin{quoting}
  The relation to previous computational models and experimental data
  of the primate ACC and rodent mPFC should be better framed and
  discussed, since it is not clear how the present simulations of one
  particular rat mPFC electrophysiological study can
  relate/complete/replace previous theories of the role of primate ACC
  in cognitive control, as the authors claim in the introduction; the
  discussion should be extended to thoroughly address each raised
  issue.
\end{quoting}

Thanks for this comment.
We must admit that some of the text
in the original manuscript was too bold
with regard to the impact of our model
of delayed response performance
by the rat mPFC.
The text was revised with this in mind,
and we hope that the changes
address Reviewer 1's concerns.

There are many models of ACC/mPFC function
available in the literature.
Given the scope of our study, it was not possible
to address all potential models.
We chose to focus on the recent PRO model
by \citetext{Alexander2011} and included
comparisons of our results
to that model in Section 5.1 and Figure 6.
We have also included Section 5.2 in the discussion
to relate our results to those from
reinforcement learning based approaches like
\citetext{Behrens2007,Alexander2011,Khamassi2012}.

\begin{quoting}
  In the introduction, the objective of the study and its potential
  contribution to important neuroscience debates should be better
  formulated. First, the authors claim that "there are currently no
  mathematical accounts of ACC function that would explain the myriad
  of behaviors modulated by the ACC." But the model presented here
  does not attempt to explain a myriad of behaviors. It only
  reproduces data from a single study (Narayanan and Laubach, 2009).
\end{quoting}

The reviewer is correct, and we should not
have written this text so strongly.
The introduction has been thoroughly
revised in the new manuscript.

\begin{quoting}
  Although I agree that there currently is not any complete model of
  ACC functions, there are several recent computational models that
  have importantly extended the initial models and that the authors do
  not seem aware of (Behrens et al., 2007; Alexander and Brown, 2011;
  Khamassi et al., 2013). I think the manuscript would greatly benefit
  from clearly stating what is new compared to these previous models.
\end{quoting}

As mentioned above, a detailed comparison
to \citetext{Alexander2011} is included
and the other two papers are now cited and discussed.

\begin{quoting}
  Can these models already explain part of the results of (Nayanan and
  Laubach, 2009)? Even if yes, I think that the dynamical aspect of
  the present model provides something more that should be better
  highlighted. For instance, I think that both the models of Alexander
  and Brown 2011 and Khamassi et al. 2013 could account for the ACC
  activity reported in Narayanan and Laubach 2009 that "change
  significantly after errors, and appear to contain information that
  an error occurred". This should be discussed.
\end{quoting}

Reviewer 1 is correct that the models presented in
\citetext{Alexander2011} and \citetext{Khamassi2012}
would be able to account for some elements
of the \citetext{Narayanan2009} data.
However, we believe that our ability
to generate spike data that can be
directly compared to the experimental data
is a unique contribution,
in additional to the dynamical aspect
that the reviewer notes.
While the information
being tracked is similar between
all of these models
(as shown in the revised manuscript's Figure 6),
we interpret that information
from a dynamical systems perspective
which provides an alternative hypothesis
of how that information might come about
(as Reviewer 1 states).
We have highlighted the dynamical aspect of the model
(particularly in Section 5.2, but also throughout),
and thank Reviewer 1 for pointing out
the similarities between these models.

\begin{quoting}
  It is not true that since ``no existing models have been implemented
  with spiking neurons [...] no models can be analyzed with the same
  methods used to analyze single neuron recordings.'' Simulations of
  simple non-spiking neural network models or even of algorithmic
  models can be analyzed with methods that consider neurons' firing
  rate in different time bins.
\end{quoting}

We respectfully disagree on this point.
While it is true that
our analyses could be performed with
firing rates in time bins,
there are more detailed analyses that can only
be performed with raw spike data.
For neural network models with
real-valued output from each neuron,
a representation scheme must be used
to represent that real-valued output
with spiking neurons;
firing rate alone is rarely sufficient.
For rate models that emit firing rates
and use a representation scheme,
the temporal dynamics in a spiking model
are significantly different;
a neural integrator, for example,
requires about twice as many
spiking neurons than rate neurons
to obtain similar performance.

In short, there are a variety
of differences between rate and spiking models
that require additional analyses
or ``translations'' of rate models
for comparison with spike data,
which are not required of spiking models.

\begin{quoting}
  I think the important thing here is that some characteristics of
  recorded neural activities can be seen as emerging properties of a
  dynamical system which a spiking neural network model enables to
  reproduce.
\end{quoting}

We agree with this statement,
and have attempted to highlight this
in the revised discussion.
As a general statement on neural modeling,
we believe that the aim should not be
to reproduce neural activities,
but instead to produce a working mathematical model.
By implementing the model in a spiking neural system,
we ensure that the model is robust
to the temporal dynamics of spiking neural networks,
and are able to compare directly
to recorded neural activities.
In this sense, the neural activities can be thought of
as ``emerging'' from the neural realization
of the mathematical model.

\begin{quoting}
  I think there is a major issue that needs to be solved concerning
  the acceptability or non-acceptability of considering that the mPFC
  (more precisely the prelimbic cortex (PL)) in rodents is the same
  thing as the ACC in primates. The state of the art of related
  previous computational models provided in the introduction only
  discusses models of ACC function in primates. It does not discuss
  any computational model of PL in rodents (e.g. see Koene and
  Hasselmo, 2005; Martinet et al., 2011). Classically, the rodent PL
  is considered as subserving homologous functions to the dorsolateral
  prefrontal cortex (DLPFC) in primates (Uylings et al., 2003). But
  people agree that the PL in rodents is less differentiated than the
  DLPFC in primates, and PL may integrate functions homologous to both
  the primate DLPFC and ACC. However, it should be taken into
  consideration that the more dorsal part of mPFC in rodents (named
  ACd/Fr2 in the Paxinos) could also partly integrate some functions
  homologous to the ACC in primates. Globally the authors should give
  more justifications. Cite for instance Seamans et al. (2008).
  Nevertheless, it is not straightforward to draw conclusions about
  ACC functions in primates from a model that has been calibrated to
  reproduce electrophysiological activities from a single study in the
  rodent PL (i.e. Narayanan and Laubach, 2009). Maybe the paper would
  benefit from a rewriting of the Introduction mainly describing
  previous experimental results and computational models of PL in
  rodents (although still briefly mentioning the implications for the
  understanding of the role of ACC in primates), and discussing in
  details the possible homologies with the ACC in primates in the
  Discussion.
\end{quoting}

The terms used for describing the medial frontal cortex
across species continues to be an issue for us (ML)
and many other researchers.
The original manuscript used
the term ``anterior cingulate cortex,'' or ACC.
Our rationale for this was that anatomists
working on the primate ACC
have commonly included the prelimbic cortex
(area 32) in the ``ACC''
(for example, studies from the Barbas lab;
\citenoparens{Medalla2009}).
However, there is an unfortunate controversy
about the rat prefrontal cortex,
due in part to the Uylings review
that the reviewer mentioned.
One of us (ML) has spent much time
discussing this issue
in conferences and workshops
(see \citenoparens{Laubach2011}).
We would like to avoid debate about this issue
in our paper, as it has little to do
with the validity of a double integrator network
as a candidate model
for explaining computations
performed by the rodent PFC
in controlling timed actions.
We chose to use a more generic term
(medial prefrontal cortex, or mPFC),
which is commonly used in rodent studies,
to describe the cortical region in the revised manuscript.
We have also limited the scope of our claims.

Following the reviewer's suggestion,
we added text to the Introduction
to review the series of studies
done by the Laubach lab
that led to collaboration
with the Eliasmith lab
on the double integrator network.

\begin{quoting}
  The presentation of previous results obtained by Narayanan and
  Laubach 2009 is too short and is not described with enough details
  to enable the reader to fully understand without reading the
  original paper. The authors should dedicate more space to describe
  the results displayed in Figure 2 and to clearly explain the
  interpretation they make. In particular, what does figure 2F tell
  us? Also, we would like to see on figure 2 the onset of the correct
  and error related PCS at the time of the feedback.
\end{quoting}

Thanks for raising this issue.
We agree that more should have been done to make Figure 2 clear,
and panel F was not sufficiently explained
in the original paper.
A paragraph was added to Section 2.5.1
to discuss the plots in panel F,
and the timing of reward feedback
has been noted in the figure.
Importantly, we note that
the analysis was only done on correct trials,
and so feedback was at the same time
for bot the post-correct and post-error trials.

\begin{quoting}
  The model should be described in more details both for specialists
  who would like to reimplement it themselves and for experimentalists
  who will read the paper. First, the terms used in equation (6)
  should be better explained. Second, it is not always clear which
  computations are specific to the proposed computational model and
  which computations are general to the NEF framework to implement any
  dynamical system and ``any neuron model''. To make the message
  clearer, I think it is important here to stick to the description of
  the proposed computational model, and to mention in a clearly
  separated paragraph what describes properties of the NEF (and of
  Nengo) that could be interesting to a reader that would like to
  program his own model.
\end{quoting}

Some confusing and unnecessary aspects of the model
have been simplified.
For example, the \textit{Holding} and \textit{Trigger} populations
were removed.
They originally existed in order to extend
the model to more general situations
in which motor actions may be
contingent on body state
through visual or proprioceptive feedback;
however, we have instead constrained the model
and linked the results only to the simple RT task
in order to simplify explanation
and avoid overgeneralizing the model
to tasks we have not modeled.

We have made several improvements to
the methods sections on
the dynamical systems and the NEF.
Included in those improvements
is a rewritten Equation (6)
that now uses matrix notation
to make the mathematical operation
(least-squares minimization) more clear.
While it was also the case
in the original paper
that NEF-specific equations
were only listed in one section,
we hope that with these improvements
it is more clear that all of the equations
in Section 2.4 are specific to the NEF,
and all equations in other sections
are part of the dynamical systems
that are simulated both directly
and implemented in spiking networks with the NEF.

Due to the lack of supplementary material,
it is difficult to include
a sufficiently detailed explanation of the NEF
to make it easily reproducible
from this paper alone.
We have opted for a comprehensive
explanation that includes
all of the equations necessary
to reproduce the results,
but with shorter explanations
than can be found
in other resources on the NEF.
However, we believe that the open source nature
of Nengo and the scripts that we use
to generate the model
make it possible for specialists
to reproduce the model.

\begin{quoting}
  The authors should avoid using the term "model" for many different
  things, since this can be very confusing for experimentalist readers
  (i.e. the computational model; "any neuron model"; "for each
  instance of each model"; ...).
\end{quoting}

The many levels of modeling that are used in this study
makes an overuse of the word ``model''
difficult to avoid;
we thank the reviewer for pointing this out.
We have attempted to minimize
the use of the word model.
Specifically, we refer to the dynamical systems
(the double integrator, cue-responding,
and adaptive control systems) and their neural
implementations as ``networks.''
All uses of the term ``model'' are prefaced
with a identifying term such as
``neuron'' or ``network'' in order
to avoid ``model'' losing its meaning
as a mathematical idealization.

\begin{quoting}
  The authors mention a mechanism for switching between an aggressive
  timing strategy and a conservative cue-response strategy. What is
  the programmed criterion for switching? Please write the used
  equations.
\end{quoting}

Thanks to Reviewer 1 for pointing out this glaring omission.
We have included the switching criterion as Equation (5)
in the revised manuscript.

\begin{quoting}
  It is important to provide analyses of the strengths and weaknesses
  of the proposed computational model, so that the reader can really
  understand what enables and what does not enable the model to
  succeed in reproducing the experimental data. What is the
  sensibility to changes in parameter values? Could the authors
  illustrate the quality of reproduction of experimental data as a
  function of different parameter sets?
\end{quoting}

We have attempted to explain the meaning of the parameters
through the newly added text in Section 3.1,
and its associated analyses in Figure 4.
We have also given more context to these parameters
in the explanations of the dynamical systems
in Sections 2.2 and 2.3.

Aside from the three parameters we set,
the model's success is dependent solely
on the ability of the integrative populations
to retain their represented values over time.
In our case, model fits are generally good
(Pearson $R^2 > 0.7$)
in any case when the random neural parameters
produce good integrators.
The ALIF neuron introduces additional dynamics
that make creating neural integrators
more difficult than
neural integrators made up of LIF neurons,
but these difficulties can be overcome
with larger populations.
The original double integrator paper
\cite{Singh2006}
discusses the robustness
of NEF-created neural integrators
in the double integrator context,
and \citetext{Eliasmith2005} discusses this issue
for many NEF-constructed attractor networks.
More detailed analyses can be found
in \citetext{Eliasmith2003}.

\begin{quoting}
  Related to this issue, on page 11, an important question is how much
  of parameter tuning was necessary to obtain the same PCs as
  experimental data? It is good to reproduce the experimental results.
  But it is even better to show analyses of different regimes or
  states the model can produce with different parameters, and to
  highlight which of these states or regimes may the experimental data
  correspond to.
\end{quoting}

We now discuss in the paper
how each parameters affects the model,
and show simulations with different sets of parameters
(Table 1 describes the ALIF parameters;
Table 2 and Figure 4 describe the
dynamical system parameters).
Section 3.1 discusses the results
of the simulations
and lists the parameters values
that were chosen.
Importantly, the parameter choices
did not require fitting,
and for most parameters,
a range of values would produce
functionally identical results.

For the experimental model fits in Figure 7,
several models were generated and the best
experimental fit was chosen.
It is possible to show the
average fits for
several instantiations of the model,
but this is not very informative because
increasing the number of neurons in the model
produces networks that fit the data well
nearly all the time,
despite many randomly generated parameters.
As noted above, the performance of the integrators
is the critical factor to the performance
of the entire model.

\begin{quoting}
  In relation to the previous point, it seems quite ad hoc and looks
  like a strong simplification to hand tune the model so that "reward
  delivery [...] resets the system to a starting state" while "an
  error trial [drives it] to a low state". Shouldn't different
  reactions of the model to correct versus incorrect trials emerge
  through learning (such as in reinforcement learning-based models of
  ACC; e.g. Alexander and Brown, 2011; Khamassi et al., 2013)? Or
  shouldn't they at least produce symmetrical effects on the model
  that it should exploit to produce different reactions? Isn't this a
  strong help provided a priori to the model to make it react in the
  expected way? Maybe I am wrong to think that this is a
  simplification. Maybe this is equivalent to the priors put in the
  cited previous computational models. But this should be at least
  discussed by the authors. For instance, doesn't such simplification
  prevent generalizations of the model to tasks involving a sequence
  of rewards and/or punishments (e.g. Tanaka et al., 2004)?
\end{quoting}

This is an important point that identifies
the central difference between
models created with the NEF
and many other computational modeling approaches,
including reinforcement learning-based approaches.
What Reviewer 1 calls hand tuning and simplification
is essentially our dynamical systems hypothesis
of how to solve the reaction time task
with a control system operating in continuous time.
While it may seem a simplification
for reward to drive the system to the origin
and error to drive it to a low part of the state space,
these reflect the operations
that are possible (and, indeed, natural) in dynamical systems.
We use these operations to implement a dynamical system
that performs the task
as we hypothesize the rodent performs the task.
We can then implement that dynamical system
in a network of spiking neural populations
through the principles of the NEF.
This spiking neural network models
the neural circuit of a rodent
that has already learned the task;
going from random connectivity to this network
is a separate branch of research.
We believe that providing the end-point of learning
is a great help to modelers
aiming to learn this kind of neural circuit
(and have shown an example of this
in other work; e.g., \citenoparens{MacNeil2011}).

In other words, the NEF provides three biologically plausible
principles for network construction
through which we can produce a neural circuit.
We have presented a way in which those principles
can be used to generate a network
that performs the simple RT task.
We have attempted to
emphasize these consideration
in various parts of the revised discussion,
most notably Sections 5.1 and 5.2.

\begin{quoting}
  Some simulation results are too briefly presented -- sometimes
  making it difficult to understand for the reader -- and some are not
  analyzed with enough deepness. On page 10, the results displayed in
  figure 4 should be better explained in the main text. A few lines
  below, in contrast to what the authors wrote, it is not clear for
  the reader from these trajectories that the time of the cue can be
  predicted. During correct trials, the cue has already occurred to
  make ACC in the right starting position for the next trial. Please
  explain what you mean here.
\end{quoting}

Simulations of the dynamical system are expanded
and simplified, as discussed below.
We have included two new figures (4 and 5)
that aim to help with understanding
how the simulations work under different sets of parameters.
More detailed equations throughout Sections 2.2 and 2.3
are also included in the revised manuscript.

\begin{quoting}
  The authors should show, either on figs 4 and 5 or on supplementary
  figures, correct, premature and late trajectories starting from the
  lower-left starting position (following an error).
\end{quoting}

Figure 4 in the revised manuscript now presents trajectories
from many different initial conditions.
It also shows how the result of the last trial
determines the initial conditions
of the next trial.

\begin{quoting}
  Can the states be decoded in the same manner when starting from such
  a different position?
\end{quoting}

The state can be decoded,
but the cue cannot be accurately predicted
unless the system starts at the origin
under normal system behavior.
This is why cue prediction does not occur
in post-error trials;
the system starts in the lower part
of the state space instead of the origin.
This is shown and expanded on in Figure 4
and Section 3.1.

\begin{quoting}
  On page 11, the authors show that the double integrator model has
  the same PCs as the experimental data. Could the authors also show
  whether the cue-responsive model also produces similar PCs as
  experimental data?
\end{quoting}

The cue-responsive network does not contain any integrative
populations, so it is highly unlikely that
the PCs would be the same as the experimental data.
However, there is an open question
(which was raised by Reviewer 2)
as to whether other network structures
would result in the same principal
components as the double integrator network.
In order to address this question,
we have added Figure 8 in the revised manuscript,
which shows that
a single integrator does not produce
similar PCs as experimental data.

\begin{quoting}
  In Narayanan and Laubach (2009), premature and late cases are
  combined in a single PC1 and PC2 couple of curves for "post-error".
  Do the PCs of the original data (from Narayanan and Laubach 2009)
  still look like the model's curves when premature and late cases are
  separated?
\end{quoting}

In \citetext{Narayanan2009}, the PC analysis
was only done using correct trials.
Error trials were not included in the analysis.
PCA done on premature trials
deviate from those based on correct trials
due to the rats responding early in the foreperiod.
PCA on late trials do not differ much
from those on correct trials,
but lack modulation during the normal period
when rats release the lever
at the end of the foreperiod.

\citetext{Narayanan2008},
which was mistakenly not cited in the previous
version of the paper, reported effects
for premature and late responses,
and showed that single units
(not population activity)
tracked outcomes during
the intertrial interval.

Importantly, while neural activity
tracked outcomes during the intertrial interval,
that activity did not differentiate
between the type of error that occurred
\cite{Narayanan2008}.
Therefore, we combine post-premature
and post-late trials when analyzing
the experimental data
in order to increase
the number of trials being analyzed,
and therefore increasing
the statistical power of the analysis
For the simulated data,
we analyze the two separately
in order to show that the model also
does not differentiate between
the two cases;
since we can simulate as many
trials as we want,
we do not need to combine
the types of error trials
in order to increase statistical power.

\begin{quoting}
  Could the authors explain how late trials can occur in the adaptive
  control model nearly as often as the other model while the former
  anticipates and responds faster than the latter?
\end{quoting}

Late trials can occur in exactly the same way
as in the cue-responding model post-error,
as the adaptive control model takes
a cue-responding strategy in that case.
Like in the cue-responding case,
late trials often correspond to non-responses,
in which the release should have occurred,
but release sensitive neurons were
not active for long enough to actually
effect the full lever press.
These errors occur at approximately
the same frequency in the cue-responding
and adaptive control models.

\begin{quoting}
  Can the model also reproduce activities during the second version of
  the task in Narayanan and Laubach 2009 where cues are presented only
  on half of the trials?
\end{quoting}

Because the cue is predicted by the double integrator network,
the model can accomplish the task even if the cue
is not actually present.
The model would not change
or produce different PCs in this case,
so we did not discuss the second version of the task
in order to keep the focus of the paper clear.

\begin{quoting}
  Do the simulation results really demonstrate that the ACC cannot
  inhibit the motor command? (Page 9). If ACC is x1+x2, then what
  brain structure correspond to H+T+P/R units? If it is also ACC, the
  authors need to first validate their ability to reproduce recorded
  ACC activity. If not, then how does the model prove that ACC is not
  inhibiting motor responses? A convincing demonstration would have
  been to simulate the two alternative hypotheses and show that one
  fails to reproduce some experimental evidence. Otherwise, the
  authors should not sell this part of the work (page 9) as a strong
  demonstration, but rather just say that with this ACC model, it is
  not necessary to exert inhibition on motor cells to work.
\end{quoting}

We have removed sections
discussing the motor inhibition hypothesis
from the paper.
We thank Reviewer 1 for pointing out
the weaknesses of this argument;
we believe that the revised discussion
proposes more concrete predictions
that come from the model.

\begin{quoting}
  The prediction extracted from the model on page 12 ("our model
  predicts that fewer premature releases would occur") appears like a
  trivial prediction. It seems to the reader that there is no need of
  a computational model to predict this. Either expand on this
  prediction or give arguments to convince the reader that this is not
  a trivial prediction.
\end{quoting}

We agree that this prediction is trivial,
and have removed it from the revised manuscript.

\begin{quoting}
  The model should not only be evaluated in its ability to reproduce
  neurophysiological recordings, but also in its ability to reproduce
  rat behavior simultaneously recorded. For instance, on page 10, it
  is stated that "if the previous trial was an error, the double
  integrator will not be following the trajectory that predicts the
  time of the cue, and therefore will not cause a lever release." Is
  it what rats do in Narayanan and Laubach (2009)? Please quantify
  such behavior and other behaviors of the model and compare with rat
  behavioral data.
\end{quoting}

Unfortunately, it is impossible to determine if rodents
are predicting the time of the cue given their behavior alone;
we could estimate this by analyzing
the reaction times on correct trials
following an error,
but due to amount of variability in reaction times,
there are not enough instances of post-error correct trials
to definitively state that reaction times are slower
on these trials compared to post-correct correct trials.

We provide a summary comparing
the behavioral statistics of
the spiking cue-responding and adaptive control networks
to the experimental subjects in Figure 10.
We believe this is the most appropriate evaluation
given the data available.

\begin{quoting}
  The manuscript would be strengthened if the authors could provide a
  series of testable experimental predictions that would allow to
  further test/refute the proposed computational model. In particular,
  what further experiments are needed to decide between the two
  mentioned alternative hypotheses about ACC inhibiting motor
  responses or this inhibition being elsewhere?
\end{quoting}

We have added Section 5.3 listing several predictions.
Since we do not discuss the motor inhibition hypothesis,
we do not propose an experiment relating to this hypothesis.
However, we do make predictions about
how neurons might respond in tasks
with different temporal dynamics,
and discuss experiments that can be done
to confirm or refute our predictions.

\begin{quoting}
  The discussion about the necessity to differentiate between
  adaptability due to recurrent activity and adaptability due to
  synaptic weight changes is very interesting and should be further
  elaborated and extended. In particular, previous ACC models rely on
  reinforcement learning mechanisms that would induce adaptability due
  to synaptic weight changes (Alexander and Brown, 2011; Khamassi et
  al., 2013). Here the model produces adaptability due to recurrent
  activity. It would be useful to discuss whether some of the results
  produced by these previous models may be also explained by recurrent
  activity. If yes, could the authors think of an experimental
  protocol that would enable to decide which of the alternative models
  is right?
\end{quoting}

We agree that differentiating
between adaptability
through recurrent activity
and through synaptic weight changes
is an important issue
that we think will become increasingly important
as more complex mPFC models are proposed.
We have added Section 5.2 to discuss
this differentiation.

In the revised discussion, we have proposed two
new experiments that could allow for
direct comparisons between neural data
and the simulated networks.
First, one could record as rats are initially trained
to perform the task and compare
the dynamics of the mPFC (and motor cortex)
with a double integrator network.
A challenge for the theory side would be
to evaluate various types of RL algorithms
that might enable a match
between theory and data.
Second, one could train rats to perform
a variant of the simple RT task
using multiple foreperiods,
and test the rats in sessions with
interleaved and blocked foreperiods.
If the mPFC neurons
and double integrator neurons track
the expected timing of action,
as we predict, then there would be
shifts in the temporal expectations
following each switch in block
(e.g., from 0.5 to 1.5 second foreperiods
over blocks of 30 trials).
To our knowledge,
no group has carried out such studies.
However, they could allow for
direct comparisons between
brain activity and theory.

\begin{quoting}
  The discussion about possible extension of the model to account for
  data in the Stroop task should be discussed in relation to previous
  models. What would such extension of the authors' model add to
  previous simulations with the model of Dehaene and Changeux (1998)?
  Globally, the discussion should be extended to thoroughly address
  each raised issue.
\end{quoting}

We have removed the possible extension of the model
to the Stroop task and instead highlighted
the model's specificity to the simple RT task.
However, we have also highlighted in Section 5.2
that the process of creating
dynamical system for performing the simple RT task
could also be done for many other tasks
and implemented using the same methods.

\section*{Reviewer 2}

We would like to thank Reviewer 2
for noting the interesting point that
the double integrator model and the published
mPFC recordings share the same
population statistics, and for inquiring
about some details in the neural circuit model.
We believe that we have been able to
address all of his or her concerns,
and, as above for Reviewer 1,
the manuscript is much improved
thanks to his or her suggestions.
In what follows,
we detail our responses to each comment.

\begin{quoting}
  I think the most compelling result in the paper is in figure 6 --
  i.e. that the network has the same principal components as the
  neural data. One question I have is to what extent this result is
  unique to a double integrator or could other networks produce these
  kind of PCs? My feeling is you have to have something like a double
  integrator but I would like to see some control simulations that
  show that. So, for example, would a single integrator network have
  those PCs? Would a random network with the same inputs have those
  PCs?
\end{quoting}

This is an interesting point,
especially when using principal component analysis
on filtered and Z-scored spike trains.
To address this issue, we added
principal component analysis results
from a single integrator network (Figure 8),
and show that it does not have
the same PCs as the double integrator
or the neural recordings.

\begin{quoting}
  I'd like to see slightly more in section 2.3.2. In particular what
  are the exact update equations for the network? In particular what
  is the exact form of $G$ what are the values of $J_{bias}$ etc ... I
  realize these details are in the referenced paper and book but I
  think it is appropriate to have them in an abridged form here.
\end{quoting}

The update equation for $G$ has been included in Section 2.4
(which was formerly Section 2.3.2),
and a table of the parameters used
(that discusses how $J_{bias}$ is determined)
has been included as Table 1.

\begin{quoting}
  I might have missed this - but how many neurons are in the model?
  You say you sample 174 but I'm not sure how many are in the actual
  model?
\end{quoting}

Thanks for catching this issue.
The simulated networks had 1200 neurons;
this has been stated in the revised text
and in Figure 3.

\begin{quoting}
  Regarding the fact that the behavior of the model is slightly too
  good (not enough errors) -- you mention adding noise to the model
  might solve this problem. I think it would be worth reporting
  whether it does in practice. Neural integrators can be fickle beasts
  so I don't think this is a given.
\end{quoting}

Our original mention of noise was in reference
to the spike density functions plotted
in (now) Figures 7 and 8.
While the principal components matched closely,
the spike densities revealed that
the simulated data is much cleaner
than the experimentally recorded data.
As Reviewer 2 suggests, we added noise
($J_\eta$, now included in equations
in Section 2.4)
and recreated Figure 7 (formerly Figure 6).
As can be seen in the revised manuscript,
the simulated spike densities now very closely resemble
the experimental spike densities,
while still maintaining a very close
principal component fit.

In regards to the model performing too well behaviorally,
this is not an issue of the noisiness of the system,
but on other modeling decisions.
In particular, we could have varied $\beta$
in order to obtain more premature errors,
which would match the behavior
of experimental subjects more closely.
However, rerunning the simulations
that are summarized in Figure 10
takes several days.
Since having a perfect fit
to the experimental performance
is not necessary to match other quantities
like reaction times,
we chose an acceptable value for $\beta$
based on direct simulations
of the dynamical system and
did not vary it while simulating the spiking implementation.

\begin{quoting}
  I think it would be nice if you could generate some testable
  predictions of the model. For example, can you say anything about
  the character of neurons that contribute to each of the principal
  components ... is there any difference between them in the model in
  terms of connectivity (are PC2 neurons more highly connected than
  PC1 neurons) and sign (are PC1 neurons more likely to be
  inhibitory). What about microstimulation or inactivation? Similarly
  for the holding, trigger and press release neurons - how would you
  know if you were recording them? As I understand it they might not
  look like figure 8 because that's the decoded signal not firing
  rate. These are just suggestions and I will leave it to the authors
  to try and find something - but the nice thing about a model is you
  can interrogate it and I think it would improve the paper to put
  down some firm predictions that can be tested.
\end{quoting}

Thanks for this suggestion,
which was also raised by Reviewer 1.
Testable predictions have been added in Section 5.3.

\begin{quoting}
  Will the Nengo scripts be available through J Neuroscience? I'm not
  quite sure what the policy is for this kind of material but it would
  be nice if it were available with the article as the github may
  change over time ...
\end{quoting}

A link to the files used
for the submitted version of the manuscript
is provided in the revised manuscript.
This link will always point to the
version of those files
used to generate the revised manuscript
even if changes are subsequently
made to those files.
Additionally, we have also included
at that link a complete copy
of the Nengo simulator
used to produce the model,
in case changes to Nengo also
affect model behavior.
With this, any computing environment
with Java should be able to
run the simulations,
and a computing environment
with Python and several open source libraries
(listed with the analysis scripts)
should be able to run
the analysis and generate
the plots shown in the paper.

\bibliographystyle{namedplus}
\bibliography{jneurosci2013}

\end{document}

### 2013-11-14 paper/jneurosci2013.tex
Removed some packages from .tex
Revised version submitted Nov 29
- Responded to JNeurosci reviews
- Now using doit to analyze, plot etc
- Many other improvements; see the diff

@@
 \usepackage{caption}
-\usepackage{soul}
 \usepackage{namedplus}
 \usepackage[usenames,dvipsnames]{color}
-\usepackage{graphicx,float,wrapfig}
+\usepackage{graphicx,float}
 \usepackage{listings}
@@
 \usepackage{hyperref}
-
+\usepackage{booktabs}
 \hypersetup{
@@
   linkcolor = blue, %Colour of internal links
-  citecolor = red %Colour of citations
+  citecolor = blue %Colour of citations
 }
@@
 \title{A spiking neural integrator model of the adaptive control of
-  action by the anterior cingulate cortex}
-\date{Abbreviated title: A spiking model of ACC adaptive control}
+  action by the medial prefrontal cortex}
+\date{Abbreviated title: A spiking model of mPFC adaptive control}
 \author{%
@@
   Mark Laubach \\
-  The John B. Pierce Lab \\
-  New Haven, CT \\ ~ \\
+  The John B. Pierce Lab and Dept. of Neurobiology \\
+  Yale University School of Medicine \\ ~ \\
   Chris Eliasmith \\
@@
 
-\textbf{Pages:} 24. \textbf{Figures:} 8. \textbf{Tables, multimedia, 3D models:} 0
-
-\textbf{Word counts:} Abstract: 164, Introduction: 445, Discussion: 479
+\textbf{Pages:} 32. \textbf{Figures:} 10. \textbf{Tables:} 2. \textbf{Multimedia, 3D models:} 0
+
+\textbf{Word counts:} Abstract: 199, Introduction: 486, Discussion: 1499.
 
@@
 \begin{abstract}
-  Reaction times can be improved by preparing movements
-  in advance and by learning the expected timing
-  of action-imperative stimuli.
+  Subjects performing simple reaction time tasks
+  can improve reaction times by learning the expected timing
+  of action-imperative stimuli
+  and preparing movements in advance.
   Success or failure on the previous trial
-  can be an important factor for determining
-  whether an animal will attempt to time
-  the stimulus or simply wait for it to occur
+  is often an important factor for determining
+  whether a subject will attempt to time
+  the stimulus or wait for it to occur
   before initiating action.
-  The anterior cingulate cortex (ACC) has been implicated
+  The medial prefrontal cortex (mPFC) has been implicated
   in enabling the top-down control of action
   depending on the outcome of the previous trial.
-  Analysis of ACC activity suggests that
-  neural integration is a key mechanism
+  Analysis of spike activity from the rat mPFC
+  suggests that neural integration is a key mechanism
   for adaptive control in precisely timed tasks.
   We show through simulation
-  that a spiking neural circuit
+  that a spiking neural network
   consisting of coupled neural integrators
   captures the neural dynamics
-  of the experimentally recorded ACC.
-  We expand on this coupled integrator model
-  to construct a spiking neural circuit that
+  of the experimentally recorded mPFC.
+  Errors lead to deviations in the normal
+  dynamics of the system,
+  a process that could enable learning
+  from past mistakes.
+  We expand on this coupled integrator network
+  to construct a spiking neural network that
   performs a reaction time task by following
-  either a timing or cue-response strategy,
-  and show that it performs a simple reaction time task
-  with similar reaction times as experimental subjects.
+  either a cue-response or timing strategy,
+  and show that it performs the task
+  with similar reaction times as experimental subjects
+  while maintaining the same spiking dynamics
+  as the experimentally recorded mPFC.
 \end{abstract}
@@
 
-The anterior cingulate cortex (ACC) has been implicated
-in a wide variety of functions,
-including conflict monitoring
-\cite{Sheth2012,Aarts2009,Botvinick2004,VanVeen2004},
+The medial prefrontal cortex (mPFC) has been implicated in regulating reinforcement learning parameters
+\cite{Amiez2005,Khamassi2012},
+conflict monitoring
+\cite{Botvinick2004,VanVeen2004,Aarts2009,Sheth2012},
 performance monitoring
-\cite{Brown2005,Carter1998,Fecteau2003,Modirrousta2008,Rushworth2004,Histed2009,Hyman2013,Horst2012},
-anticipatory control \cite{Aarts2008,Koyama2001},
+\cite{Carter1998,Fecteau2003,Rushworth2004,Brown2005,Modirrousta2008,Histed2009,Alexander2011,Horst2012,Hyman2013},
+anticipatory control \cite{Koyama2001,Aarts2008},
 arousal of the sympathetic nervous system
@@
 and control of action timing
-\cite{Muir1996,Mulert2003,Naito2013,Narayanan2009,Risterucci2003}.
+\cite{Muir1996,Mulert2003,Risterucci2003,Narayanan2009,Naito2013}.
 Conceptual models have been proposed
-that would allow the ACC to be involved
+that would allow the mPFC to be involved
 in several of these functions,
 depending on behavioral context
-(e.g., \citenoparens{Botvinick2004,Luu2003}).
-However, while there have been
-attempts at modeling the ACC
-using artificial neural networks
-\cite{Botvinick2004,Brown2005},
-there are currently no mathematical
-accounts of ACC function that would
-explain the myriad of behaviors
-modulated by the ACC.
-As well, no existing models
-have been implemented with spiking neurons,
-and therefore no models can be analyzed
-with the same methods used
-to analyze single neuron recordings.
-
-In this paper, we propose that a
+(e.g., \citenoparens{Luu2003,Botvinick2004}).
+Mathematical models
+based on reinforcement learning
+explain experimental results
+at several levels
+(e.g., \citenoparens{Alexander2011,Khamassi2012}).
+However, these reinforcement learning models
+have not been shown to explain spiking activity
+of neural ensembles recorded during
+temporally constrained behavioral tasks.
+Here, we present a dynamical spiking neural model
+that relates directly to
+both neural and behavioral data
+in a simple mPFC-dependent task.
+
+Specifically, we propose a
 two-dimensional dynamical system
-can account for ACC activity
-related to error monitoring and motor action timing.
-We test the model in a simple reaction time task
-with a predictable foreperiod and
-show that the mathematical model
-enables precise timing of actions,
-and changes behavior depending
-on the outcome of the last trial.
-The dynamical system is based
-on a multi-dimensional attractor
-network proposed by \citetext{Singh2006},
-and is implemented in a spiking neural network
-using the Neural Engineering Framework
-(NEF; \citenoparens{Eliasmith2003}).
-We show, using identical analyses
-to those employed on the experimental data,
-that simulated neural activity from this circuit
-closely matches recorded ACC activity
-and that the model can be combined with
-a simple motor control model
-to produce behavior consistent
-with behavioral studies of simple reaction time tasks.
-
-Our model also provides an alternative hypothesis
-of how the ACC accomplishes the control of action timing.
-\citetext{Narayanan2009} hypothesized
-that ACC activity inhibits motor responses
-until the time of the cue.
-Our model instead proposes that
-ACC activity encodes the duration of the action
-and that inhibition is carried out elsewhere.
-
-In this paper we apply our
-mathematical account of ACC function
-to explain error monitoring and
-the control of action timing.
-However, we suggest that the dynamical modelling
-method we employ is general and can be extended
-to explain other theories
-of high-level ACC function.
+that can account for mPFC activity
+related to error monitoring and action timing
+in rodents performing a simple reaction time task.
+Neural recordings in the motor cortex
+during the acquisition of this task
+reveal that an encoding of forthcoming errors
+develops with learning \cite{Laubach2000}.
+These activities depend on processing in the mPFC
+\cite{Narayanan2006Neuron},
+where neurons encode errors
+both prospectively and retrospectively
+\cite{Narayanan2006Neuron,Narayanan2008}.
+Analyses of population activity during the task
+show that both the mPFC and motor cortex
+are dominated by
+relatively slow fluctuations in firing rates,
+starting from the initiation of the trial
+and terminating when feedback is given
+about the trial's outcome
+\cite{Narayanan2009}.
+
+Here, we show that a specific type of network structure,
+called a double integrator network \cite{Singh2006},
+can play a central role in
+accounting for reaction time task performance.
+With analyses identical
+to the experimental study,
+we show that the neural activity produced
+by the simulated spiking network closely matches
+the experimental data from the
+\citetext{Narayanan2009} study.
+Furthermore, we propose a feedforward control system
+that can perform the simple reaction time task
+and show that the double integrator network
+can be used to modify the behavior
+of the control system based on
+the outcome of the previous trial.
+When the control system is implemented
+in a spiking neural network,
+the modified behavior closely matches
+the behavior of experimental subjects
+performing the same reaction time task.
 
@@
 
-\subsection{The reaction time task} \label{sec:simple-rt}
-
-Twelve male Long-Evans rats
-were trained to press down on a lever
+\subsection{Reaction time task} \label{sec:simple-rt}
+
+The experimental data analyzed in this study is the same
+as that in \citetext{Narayanan2009},
+where more detailed explanation of the experimental
+methods can be found.
+In that study, twelve male Long-Evans rats
+were trained to press down a lever
 for a fixed amount of time (the ``foreperiod'';
@@
 using standard methods \cite{Laubach2000}.
-After the foreperiod, a cue is presented
-(an auditory tone).
+After the foreperiod, an auditory cue is presented.
 The subject then has 0.6 seconds (the response window)
@@
 If the lever is not released within the response window,
-the trail is classified as a ``late error,''
+the trial is classified as a ``late error,''
 and is also penalized by a time-out with the house lights extinguished.
-Figure~\ref{fig1} presents a schematic view of the task.
-Subjects were considered to have learned the task
-when they performed in correctly $> 60 \%$
-of the time.
-
-The same reaction time task was simulated
+Figure~\ref{fig1}A presents a schematic view of the task.
+
+Rats included in the study reached performance criterion
+of $> 60 \%$ correct trials in 9.75 $\pm$ 1.5 sessions
+(1748 $\pm$ 343 trials).
+Microwire electrode arrays were then implanted
+into the prelimbic region of the rat cerebral cortex,
+which is generally considered to be
+part of the medial prefrontal cortex
+\cite{Laubach2011} and the anterior cingulate cortex
+in particular \cite{Medalla2009}.
+Data were recorded using
+a Plexon Many Neuron Acquisition Program
+for 3 sessions (567 $\pm$ 98 trials).
+
+\subsubsection{Simulation}
+
+In order to provide appropriate
+environmental input for
+the networks described
+in subsequent sections,
+the same reaction time task is simulated
+using a finite state machine.
+The task can be in one of seven states~($S$):
+trial start~(TS), foreperiod~(FP),
+timeout~(TO), cue~(C), response window~(R),
+reward~(Rw) and intertrial interval~(ITI).
+State transitions can be seen in
+the graphical depiction of
+the finite state machine in Figure~\ref{fig1}C.
+
+The auditory cue, house lights, and reward signals
+are provided by the environment
+based on the current task state.
+\begin{equation} \label{eq:env}
+  u_C =
+  \begin{cases}
+    1 &\text{if } S = \text{C} \\
+    0 &\text{otherwise}
+  \end{cases}
+  \quad
+  u_{TO} =
+  \begin{cases}
+    1 &\text{if } S = \text{TO} \\
+    0 &\text{otherwise}
+  \end{cases}
+  \quad
+  u_{Rw} =
+  \begin{cases}
+    1 &\text{if } S = \text{Rw} \\
+    0 &\text{otherwise}
+  \end{cases}
+\end{equation}
+The lever position does not depend on task state,
+but instead on press ($u$) and release ($u_r$)
+signals provided by a control system
+(see section~\ref{sec:control}).
+\begin{equation} \label{eq:lever}
+  L = \int [-u(t) + u_r(t)] \, dt
+\end{equation}
+
+\subsection{Double integrator network} \label{sec:dint}
+
+A key finding of the \citetext{Narayanan2009}
+study was that the temporal profiles
+of the leading principal component
+from the mPFC resembled the cumulative sum
+of the second component, and vice versa.
+This finding suggested that population activity
+in the mPFC might reflect a process of integration.
+Using the Nengo simulator \cite{BekolayInpress},
+a number of candidate networks
+were investigated and one particular network
+was found to resemble the dynamics
+of the mPFC recordings.
+This network is called a ``double integrator network,''
+and was first proposed in the context of working memory
+by \citetext{Singh2006}.
+We hypothesize that the double integrator network
+proposed by \citeauthor{Singh2006}
+will exhibit the same neural dynamics
+as the experimental mPFC
+when modified to respond
+to events occurring during the simple RT task.
+
+In order to test this hypothesis,
+we extend the double integrator network
+by providing it with multiple inputs
+that represent events occurring during
+the simple RT task
+(including those in Equation~[\ref{eq:env}]),
+while maintaining the same underlying dynamics.
+By driving the double integrator network
+to certain parts of the state space,
+we make it possible to predict
+the state of the system
+when the cue occurs during the
+reaction time task.
+
+\subsubsection{Dynamics} \label{sec:dynamics}
+
+The underlying dynamics of the double integrator network
+are captured in the following equations:
+\begin{align*}
+  \dot{x}_1 &= u \\
+  \dot{x}_2 &= \beta x_1,
+\end{align*}
+where $u$ represents an input signal
+and $\beta$ is a scalar parameter representing the strength of
+the connection between the two integrators.
+This results in a two-dimensional system
+in which $x_1$ integrates its input,
+maintaining the accumulated value over time, and
+$x_2$ integrates the value of $x_1$.
+This is the system proposed by \citetext{Singh2006},
+who predicted that this dynamical system is capable
+of time tracking;
+because $x_2$ integrates $x_1$,
+the amount of time elapsed since
+$x_1$ changes can be determined
+based on how much $x_2$ has changed.
+
+In \citetext{Singh2006}, neural recordings
+were analyzed during a single trial
+of the vibrotactile discrimination task.
+This corresponds to a single trial
+of the simple RT task
+in which the input, $u$,
+is the lever press,
+and the initial state, $x$, is $(0, 0)$.
+However, in the simple RT task we perform
+many consecutive trials.
+We therefore use other inputs from
+the environment to match
+the trajectories tracked by
+the experimental principal components
+(shown in Figure~\ref{fig2}).
+
+Since the principal components
+resemble the double integrator network
+in the post-correct case,
+we aim to reset the system to the origin point
+after correct trials.
+Figure~\ref{fig2}D indicates that
+the principal components return to their
+original state after approximately 2 seconds,
+which is the time at which
+the subject receives reward.
+Therefore, we hypothesize that
+the trial's outcome (i.e., reward delivery
+or a timeout) resets the system
+to some particular state,
+as proposed in \citetext{Narayanan2009}.
+In the case of reward delivery,
+we implement this hypothesis
+by using the reward delivery signal, $u_{Rw}$,
+to drive the system toward $(0, 0)$.
+\begin{align*}
+  \dot{x}_1 &= u - R \, u_{Rw} \, x_1 \\
+  \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
+\end{align*}
+where $R$ is a scalar parameter
+indicating how strongly the system
+is driven to the origin point.
+
+In the post-error case,
+the principal components
+in Figure~\ref{fig2}D are
+not at the origin point.
+Since the lever press increases $x_1$,
+when $x_1 > 0$ we can conclude that
+the subject is currently performing the task.
+We therefore hypothesize that $x_1 < 0$
+indicates that an error has recently occurred.
+We implement this hypothesis
+by using the house lights signal,
+$u_{TO}$, to decrease $x_1$.
+\begin{equation} \label{eq:dint}
+  \begin{aligned}
+    \dot{x}_1 &= u - R \, u_{Rw} \, x_1 - E u_{TO} \\
+    \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
+  \end{aligned}
+\end{equation}
+where $E$ is a scalar parameter
+controlling how quickly $x_1$ decreases.
+Note that $x_2$ is not directly driven by $u_{TO}$;
+however, because of the $\beta x_1$ term,
+$x_2$ will also decrease.
+
+Finally, Figure~\ref{fig2}D shows that
+the principal components are unstable
+during the intertrial interval.
+To model trial-by-trial cycles of
+excitability associated with task engagement,
+a slow oscillation ($\sim0.2$ Hz)
+was applied to the networks.
+The final dynamics of the double integrator are
+\begin{equation} \label{eq:dint-o}
+  \begin{aligned}
+    \dot{x}_1 &= u - R \, u_{Rw} \, x_1 - E u_{TO} + \sin(0.2 \cdot 2 \pi) \\
+    \dot{x}_2 &= \beta x_1 - R \, u_{Rw} \, x_2,
+  \end{aligned}
+\end{equation}
+where $\beta$, $R$, and $E$ are positive scalars.
+The effects of changing these free parameters
+are explored in Section~\ref{sec:params}.
+
+\subsection{Control networks} \label{sec:control}
+
+In order to evaluate the double integrator network's
+ability to influence action timing,
+we created a feedforward control network
+that performs the simple RT task
+by releasing the lever
+when the cue arrives;
+we refer to this feedforward system as the cue-responding network.
+We connected the double integrator network
+to the cue-responding network such that
+the lever can also be released
+by the double integrator network;
+we refer to this combined system as the adaptive control network.
+
+\subsubsection{Cue-responding network} \label{sec:cue-responding}
+
+The cue-responding network
+mimics how an experimental subject
+would react to the cue
+if the time of the cue cannot be predicted.
+The trial starts after
+the intertrial interval ends ($t_s$),
+at which point the lever is pressed.
+\begin{equation*}
+  u(t) =
+  \begin{cases}
+    1 &\text{if } t > t_s \text{ and } L > -1 \\
+    0 &\text{otherwise}.
+  \end{cases}
+\end{equation*}
+
+The lever is released
+once the cue occurs at $t_c$.
+\begin{equation} \label{eq:cue-responding}
+  u_r(t) =
+  \begin{cases}
+    1 &\text{if } t > t_c \text{ and } L < 1\\
+    0 &\text{otherwise}
+  \end{cases}
+\end{equation}
+
+\subsubsection{Adaptive control network}
+
+There is a fixed lower bound
+on reaction times in the cue-responding network
+based on the time taken to detect the cue
+and for $u_r$ to integrate to 1;
+with direct simulation of Equation~\eqref{eq:cue-responding}
+there is no delay between cue onset
+and cue detection,
+but in a spiking neural implementation
+there are signal transmission delays
+between each neural connection.
+Since the goal is to release the lever
+within the response window,
+the cue-responding strategy
+may be too slow if there are too many intervening
+connections between
+the neural population detecting the cue,
+and the neural population driving
+the muscles that eventually
+effect lever release.
+
+The double integrator network described
+previously can be combined with the
+cue-responding network to implement
+a faster strategy because
+the state of the dynamical system being tracked by
+the double integrator (Equation [\ref{eq:dint-o}])
+is approximately the same on each post-correct trial.
+If the previous trial was a correct trial,
+then the double integrator's state
+approaches (1, 1) as the time of the cue approaches,
+for some values of $\beta$.
+We therefore add an additional term
+that effects lever release
+if the predicted cue time is approaching.
+\begin{equation} \label{eq:release}
+  u_r = u_r + \frac{1}{1 + e^{-a (x_2 - b)}}
+\end{equation}
+The new term defines a sigmoid that increases
+from 0 to 1 when $x_2 \approx b$.
+The slope of the smooth increase
+from 0 to 1 is controlled by $a$;
+for very large $a$, the sigmoid
+is essentially a step function.
+We choose $a = 20$ and $b = 0.9$,
+which corresponds to a sigmoid
+that starts increasing
+when $x_2 = 0.6$ and reaches 1 when $x_2 = 1.2$.
+
+If the previous trial was an error trial,
+the double integrator
+will follow a different trajectory
+in which $1 / 1 + e^{-a (x_2 - b)} \approx 0$
+throughout the trial,
+and therefore the control system
+will only release the lever
+once the cue occurs.
+This allows the adaptive control network
+to adopt an aggressive prediction strategy
+if the last trial was successful,
+and to adopt a conservative cue-response strategy
+if the last trial was not successful.
+
+% It is likely that learning
+% takes place after error trials
+% in order to perform better
+% on subsequent trials.
+% Incorporating this in the adaptive control network
+% would be possible by modifying $\beta$
+% when errors occur;
+% however, that extension to the adaptive control network
+% is outside the scope of this paper.
+
+\subsection{Neural Engineering Framework} \label{sec:nef}
+
+The previous sections describe
+a dynamical system hypothesis
+of the mPFC and surrounding areas
+of cortex that together
+perform the simple reaction time task
+adaptively depending on the outcome
+of the previous trial.
+While these can be numerically simulated
+to observe ideal system behavior,
+we construct a spiking neural network
+implementing the above equations
+in order to test the hypothesis
+in a biologically plausible setting
+that approximates the direct numerical simulation.
+Testing this neural implementation
+allows us to determine if the approximation
+is able to perform the task
+in a manner comparable to the real brain
+by producing simulated neural data
+that can be compared directly with
+experimental neural data.
+We construct spiking neural networks
+using the principles of the Neural Engineering Framework
+(NEF; \citenoparens{Eliasmith2003})
 in the Nengo simulation environment
-using a finite state machine to track
-task state, and scalar values between -1 and 1
-to indicate the state of house lights,
-lever position, auditory cue, and reward.
-The Nengo scripts simulating this environment
-are available at
-\url{https://github.com/tbekolay/jneurosci2013}.
+\cite{BekolayInpress}.\footnote{The Nengo scripts and simulation
+  package simulating the networks and their environment are available
+  at \url{https://github.com/tbekolay/jneurosci2013/releases/tag/2013-11-29}.}
+Using the NEF, each signal
+from the dynamical systems above
+are represented with a population of spiking neurons,
+and transformed through the connections
+between those populations
+(see Figure~\ref{fig3}).
+
+The NEF enables the simulation of dynamical systems
+by making the assumption that populations of spiking neurons
+represent real-valued vectors,
+and using a least-squares optimal method
+to compute connection weights that
+transform those representations
+through the connections between
+populations of neurons.
+A population can have recurrent connections
+in order to produce dynamics,
+such as the integrative populations
+employed in this study.
+
+The NEF's representation scheme
+is a kind of population coding
+\cite{Georgopoulos1986,Salinas1994},
+extended to $n$-dimensional vector spaces.
+Each neuron in a population is sensitive
+to a particular direction,
+called the neuron's \textit{encoder}, denoted $\mathbf{e}$.
+The activity of a neuron can be expressed as
+\begin{equation} \label{eq:activity}
+  a = G[J(\mathbf{x})], \quad
+  J(\mathbf{x}) = \alpha \mathbf{e} \cdot \mathbf{x} + J_{bias} + J_\eta,
+\end{equation}
+where $G[\cdot]$ is a spiking nonlinear neural activation function,
+$\alpha$ is a scaling factor (gain) associated with the neuron,
+$\mathbf{e}$ is the neuron's encoder,
+$\mathbf{x}$ is the vector to be encoded,
+$J_{bias}$ is the background current of the cell
+when $\mathbf{x} = 0$,
+and $J_\eta$ is noise current injected
+to match experimental neural variability.
+The encoded value, $\mathbf{x}$,
+can be estimated linearly:
+\begin{equation*}
+  \mathbf{\hat{x}}(t) = \sum_i \mathbf{d}_i a_i(t),
+\end{equation*}
+where $a_i$ is the activity of neuron $i$
+and $\mathbf{d}_i$ is a decoding weight
+determined by solving a least-squares minimization
+of the difference between the decoded estimate
+and the actual encoded vector.
+
+The least-squares minimization is solved
+with the following equations.
+$\mathbf{X}$ is a set of samples from
+the representational range of
+$\mathbf{x}$ (e.g., samples from $\mathcal{U}(-1, 1)$
+in a typical scalar case).
+\begin{align*}
+  \mathbf{A} &= \begin{bmatrix}
+                 a_0(\mathbf{X}) \\
+                 a_1(\mathbf{X}) \\
+                 \vdots \\
+                 a_n(\mathbf{X})
+               \end{bmatrix} \\
+  \mathbf{d} &= \mathbf{\Gamma^{-1} \Upsilon}, \quad \text{where }
+  \mathbf{\Gamma} = \mathbf{AA^T} \quad \text{and } \mathbf{\Upsilon} = \mathbf{AX^T}
+\end{align*}
+
+Neural activity is interpreted as a
+filtered spike train.
+\begin{equation*}
+  a_i(t) = \sum_s h(t - t_s) = \sum_s e^{-(t - t_s) / \tau_{PSC}},
+\end{equation*}
+where $h(\cdot)$ is an exponential filter
+modeling postsynaptic current that is
+applied to each spike,
+and $s$ is the set of all spikes occurring
+before the current time $t$.
+
+Connection weights implementing a linear transform
+of the encoded vectors of a population
+can be computed as
+\begin{equation*}
+  \omega_{ij} = \alpha_j \mathbf{e}_j L \mathbf{d}_i,
+\end{equation*}
+where $i$ indexes the input population,
+$j$ indexes the output population,
+and $L$ is a linear operator.
+Nonlinear transformations can be computed
+by solving for decoders that minimize
+the difference between the decoded estimate and
+a function of the encoded vector;
+i.e., $\mathbf{\Upsilon} = \mathbf{A} f(\mathbf{X^T})$.
+
+Dynamical systems of the form
+$\dot{\mathbf{x}} = A(\mathbf{x}) + B(\mathbf{u})$
+can be implemented by connecting a population
+to itself with the linear operator
+$L = A$ on the recurrent connection,
+and receiving input $\mathbf{u}$
+from other populations and
+setting $L = B$ on those connections.
+
+The above principles are not dependent
+on a particular neuron model, $G[\cdot]$.
+In this study, we use the
+adaptive leaky integrate-and-fire (ALIF) model
+\cite{Koch1999} to match \citetext{Singh2006}.
+This model is governed by the equations:
+\begin{align*}
+  \frac{dV}{dt} &= - \frac{V (1 + RG_{adapt}) - J(\mathbf{x})
+    R}{\tau^{RC}} \\
+  \frac{dG_{adapt}}{dt} &= - \frac{G_{adapt}}{\tau_{adapt}} \\
+  \text{if } V > V_{th}&, \:\: G_{adapt} = G_{adapt} + G_{inc},
+\end{align*}
+where $R$ is the leak resistance,
+$\tau^{RC}$ is the RC time constant,
+and $G_{adapt}$ tracks how much
+the conductance is modulated over time.
+$G_{inc}$ and $\tau_{adapt}$
+are parameters affecting $G_{adapt}$.
+
+For each instance of each spiking neural network,
+the parameters of each ALIF neuron
+are randomly selected from
+a biologically plausible range
+(e.g., the maximum firing rate,
+$\mathcal{U}(10\text{Hz}, 50\text{Hz})$,
+produces average firing rates
+that match \citenoparens{Narayanan2009}).
+The encoders, $\mathbf{e}_i$, are
+randomly selected from
+a uniform distribution of directions
+in the encoded vector space.
+Table~\ref{tab:nef-param} summarizes
+the parameters specific
+to the neural implementation.
+All other parameters are described
+in Sections~\ref{sec:dint} and \ref{sec:control}.
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{cccccc}
+      \toprule
+      Maximum firing rate & $J_\eta$ & $\tau^{RC}$ & $\tau_{ref}$ &
+        $\tau_{adapt}$ & $G_{inc}$ \\
+      \midrule
+      $\mathcal{U}(10\text{ Hz}, 50\text{ Hz})$ &
+      $\mathcal{U}(-0.2\text{ nA}, 0.2\text{ nA})$ & 20ms & 1ms & 10ms &
+        $\mathcal{U}(0.001, 0.02)$ \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{Parameters used in the neural implementation
+    of the double integrator and control networks.
+    Maximum firing rate and $J_\eta$ are ensemble-level parameters
+    as they apply to any neuron model $G[\cdot]$.
+    The maximum firing rate is used to determined the gain ($\alpha$)
+    and bias currents ($J_{bias}$) associated with neuron
+    (and used in Equation~[\ref{eq:activity}]).
+    All other parameters apply specifically to the ALIF neuron model.}
+  \label{tab:nef-param}
+\end{table}
 
@@
 
+Both experimental and simulated experiments
+produced data at the neural and behavioral levels.
+The analyses below were done
+on both experimental and simulated data,
+unless otherwise noted.
+
 \subsubsection{Neural data}
 
-The experimental data analyzed in this study is the same
-as that in \citetext{Narayanan2009}, in which
-the authors recorded the anterior cingulate cortex (ACC) and motor cortex
-of rodents performing this simple reaction time task.
-(The ACC recordings were made in the prelimbic cortex,
-which is a component of the ACC.
-The area was referred to as the dorsomedial PFC
-in the original paper.)
-\citeauthor{Narayanan2009} used principal component analysis
-(PCA) to characterize common firing patterns
-in the ACC during the task.
-
+Experimental spike trains from Long-Evans rats
+were determined by online identification
+with an oscilloscope and audio monitor,
+and offline spike sorting
+using Plexon software.
+Spike sorting was based on principal component analysis
+(PCA; described below) and waveform shape.
+In all, 174 single units were identified
+in rodent mPFC and analyzed together
+\cite{Narayanan2009}.
+
+Simulated spike trains were generated
+by the Nengo simulator.
+The populations representing signals
+in the double integrator and control
+networks (see Sections~\ref{sec:dint} and \ref{sec:control})
+contained 1200 adaptive leaky integrate-and-fire neurons
+(see Figure~\ref{fig3});
+174 of the neurons in an analyzed population
+were randomly selected
+in order to match the number of units
+from the experimental study.
+
+Spike trains are analyzed
+with principal component analysis (PCA),
+following procedures used in \citetext{Narayanan2009}
+to characterize common firing patterns
+in the mPFC during the task.
 In the PCA, only neurons with an average
@@
 Spike trains from 4 seconds prior to and 4 seconds after
-each press event on a correct trial are isolated.
-Each perievent spike train is binned in 1 ms bins and is then
+each press event are isolated.
+The perievent spike train is binned in 1 ms bins and
 convolved with a Gaussian filter ($\sigma = $ 25 ms;
@@
 the normalized average response of a neuron
-over the perievent epoch.
+4 seconds before and after a lever press.
 Singular value decomposition
@@
 to isolate the principal components,
-which are then normalized to Z-scores.
+which are normalized to Z-scores.
 The amount of variance accounted for
@@
 
-As shown in Figure~\ref{fig2},
+As previously described in \citetext{Narayanan2009}
+and summarized in Figure~\ref{fig2},
 the first two principal components
 of the experimental data
-are modulated around the task events
+are modulated around the lever press
 and appear to track
@@
 \citeauthor{Narayanan2009} also found that the neural activity
-in the ACC changed significantly depending on
+in the mPFC changed significantly depending on
 the outcome of the previous trial;
@@
 is highly correlated with the integral
-of the second principal component,
-and vice versa.
+(i.e., cumulative sum) of the second principal component
+during the post-correct trial
+(Figure~\ref{fig2}E).
+Since the first two principal components
+account for over half of the variance
+in the neural data (Figure~\ref{fig2}B),
+we hypothesize that the mPFC
+represents the task state (PC1)
+and the integral of the task state (PC2).
+Implementing integration
+in a population of neurons has
+been widely explored
+\cite{Seung1996,Shadlen2001,Wang2002,Eliasmith2005}.
 In the post-error case,
-these correlations are significantly weaker.
-This result suggests that
-the ACC is performing some kind of integration,
-and that integration depends
-on prior behavioral outcomes.
+these correlations are significantly weaker,
+indicating that the integration
+taking place in the mPFC
+depends on prior outcomes.
+The firing patterns of individual neurons,
+related to the two leading PCs,
+are shown for post-correct trials in Figure~\ref{fig2}F.
+Nearly equal fractions of neurons
+became more or less active at times
+when the PCs were maximally positive or negative.
+In these plots, neural activity
+is synchronized to the start of the trial (time 0).
+The cue was presented at 1 second.
+The animals were required to respond
+to the cue with a reaction time less than 0.6 second,
+and feedback about the outcome
+was given approximately 0.1 second later.
+The range of feedback times are indicated
+near the upper right corner of each plot.
+Importantly, the trials used for this analysis,
+and shown in Figure~\ref{fig2}, were all correct trials.
+The only difference between trial types
+was the outcome of the previous trial
+(correct or incorrect).
+As in \citetext{Narayanan2008},
+neural activity did not track
+the type of error that occurred
+(premature or late).
+Population activity was not analyzed
+for these subtypes of errors,
+as they were present in unequal fractions of trials
+over rats and were much less frequent
+than the previously correct trials.
 
@@
 In both the experimental and theoretical studies,
-the following relevant event times are recorded:
-lever press, lever release, reward delivery,
-trigger stimulus onset, and house light extinguishing.
-Sequences of event times were used to
-identify the result of each trial.
-Filtering the set of event timestamps
-for these sequences identifies each trial outcome.
+we record five relevant event times:
+lever press ($t_p$), lever release ($t_r$), reward delivery ($t_{Rw}$),
+cue onset ($t_c$), and house lights extinguishing ($t_{TO}$).
+A correct trial is identified by
+the sequence lever press, cue onset, lever release, reward delivery;
+a premature trial is identified by
+lever press, lever release, house lights extinguishing;
+and a late trial is identified by
+lever press, cue onset, house lights extinguishing.
+Trial outcomes are identified
+from the overall sequence of events
+by filtering for these sequences.
 Reaction times are calculated
@@
 and are defined as the difference between
-trigger stimulus onset and lever release.
-
-\subsection{Double integrator model}
-
-The double integrator model
-was proposed in the context of working memory
-by \citetext{Singh2006}.
-Their model uses multidimensional neurons
-in integrative populations
-to exhibit time-varying activity
-during a vibrotactile discrimination task.
-It relies on two coupled integrators,
-whose activity resembles
-the two strong principal components
-in the ACC data.
-The model employs
-the Neural Engineering Framework (NEF; \citenoparens{Eliasmith2003})
-to implement a spiking neural network
-that captures the dynamics of the
-integrative dynamical system.
-We hypothesize that a model
-similar to that proposed by \citeauthor{Singh2006}
-will exhibit the same neural dynamics
-as the experimental ACC in the simple RT task.
-
-\subsubsection{Dynamics} \label{sec:dynamics}
-
-The dynamics of the double integrator model
-are captured in the following dynamics equations:
-\begin{align*}
-  \dot{x}_1 &= u \\
-  \dot{x}_2 &= \alpha x_1,
-\end{align*}
-where $u$ represents an input signal
-and $\alpha$ is the strength of
-the connection between the two integrators.
-This results in a two-dimensional system
-in which $x_1$ integrates its input,
-maintaining the accumulated value over time, and
-$x_2$ integrates the value of $x_1$.
-\citetext{Singh2006} have proposed that this system
-enables time tracking.
-Because $x_2$ integrates $x_1$,
-the amount of time elapsed since
-$x_1$ changes can be determined
-based on how much $x_2$ has changed.
-
-In the simple RT task,
-we want to keep track of how much time
-has elapsed since the subject
-has pressed the lever.
-While the lever press event could be
-used as the input to the system,
-since we aim to embed this system
-in a spiking neural model,
-we instead use the activity of the motor neurons
-that effect the lever press as
-the input, $u$.
-The system is also modified by signals
-representing the outcome of a particular trial;
-on a correct trial, reward delivery ($R > 0$)
-resets the system to a starting state.
-On an error trail, a second input signal ($E < 0$)
-drives $x_1$ to a low state.
-When implemented in a spiking network,
-this causes $x_2$ to saturate at a low value,
-effectively breaking the integration,
-as is seen in Figure~\ref{fig2} in post-error trails.
-This results in the following overall dynamics.
-\begin{align} \label{dint}
-  \dot{x}_1 &= u + E - R x_1 \nonumber \\
-  \dot{x}_2 &= \alpha x_1 - R x_2.
-\end{align}
-See Figure~\ref{fig3} for a graphical
-depiction of the double integrator model.
-
-The dynamics of Equation \eqref{dint}
-are predictable for certain values
-of $\alpha$, $E$, and $R$.
-$\alpha$ defines how quickly
-the system drifts during the task;
-we set it such that $x_2 \approx 1$
-at the time of release,
-which occurs when $\alpha = 0.04$.
-If $R$ is sufficiently big,
-then when reward is delivered
-the system will be driven to the origin.
-This occurs when $\int_0^{t_R} R\,dt \ge 1$.
-We set the length of reward delivery,
-$t_R$, to 1 second to reflect the pace
-of the experimental subjects,
-and we set the reward magnitude,
-$R$, to 0.15 so that $\int_0^{t_R} R\,dt \ge 1$.
-Larger values of $R$ would be functionally identical,
-as $x_1 = x_2 = 0$ when $R \ge 0.15$ for $t_R$ seconds.
-If $E$ is sufficiently low,
-then when error is detected,
-the system will be driven to
-the lowest $x_1$ value that can
-be represented by the population
-of spiking neurons.
-This exploits the saturating nature
-of spiking neurons in order to
-obtain stability in this
-mathematically unstable system.
-If $\int_0^{t_E} E\,dt \le -1$,
-then the system will saturate
-at a low value for $x_1$.
-We set $t_E = 2 s$ and $E = -0.06$.
-
-To summarize, there are five parameters
-in this dynamical system:
-$t_R$ and $t_E$ are set based on the pace
-of experimental subjects;
-$R$ and $E$ have little behavioral
-consequence if $R \ge 0.15$ and
-$E \le -0.06$ and are set
-to match the experimental principal components;
-$\alpha$ is a free parameter
-that critically affects the model's success
-and is set such that $x_2 \approx 1$
-at the time of release.
-
-In addition to the above system,
-we inject a slowly oscillating signal into $x_1$.
-Having made the parameter choices
-described above,
-the system is stable during the intertrial
-interval in both post-correct and post-error
-cases.
-However, the experimental data shown in
-Figure~\ref{fig2} exhibits some activity during
-the intertrial interval;
-we capture this additional activity
-by introducing a simple oscillation,
-$O = \sin(2 \pi f t + \phi)$,
-where the frequency $f$ is low ($\le 0.4$Hz).
-While there is experimental evidence for
-a slow oscillations like $O$ in various brain regions
-\cite{Buzsaki2004}, this injected oscillation
-does not affect the behavioral performance of the network,
-but instead models ACC activity
-not being explicitly modelled in
-the remainder of the circuit.
-This activity could reflect
-other functions that the ACC is performing,
-or could reflect the overall pace
-of the simple RT task.
-The final dynamics are
-\begin{align} \label{dint-full}
-  \dot{x}_1 &= u + E + O - R x_1 \nonumber \\
-  \dot{x}_2 &= \alpha x_1 - R x_2.
-\end{align}
-
-\subsubsection{Neural Engineering Framework} \label{sec:nef}
-
-Equation~\eqref{dint-full} is implemented
-in a spiking neural network using
-the principles of the Neural Engineering Framework
-(NEF; \citenoparens{Eliasmith2003}).
-The NEF enables the simulation of dynamical systems
-like Equation~\eqref{dint-full} by
-making the assumption that populations of spiking neurons
-represent real-valued vectors,
-and using a least-squares optimal method
-to transform those representations
-through the connections between
-populations of neurons.
-A population can have recurrent connections
-in order to produce dynamics,
-such as the neural integrators
-employed in this study.
-
-The NEF's representation scheme
-is similar to population coding,
-as proposed by \citetext{Georgopoulos1986},
-but extended to $n$-dimensional vector spaces.
-Each neuron in a population is sensitive
-to a particular direction,
-called the neuron's \textit{encoder}.
-The activity of a neuron can be expressed as
-\begin{equation}
-  a = G[\alpha \mathbf{e} \cdot \mathbf{x} + J_{bias}],
-\end{equation}
-where $G[\cdot]$ is a nonlinear neural activation function,
-$\alpha$ is a scaling factor (gain) associated with the neuron,
-$\mathbf{e}$ is the neuron's encoder,
-$\mathbf{x}$ is the vector to be encoded, and
-$J_{bias}$ is the background current of the cell
-when $\mathbf{x} = 0$.
-The currently encoded value, $\mathbf{\hat{x}}$,
-can be estimated linearly:
-\begin{equation}
-  \mathbf{\hat{x}}(t) = \sum_i \mathbf{d}_i a_i(t),
-\end{equation}
-where $\mathbf{d}_i$ is the neuron's decoder,
-and $a_i$ is activity of neuron $i$.
-
-Neural activity is interpreted as a
-filtered spike train.
-\begin{equation}
-  a_i(t) = \sum_s h(t - t_s) = \sum_s e^{-(t - t_s) / \tau_{PSC}},
-\end{equation}
-where $h(\cdot)$ is the exponential filter
-applied to each spike,
-and $s$ is the set of all spikes occurring
-before the current time $t$.
-
-Decoders are found through a least-squares minimization
-of the difference between the decoded estimate
-and the actual encoded vector.
-\begin{equation}
-  \mathbf{d} = \Upsilon^{-1} \Gamma \hspace{1.8em}
-  \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
-  \Upsilon_j = \int a_j \mathbf{x} dx.
-\end{equation}
-
-Connection weights to linearly transform these encoded vectors
-can then be computed as
-\begin{equation}
-  \omega_{ij} = \alpha_j \mathbf{e}_j L \mathbf{d}_i,
-\end{equation}
-where $i$ indexes the input population,
-$j$ indexes the output population,
-and $L$ is a linear operator.
-
-% We only do linear transformations I think?
-
-% Connection weights computing a function
-% between two populations connected in a feedforward manner
-% $f(\mathbf{x})$ can be determined by
-% solving for a set of decoding weights
-% for that function,
-% \begin{equation}
-%   \mathbf{d}^{f(\mathbf{x})} = \Upsilon^{-1} \Gamma \hspace{1.8em}
-%   \Gamma_{ij} = \int a_i a_j dx \hspace{1.8em}
-%   \Upsilon_j = \int a_j f(\mathbf{x}) dx,
-% \end{equation}
-% and then computing the following weight matrix, $\omega$;
-% \begin{equation}
-%   \omega_{ij} = \alpha_j \mathbf{e}_j L \mathbf{d}^{f(\mathbf{x})}_i,
-% \end{equation}
-% where $i$ indexes the input population,
-% $j$ indexes the output population,
-% and $L$ is a linear operator.
-
-Dynamical systems of the form
-$\dot{\mathbf{x}} = A(\mathbf{x}) + B(\mathbf{u})$
-can be created by connecting a population
-to itself with the linear operator
-$L = A$ on the recurrent connection,
-and receiving input $\mathbf{u}$
-from other populations and
-setting $L = B$ on those connections.
-
-These principles can be used for any neuron model,
-$G[\cdot]$. In this study, we use the
-adaptive leaky integrate-and-fire (ALIF) model
-to match \citetext{Singh2006}.
-For each instance of each model,
-the parameters of each ALIF neuron
-is randomly selected from within
-a biologically constrained range.
-The encoders, $e_i$, are randomly selected
-directions in the encoded vector space.
-All other parameters in the simulation
-are computed from these parameters,
-or are described in section \ref{sec:dynamics}.
-
-\subsection{Adaptive control circuit}
-
-\citetext{Narayanan2009} hypothesized
-that ACC activity could be used
-to inhibit motor responses
-until the time of the cue.
-An alternative view is that neuronal activity
-serves to encode the duration of the action
-and that inhibition is carried out elsewhere.
-To examine these hypotheses,
-we constructed a network that
-performs the simple RT task
-using the non-adaptive strategy
-of responding to cues.
-We then combined that network
-with the double integrator network
-to give an overall ``adaptive control circuit''
-that is able to switch between
-an aggressive timing strategy
-and a conservative cue-response strategy
-depending on the outcome of the previous trial.
-
-\subsubsection{Cue-responding circuit} \label{sec:cue-responding}
-
-The cue-responding circuit is composed
-of three populations
-of adaptive leaky integrate-and-fire neurons
-(see Figure~\ref{fig3}, grey box only).
-The population labelled \textit{Trigger}
-is sensitive to the cue.
-The population labelled \textit{Holding}
-tracks whether the lever is currently being held down.
-The population labelled \textit{Press/Release}
-causes lever presses and releases
-through activation of different subsets
-of its neurons.
-The actual lever press and release
-is delayed by $\sim$200ms,
-the approximate amount of time
-it takes for muscles to respond
-to sensory input \cite{Kawato1999}.
-
-The \textit{Holding} neurons project
-to the \textit{Trigger} population
-to provide the behavioral context
-(i.e., \textit{can} I release the lever?)
-The \textit{Trigger} neurons send projections
-to the \textit{Press/Release} neurons
-upon a cue that occurs in the correct context
-(i.e., \textit{should} I release the lever?)
-
-The context-driven design of this circuit
-provides an alternative theory to
-the prepotent inhibition model
-proposed by \citetext{Narayanan2009}.
-They suggest that a motor command
-is activated at the same time
-as an inhibitory waiting signal,
-and it is the removal of that inhibition
-that produces the motor command.
-This model proposes that
-an excitatory context signal is activated,
-and the pairing of that context signal
-with the cue produces the motor command.
-This leads to an alternative hypothesis
-for the function of ACC activity:
-the ACC tracks the behavioral context,
-and is able to inhibit or excite motor cortex,
-but does not always necessarily do so.
-
-\subsubsection{Adaptive control circuit}
-
-A weakness of the contextual cue-responding circuit
-is that there is a fixed lower-bound
-on the reaction time
-based on the time taken to detect the cue
-and for the delay between activation
-of motor cortex and muscle activity.
-Since the goal is to produce the action
-as close as possible to the time of the cue,
-the cue-responding strategy may be too slow,
-and so a predictive approach must be employed.
-
-The double integrator model described
-previously can be combined with the
-cue-responding circuit to implement such a strategy
-(see Figure~\ref{fig3}).
-The state of the dynamical system being tracked by
-the double integrator (Equation [\ref{dint-full}])
-is approximately the same each trial when the cue occurs.
-By connecting the output of the double integrator
-to the \textit{Trigger} population,
-the double integrator can cause a lever release
-when the state of the double integrator
-is slightly before the predicted
-time of the cue.
-
-However, if the previous trial was an error,
-the double integrator will not
-be following the trajectory
-that predicts the time of the cue,
-and therefore will not cause a lever release.
-The circuit then naturally
-uses the conservative
-cue-responding strategy.
-Switching between these two strategies
-depending on the outcome
-of the previous trial
-allows for adaptive control of action timing.
-A possible extension to this model
-would incorporate synaptic plasticity
-to change the recurrent connections such that
-the cue prediction would be better
-on subsequent trials.
+cue onset and lever release.
 
@@
 
-\subsection{Model dynamics during the simple RT task}
-
-Here we describe the results of solving
-the dynamical system in equation \eqref{dint} directly
-(i.e., not using neurons).
+Here we describe the results of simulating
+the dynamical systems in Sections~\ref{sec:dint}
+and \ref{sec:control} directly.
 This provides a general characterization
@@
 of the high-dimensional neural simulation.
-Figure~\ref{fig4} demonstrates that
-this dynamical system exhibits
-the behavior described in section \ref{sec:dynamics} during
-correct, premature, and late trials.
-Note that this figure depicts equation \eqref{dint},
-which does not have the oscillatory term $O$.
-The final position of
-the trajectory becomes the starting position
-of the next trial, and therefore
-provides information about
-the outcome of the last trial.
-
-It is clear from these trajectories
-that the time of the cue
-can be predicted,
-as the ACC is only in the predictive state
-when the cue occurs on correct trials.
-It is also clear that the
-outcome of the previous trials
-is easily decoded based
-on the state after the previous trial.
-Note that premature trials
+
+\subsection{The double integrator can predict cue time} \label{sec:params}
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{ccccc}
+      \toprule
+      $\beta$ & $R$ & $E$ \\
+      \midrule
+      0.44 & 2 & 0.5 \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{Parameters used to simulate the dynamical system
+    described in Equation~\eqref{eq:dint}.
+    Appropriate values were found through simulation;
+    see Figure~\ref{fig4}.}
+  \label{tab:sim-param}
+\end{table}
+
+The purpose of the double integrator
+is to predict the time at which the cue occurs.
+Figure~\ref{fig4} shows that this is possible.
+The parameter $\beta$ controls how fast
+the second dimension of the double integrator state
+($x_2$) increases (see Equation~\eqref{eq:dint}),
+resulting in different states at $t_c$
+(see Figure~\ref{fig4}B);
+since this parameter is encoded in the strength
+of the connection between the two integrators,
+this is a natural target for learning
+after errors.
+Given a particular value for $\beta$,
+the state of the system at the time of the cue,
+$t_c$, can be predicted given
+the initial conditions of the double integrator
+at the start of the trial;
+values at $t_c$ for $\beta=0.44$ are shown
+in Figure~\ref{fig4}A.
+
+The state of the system at the start of a trial
+(i.e., the initial conditions of the dynamical system)
+depend on the outcome of the previous trial.
+Equation \eqref{eq:dint} describes how the state changes
+based on the two possible outcomes,
+reward or light extinguishing.
+For sufficiently high values of $R$
+(see Figure~\ref{fig4}C),
+the system starts the next trial
+near the origin,
+resulting in the system predicting
+the time of the cue,
+and therefore employing the adaptive strategy.
+For sufficiently high values of $E$
+(see Figure~\ref{fig4}D),
+the system starts the next trial near $(-1, -1)$,
+resulting in the system being unable to predict
+the time of the cue (see Figure~\ref{fig4}A),
+and therefore employing the cue-responding strategy.
+Note that correct and error trials
+can be distinguished,
+and therefore used to modify behavior
+and drive learning;
+however, premature trials
 and late trials cannot be distinguished
-on the next trial;
-this is consistent with the results of
+on the next trial,
+which is consistent with the results of
 \citetext{Narayanan2009}.
 
-These theoretical results suggest that
-the underlying, low-dimensional dynamics
-of the postulated circuit
+Appropriate parameter values were determined
+using these simulations
+and used for all subsequent simulations,
+unless otherwise noted.
+The values used are listed in
+Table~\ref{tab:sim-param}.
+
+\subsection{The adaptive control network can react faster}
+
+Equation~\eqref{eq:release} defines a signal
+that causes the network to release the lever
+when the predicted cue approaches.
+Figure~\ref{fig5} shows a representative
+correct trial from the cue-responding
+and adaptive control networks,
+as well as a trial in which
+the cue was mispredicted,
+and the subsequent corrected trials
+following the error.
+Importantly, Figure~\ref{fig5}B
+shows the adaptive control network
+can release the lever
+at the exact time of the cue.
+
+\subsection{The double integrator state can drive learning}
+
+The state of the system reflects the outcome
+of the previous trial, and therefore
+can affect behavior on the next trial.
+The system can also be used to drive
+learning in order to maximize long-term reward.
+While the proposed system
+does not employ learning,
+previous work has proposed that the
+mPFC drives learning of action outcomes.
+Specifically, \citetext{Alexander2011} have proposed
+the PRO model, in which the mPFC tracks the predicted values
+of action outcomes in order to compute
+prediction errors that drive learning.
+While the signals tracked
+by the double integrator model
+exist to modify action timing,
+the signals used to drive learning
+in PRO can be computed from
+the double integrator model as well,
+as shown in Figure~\ref{fig6}.
+The only difference that cannot
+be attributed to the task
+is that the state is tracked throughout
+the trial in the PRO model,
+while in the double integrator,
+reward turns off the integrator
+responsible for tracking task state.
+Despite this difference,
+the similarity in signals
+tracked by the two models
+suggests that the double integrator,
+which we use to adaptively control action timing,
+can also be used to drive learning.
+
+\section{Simulation results}
+
+The above theoretical results suggest that
+the underlying low-dimensional dynamics
+of the postulated network
 should be appropriate
-for capturing the qualitative features of ACC activity.
+for capturing the qualitative features of mPFC activity.
 We now examine
@@
 
-\section{Simulation results}
-
-\subsection{Spiking networks follow predicted trajectories}
-
-Figure~\ref{fig5} shows that
-the decoded values of a spiking neural network
-with the same dynamics as equation \eqref{dint-full}
-also take a similar trajectory through
-state space in correct, premature, and late trials.
-Despite the introduction of the oscillation
-and the additional noise present
-in spiking neural systems,
-behaviorally relevant states
-are significantly separated
-in the decoded state space of the spiking model.
-Therefore, the cue can be predicted,
-and the outcome of the previous trial
-has a significant effect
-on the trajectory taken during a trial.
-
-\subsection{The double integrator model has the same principal components as the experimental data}
-
-In order to perform an identical
-principal component analysis
-on the simulated data as was done
-on the experimental data,
-we adopted the methods of the original paper
-and randomly sampled 174 neurons
-with firing rates $> 1$ Hz from
-the spiking double integrator model.
-The spike trains of these neurons
-were convolved with the same Gaussian filter
-as used experimentally
-($\sigma = $ 25 ms)
-and then normalized to Z-scores.
-The results are shown in Figure~\ref{fig6}.
+\subsection{The spiking double integrator has the same principal components as experimental data}
+
+Figure~\ref{fig7} compares
+the two leading principal components
+of the spiking implementation of the double integrator network
+to those of the experimental data.
 The simulated principal components
@@
 \cite{Narayanan2009}:
-$R^2 = 0.96$ and $R^2 = 0.895$ for
-the first two components in the post-correct case,
-$R^2 = 0.979$ and $R^2 = 0.877$ for
-the first two components in the post-premature case,
-and $R^2 = 0.967$ and $R^2 = 0.862$ for
-the first two components in the post-late case.
-$R$ is a Pearson correlation coefficient in all cases.
-
-\subsection{The adaptive control model explains observed performance}
-
-Figure~\ref{fig7} shows the performance
+Pearson $R^2 = 0.82, 0.97, 0.95$
+for the first component
+and $R^2 = 0.91, 0.81, 0.84$
+for the second components in
+the post-correct, post-premature,
+and post-late trials respectively.
+
+\subsection{A single integrator does not have the same principal components as experimental data}
+
+We also performed principal component analysis
+on neural populations representing only one
+of the two dimensions tracked by
+the double integrator network.
+Figure~\ref{fig8}
+shows the results of the analysis
+for the population tracking $x_1$.
+While the first principal component
+closely matches the experimental data
+(Pearson $R^2 = 0.54, 0.86, 0.80$
+for post-correct, post-premature, and post-late
+trials respectively),
+the second principal component
+does not (Pearson $R^2 = 0.15, 0.26, 0.30$).
+This suggests that there exist
+neurons that are sensitive
+to both dimensions of the double integrator.
+
+\subsection{Spiking networks can predict the cue}
+
+Figure~\ref{fig9} shows that
+the decoded values of spiking neural networks
+implementing the cue-responding
+and adaptive control networks
+closely resemble the signals
+directly simulated with
+the dynamical system
+in correct, premature, and late trials.
+Despite the introduction of the oscillation
+and the additional injected noise current,
+the cue can be predicted,
+and the outcome of the previous trial
+has a significant effect
+on the system state during a trial.
+The spiking networks are capable
+of the adaptive control described
+in the theoretical results,
+as can be seen in Figure~\ref{fig9}A.
+
+\subsection{Adaptive control network matches observed performance}
+
+Figure~\ref{fig10} shows the performance
 of the experimental subjects,
-the model using only the cue-responding strategy,
-and the full adaptive model.
-The length of the simulations were
+the spiking cue-responding network,
+and the full spiking adaptive control network.
+The lengths of the simulations were
 varied to match the number of trials
@@
 is 272$\pm$48ms for the experimental subjects,
-368$\pm$34ms for the cue-responding models,
-and 260$\pm$100ms for the adaptive control model.
+365$\pm$24ms for the cue-responding models,
+and 240$\pm$62ms for the adaptive control model.
 The mean performance (\% correct trials) is
 70.8\% for experimental subjects,
-96.9\% for cue-responding models,
-and 84.0\% for adaptive control models.
+98.6\% for cue-responding models,
+and 92.5\% for adaptive control models.
 Therefore, only the adaptive control model
@@
 compared to the experimental subjects.
-The cue-responding model is significantly slower,
-but makes fewer errors than the adaptive control model.
-
-\subsection{Cue prediction causes premature releases}
-
-Figure~\ref{fig8} shows the decoded values
-of neural populations during representative trials
-of the cue-responding and adaptive control circuits.
-In panel C, the double integrator model
-causes the control model to respond faster than
-the cue-responding strategy.
-Premature release never occurs with
-the cue-responding circuit alone.
-While this cannot be directly compared
-with the experimental data,
-as we cannot determine
-whether the experimental subject
-is attempting to
-predict the time of the cue,
-it suggests that more premature
-releases will occur when the time of the cue
-can be predicted;
-if the foreperiod randomly varied, for example,
-our model predicts that fewer premature
-releases would occur.
+The cue-responding network is significantly slower,
+but makes fewer errors than the adaptive control network.
+Both simulated models make fewer errors
+than experimental subjects.
 
@@
 
-Our results indicate that the double integrator model
-allows for adaptive control of action timing.
+Our results show that a spiking implementation
+of the double integrator network
+\cite{Singh2006} can reproduce the dynamics
+of population activity in the rat mPFC
+\cite{Narayanan2009} and can be
+used to modify a feedforward control system
+in order to adaptively control
+the timing of actions.
 The ability to predict the time of the cue
@@
 of additional premature errors.
-The model that enables this adaptive control
-has very similar neural dynamics
-as the experimental ACC.
-
-One important difference between
-the experimentally recorded and simulated ACC
-is in the amount of variance
-explained by the principal components
-and the lack of noise in
-spike densities (see Figure~\ref{fig6}).
-The relative cleanness of the simulated data
-is expected for several reasons.
-Primarily, we expect that the ACC
-is sensitive to many other factors
-than are represented in the simulation.
-The simulated data is more analogous
-to the 174 neurons in the entire ACC
-that are modulated the most during the simple RT task,
-rather than the 174 neurons that happened
-to be recorded during the experiments.
-Increased variability could be simulated
-by either injecting
-noise into the simulation,
-or by representing additional
-randomly varying dimensions.
 
@@
 despite all but one performing the task well
-(see Figure~\ref{fig7}).
-Also interesting is that the model is
-not plastic; the weights determined
+(see Figure~\ref{fig10}).
+Also notable is that the model
+does not employ learning:
+the weights determined
 before the simulation do not change.
@@
 is what enables the model to adapt.
-This points to a need for models to
-differentiate between adaptability
-due to recurrent activity
-and adaptability due to
-synaptic weight changes.
-
-While we have tested our mathematical model
-of ACC function in the context of
-control of action timing in a simple RT task,
-the tracking of behavioral variables
-in a dynamical system could be used
-to perform many of the other high-level functions
-that the ACC is theorized to perform.
-For example, in the commonly studied Stroop task,
-the state space being tracked by
-the ACC could be the difference between
-the response suggested by word-reading and the
-response suggested by the color;
-the context in which colors should be read
-would be reflected in the passive dynamics
-of the ACC, which would push the ACC
-to increasingly drift towards a state
-that compensates for the word-reading response.
-Similar mappings could be proposed
-for a wide variety of ACC-monitored tasks.
-
-In conclusion, we have shown that neural integration
-is likely a key mechanism in the ACC's
-ability to control the timing of actions.
-Analysis of the activity of a neural integrator
-model closely matches that of the experimentally recorded ACC.
-That model is sufficient to allow
-a simple cue-responding network
-to react more quickly at the cost
-of additional errors,
-which matches the performance of experimental subjects.
+By defining the dynamics of other tasks
+and implementing them
+in recurrently connected neural networks,
+we can start to form general theories
+of how biological systems can learn
+to perform complex tasks.
+
+\subsection{Relation to other models}
+
+In this study, we make
+an adaptive control hypothesis
+of mPFC function in reaction time tasks.
+This hypothesis is compatible
+with previous hypotheses suggesting
+that the mPFC plays a role
+in the regulation of
+reinforcement learning parameters.
+One such example is the PRO
+model by \citetext{Alexander2011}
+(though these arguments also
+apply to related models,
+such as \citenoparens{Khamassi2012}).
+The PRO model proposes that the mPFC
+is involved in a form of
+temporal difference learning.
+As shown in Figure~\ref{fig6},
+the positive and negative reward prediction error
+computed by PRO can also be computed
+by the double integrator,
+assuming that the value function
+learned by the system steadily increases.
+
+In our model, we interpret the steadily increasing
+signal as representing time elapsed rather
+than value over time;
+this means that there is only one parameter
+that must be learned
+($\beta$, the speed at which $x_2$ increases),
+rather than learning the value at every timestep
+in the simulation.
+In addition to being an easier target for learning,
+we further suggest that time tracking
+is a more natural interpretation for
+such a ramping signal.
+Theoretically, it is more likely that
+the value signal associated with those
+prediction errors is constant
+over time rather than steadily increasing
+\cite{Niv2008}.
+We suggest that the ramping signal
+shown in \citetext{Alexander2011} might
+reflect the temporal dynamics of the task
+instead of serving as a value signal.
+
+The model presented in this study keeps track
+of the results of the previous trial,
+and therefore can only learn
+on a trial-by-trial basis.
+Some quantities like volatility,
+which is thought to be encoded
+in the mPFC \cite{Behrens2007},
+would require additional integrative
+populations to aggregate information
+across many trials.
+While we have not presented
+such a model in this study,
+we believe that a similar dynamical approach
+can be used to construct models
+that encode volatility
+and other quantities
+in spiking neural networks,
+which can be directly compared
+to neural recordings.
+
+\subsection{Adaptation and learning}
+
+A primary difference between
+the adaptive control network we have presented
+and previous hypotheses of mPFC function
+is that we do not employ learning
+to change connection weights
+during a simulation.
+Instead, we achieve adaptive behavior
+through recurrently connected
+integrative populations that
+maintain information across trials.
+We believe that this type of adaptability
+should be explored in more depth in mPFC models.
+
+We are able to construct a spiking neural network
+that performs the simple RT task
+because we have made a hypothesis of the
+dynamics governing the task.
+One issue that we have not attempted to address
+in this work is how those dynamics
+could be learned from the sensory
+information available during the task.
+We believe that this learning procedure
+is central to understanding the mPFC,
+and the brain in general.
+However, in order to characterize
+this learning procedure,
+we must first have examples
+of the endpoints of the learning procedure.
+We believe that
+the adaptive control network,
+which uses the double integrator
+to predict cue times,
+is an example of one possible endpoint.
+By understanding the dynamics
+of other tasks and how those dynamics
+might be implemented in spiking neural networks,
+we can begin to form theories of
+how spiking neural networks
+performing complex tasks might be learned.
+In previous work, we have demonstrated
+that a spike-based learning rule
+can generate a neural integrator
+like those used in this study \cite{Macneil2011},
+but learning the entire network
+described here is the subject of future work.
+
+\subsection{Limitations}
+
+Our model is specific to the simple reaction time task
+being performed by the experimental subjects,
+and therefore can only be directly
+applied to experimental data
+of subjects performing this task.
+However, we believe that the
+methods used to determine the
+dynamical system governing the task
+and to implement the dynamical system
+in a spiking neural network
+are general,
+and can be used to explain many other tasks.
+Determining the dynamics of other tasks
+in a dynamical systems framework
+can help identify general principles
+of neural function.
+
+Even in the context of the simple reaction time task,
+our model cannot be directly compared
+to all of the experimental data available.
+We have chosen to implement the model
+in a spiking neural network,
+as we believe that it provides the most
+direct comparison to
+the data recorded from the biological system.
+However, because the mappings from
+spiking activity to EEG and fMRI recordings
+are not well understood,
+we cannot directly explain
+the wealth of data from humans performing
+similar reaction time tasks.
+Previous studies of mPFC function
+in humans have found evidence
+for low-frequency rhythmic activity being
+generated when mistakes are made
+and on trials after such errors
+\cite{Cavanagh2009}.
+A recent study reported the same types
+of error-encoding cortical rhythms
+in rat mPFC \cite{Narayanan2013}
+and showed that mPFC activity constrains
+rhythmic activity in motor cortex.
+It has been suggested that the expression
+of these rhythms reflects alterations
+in cross-laminar processing
+\cite{Carracedo2013}.
+These effects could be examined by
+using specific models for different
+cell types and laminar patterns
+for connecting neurons together.
+
+\subsection{Predictions}
+
+Unlike previous mPFC models,
+we predict that mPFC inactivation
+negatively impacts behavioral performance
+of tasks that require precisely timed actions,
+in addition to disrupting
+the learning of such tasks.
+
+At the neural level,
+the spiking implementation
+of the network predicts that
+neurons in the mPFC will ramp up in activity
+over the course of a goal-directed task.
+However, we predict that those ramping signals
+are an attempt to encode
+the temporal dynamics of the task,
+not an attempt to encode action-outcome values.
+This prediction could be tested
+by recording neural ensembles
+in the mPFC and motor cortex during learning.
+One important study would be to record
+the mPFC of naive rats,
+as was done in the motor cortex
+in \citetext{Laubach2000}.
+In that earlier study, a major limitation
+was that the rats only experienced
+brief foreperiods (400--600ms) before
+the trigger stimulus, and ramping activity
+was not apparent in the PCA-style analyses
+(unpublished observations).
+A better test of this idea would be to extend
+the required timing of the action to 1 second
+and record in the two cortical areas simultaneously.
+Neural integration could develop with the increased
+length of the foreperiod,
+and the rate of integrator ``ramping''
+might track the animals' prediction of the cue time.
+Alternatively, recordings could be made
+in well-trained rats who experience
+two foreperiods presented in blocks.
+We predict that the rate of integration
+would change following block changes,
+e.g. as the foreperiod switches from
+relatively long to short.
+To our knowledge, no studies like these
+have been done.
+
+We have shown that neurons
+in mPFC are tuned to multiple aspects
+of the simple RT task
+(see Figures~\ref{fig7} and \ref{fig8}).
+Since we find neurons that are sensitive
+to both task state ($x_1$) and the
+length of time in that state ($x_2$),
+we predict that single neurons
+would be sensitive to additional
+aspects of the task in more complex tasks.
+This is due to the mPFC as a whole
+tracking the dynamics of that task
+in order to modify behavior.
+Initial support for this idea
+can be found in a recent study
+of the mPFC that used
+a delayed spatial alternation task
+and found similar dynamics
+in principal component space across
+the mPFC ensembles \cite{Horst2012}.
+This prediction could be further tested
+by using a task that
+has different temporal dynamics
+in different trial blocks or task contexts,
+and determining if there are neurons
+with loadings on many primary components
+in a principal component analysis.
+One such task was used by
+\citetext{Delatour2001}.
 
@@
 depending on the time of lever release.
+The times at which important events occur
+in the trials are labeled.
+In order, $t_s$ is the start of the trial,
+$t_p$ is when the lever is fully pressed,
+$t_c$ is the time of the auditory cue,
+$t_{Rw}$ is when reward is delivered,
+$t_{TO}$ is when house lights are extinguished,
+and $t_{ITI}$ is the start of the intertrial interval.
+(A) Schematic of each trial type.
+(B) Results of simulating each trial type
+according to Equation~\eqref{eq:env}.
+(C) Finite state machine tracking
+task state in the simulation.
 
@@
 A summary of the principal component analysis done
-in \citetext{Narayanan2009};
-the motivation for investigating \citetext{Singh2006}.
+in \citetext{Narayanan2009}.
 (A) A graphical summary of principal component analysis.
-Nomarlized perievent neural data,
+Normalized perievent neural data,
 in the form of the Z-scores of instantaneous firing rates,
@@
 to the first and second principal components.
-(D) The top three principal components
+(D) The top two principal components
 for the post-correct trials (black)
@@
 This points to neural integration
-as a potential mechanism explaining ACC activity.
+as a potential mechanism explaining mPFC activity.
 (F) Normalized spike-density functions
@@
 
-Block diagram of the adaptive control model.
+Block diagram of the adaptive control network.
 (Left) The adaptive control model contains
-the double integrator model,
+the double integrator network
 from \citetext{Singh2006}.
@@
 both press information ($u$)
-and error information ($E$)
+and error information ($E \; u_{TO}$)
 to drive $x_1$, and
-reward information ($R$)
+reward information ($R \; u_{Rw}$)
 to control the integration
 of both $x_1$ and $x_2$.
-The output of the double integrator,
-colored red, is the only component
-that interacts with the cue-responding
-circuit, by activating neurons
-in the \textit{Trigger} population.
-(Middle) The cue-responding circuit
-that is able to perform
+Lines terminating with circles indicate
+negative connections;
+lines terminating with arrowheads indicate
+positive connections.
+The output of the double integrator
+is the only component
+that interacts with the cue-responding network.
+(Middle) The cue-responding network
+is able to perform
 the simple reaction time task
-by responding to the Cue
+by responding to the Cue ($u_C$)
 provided by the environment
-(see section \ref{sec:cue-responding}).
+(see the text, Section \ref{sec:cue-responding}).
 (Right) Values that are not modeled
@@
 The lever accepts the decoded
-output of the \textit{Press/Release} population,
-lowering after \textit{Press} sensitive
-neurons are activated (following some delay),
-and raising after \textit{Release} sensitive
-neurons are activated (following the same delay).
+output of the cue-responding network.
 
@@
 
-A detailed breakdown of the dynamics and
-trajectory the model takes through state space
-for (A) correct, (B) premature, and (C) late trials
-following a correct trial.
-The injected oscillation is omitted
-in these plots;
-its time dependence makes it misleading
-to draw vector fields for a particular time.
-(A) In a correct trial, the pressing of the lever
-sends the double integrator model's state into
-the upper part of the state space.
-The connection between the two integrators
-causes the passive dynamics (panel 2)
-to push the state towards the right.
-The cue is delivered while drifting
-to the right, and the release occurs
-in the upper-right portion of the state space.
-After successful release,
-the reward turns off both integrators,
-causing the system to return to the origin point,
-where it remains until the next trial.
-(B) In a premature trial, the lever is released
-prior to the onset of the cue (panel 2).
-Premature release causes
-the house lights to extinguish,
-which drives the system to
-a point in the lower part of the state space.
-Between trials (panel 4),
-the passive dynamics of the state space
-drive the system to the lower-left portion
-of the space, where it will be
-at the start of the next trial.
-(C) In a late trial, the cue occurs,
-but the lever is not released (panel 2).
-Late release causes
-the house lights to extinguish,
-which drives the system
-to a point in the lower part of the state space.
-Between trials (panel 4), the passive dynamics
-of the state space drive the system
-to the same lower-left portion of the space
-that premature errors result in.
-Therefore, late and premature errors
-cannot be distinguished on the next trial.
+Time tracking in the double integrator model.
+$x_2$ (solid black line) ramps up over time,
+enabling the network to predict
+the time of the cue.
+Time labels indicate important events during each trial.
+$t_s$ is the start of the trial,
+$t_p$ is when the lever is fully pressed,
+$t_c$ is the time of the auditory cue,
+$t_{Rw}$ is when reward is delivered,
+$t_{TO}$ is when house lights are extinguished,
+and $t_{ITI}$ is the start of the intertrial interval.
+(A) The state of the system at $t_c$
+depends on the initial state of the system
+at the start of the trial, $t_p$.
+The initial state depends on the outcome
+of the previous trial;
+after a correct trial
+the system should start at $(0, 0)$,
+and after a premature or late trial,
+the system should start at $(-1, -1)$.
+(B) The slope of the ramping signal $x_2$
+depends on the parameter $\beta$, which
+represents the strength of the connection
+between $x_1$ and $x_2$.
+An ideal $\beta$ pushes $x_2$ far from
+the initial state,
+but within the representational range
+of a population of spiking neurons,
+which is normalized to $(-1, 1)$;
+therefore, the system should approach $(1, 1)$
+at the time of the cue,
+which occurs when $\beta \approx 0.44$.
+(C) The system is reset to $(0, 0)$
+after a correct trial by reward feedback.
+The parameter $R$ controls
+the strength of the reward feedback;
+the system is properly reset
+if $R \ge 2$.
+(D) The system is driven near $(-1, -1)$
+after an error trial by error feedback.
+The parameter $E$ controls
+the strength of the error feedback;
+the system is driven near $(-1, -1)$
+if $E \ge 1$.
 
@@
 
-Decoded values of the double integrator model
-simulated in spiking neurons.
-The spiking neural model takes
-very similar trajectories through state
-space as the mathematical model (see Figure~\ref{fig4}).
-Similar to the mathematical model,
-the spiking model represents a very different state
-after correct (left) and error (middle, right)
-trials.
+Results of control network simulations.
+Time labels indicate important events during each trial.
+$t_s$ is the start of the trial,
+$t_p$ is when the lever is fully pressed,
+$t_c$ is the time of the auditory cue,
+$t_{Rw}$ is when reward is delivered,
+and $t_{TO}$ is when house lights are extinguished.
+(A) The cue-responding network.
+The release, $u_r$ is initiated only after cue onset, $t_c$.
+(B) The adaptive control network in the ideal case.
+The release is initiated such that
+the lever (dashed black line) is raised
+at the time of cue onset.
+(C) A premature release in the adaptive control network.
+The release is initiated before the time of cue onset
+(premature release is emulated by setting $\beta = 0.94$).
+(D) The adaptive control network follows
+the cue-responding strategy after error trials,
+due to the initial state of the double integrator.
 
 \paragraph{Figure 6}
+
+(Left) Signals tracked by the PRO model \cite{Alexander2011}.
+(Right) Signals tracked by or computed from
+the double integrator network.
+In all but one case, the signals are similar,
+and differ mainly in task-related properties
+(e.g., the length and magnitude of the reward signal).
+In both models, all signals can be represented
+as vectors if appropriate for the task.
+The one divergent case is the state;
+in the PRO model, the state signal
+persists after the reward has been delivered,
+but in the double integrator network
+is turned off by the reward.
+
+\paragraph{Figure 7}
 
@@
 on the data generated by the double integrator model
-for correct presses in (A) post-correct trails,
+for correct presses in (A) post-correct trials,
 (B) post-premature trials, and (C) post-late trials.
@@
 the top two principal components
-of the model to those of the experimental ACC data.
+of the model to those of the experimental mPFC data.
 In all cases, the PCs are significantly similar
-(Pearson $R^2 > 0.86$).
+(Pearson $R^2 > 0.8$).
 The middle row of panels shows normalized
@@
 
-\paragraph{Figure 7}
-
-Behavioural performance of the
+\paragraph{Figure 8}
+
+Results of principal component analysis
+on the data generated by the integrator
+tracking $x_1$ in the double integrator model
+for correct presses in (A) post-correct trials,
+(B) post-premature trials, and (C) post-late trials.
+The top row of panels compares
+the top two principal components
+of the model to those of the experimental mPFC data.
+In all cases, the first PC is significantly similar
+(Pearson $R^2 > 0.5$),
+but the second PC is not (Pearson $R^2 \le 0.3$).
+The bottom row of panels shows normalized
+peri-event spike densities
+for 174 randomly sampled neurons from the
+simulated integrator population.
+
+\paragraph{Figure 9}
+
+(A) Decoded values of representative trials of
+the control networks simulated with spiking neurons.
+From top to bottom, the trials shown are
+a correct trial in the cue-responding network,
+a late trial in the cue-responding network,
+a correct trial in the adaptive control network,
+and a premature trial in the adaptive control network.
+(B) The decoded value of the double integrator network
+that is embedded in the adaptive control network
+on a correct trial following a correct trial.
+The system starts ($t_s$) and finishes ($t_{ITI}$)
+the trial near the origin point.
+(C) The decoded value of the double integrator network
+on a correct trial following an error trial.
+The system starts near the lower left
+portion of the state space,
+but since the trial was correct,
+finishes near the origin point.
+
+\paragraph{Figure 10}
+
+Behavioral performance of the
 experimental subjects and
@@
 The dotted black line represents
-the mean of all subjects' median reaction times.
-
-\paragraph{Figure 8}
-
-Decoded values from representative trials.
-The labels above denote events;
-`P' is a lever press event,
-`C' is a cue onset event,
-`R' is a lever release event,
-`E' is an error delivery event,
-and `Rw' is a reward delivery event.
-Names of the decoded signals correspond
-to the names of populations in Figure~\ref{fig3}.
-(A) A correct trial using the cue-responding circuit.
-In the cue-responding network, the trigger neurons
-activate only once the cue is delivered.
-Trigger neurons activate the release neurons,
-which initiate the release of the lever.
-There is a delay ($\sim 250ms$)
-between the activation of release neurons
-and the actual release of the lever.
-(B) A correct trial using the adaptive control circuit.
-In the adaptive network, when
-in the appropriate state, the cue
-can be predicted by the double integrator model,
-causing the activation of trigger---and
-therefore release---neurons before
-the time of the cue. This results
-in a faster reaction to the cue.
-(C) A premature trial using the adaptive control circuit.
-Premature release occurs when
-the double integrator model incorrectly predicts
-the time of the cue.
-This occurs when integration is too fast,
-and the trigger neurons are activated
-such that a lever release is effected
-before the time of the cue.
-Note that the non-adaptive cue-responding
-circuit cannot make premature releases,
-because releases are critically dependent
-on cue information.
-(D) A late trial using the cue-responding circuit.
-Late release occurs when
-the release neurons are not sufficiently
-active in order to effect the lever release.
-Alternatively, late release can occur
-if the release signal is not faithfully
-transmitted from the release neurons
-(modeling activity in motor cortex)
-to the intermediate populations
-(modeling activity between motor cortex and muscles)
-before interfacing with the simulated environment.
-A late release in the adaptive model
-looks identical, with little activity
-in the double integrator model that predicts cues.
-
+the mean of all subjects' median reaction times
+(272$\pm$48ms for the experimental subjects,
+365$\pm$24ms for the cue-responding models,
+and 240$\pm$62ms for the adaptive control models).
+
+\clearpage
 \section{Figures}
 
-\scalefigone{fig1}{0.5}{}
+\scalefigone{fig1}{1.0}{}
 
@@
 
-\scalefigone{fig6}{1.0}{}
+\scalefigone{fig6}{0.5}{}
 
@@
 
+\scalefigone{fig9}{0.5}{}
+
+\scalefigone{fig10}{1.0}{}
+
 \end{document}

### 2013-11-14 dodo.py

@@ -1,0 +1,155 @@
+import inspect
+import os.path
+
+from doit.action import CmdAction
+
+from scripts import base, expt, files, sim_io
+from scripts import ctrl, ctrl_plot, dint, dint_plot, ds, ds_plot
+
+
+DOIT_CONFIG = {'default_tasks': ['convert'], 'verbosity': 2}
+
+
+def task_setup_nengo():
+    """Make sure scripts will run from Nengo."""
+    actions = []
+    if not os.path.exists(files.nengo):
+        actions.append('mv ' + files.home + '/nengo-* ' + files.nengo)
+    if not os.path.exists(files.nengoscripts):
+        actions.append('ln -s ' + files.jneuro + ' ' + files.nengoscripts)
+    return {'actions': actions, 'targets': [files.nengo, files.nengoscripts]}
+
+
+def task_sim_dint():
+    """Run a double-integrator simulation."""
+    return {'actions': [files.nengo_cl + ' ' + files.nef_dint
+                        + ' %(nperd)s %(seed)s'],
+            'params': [{'name': 'nperd', 'short': 'n', 'default': '600'},
+                       {'name': 'seed', 'short': 's', 'default': '933'}]}
+
+
+def task_sim_ctrl():
+    """Run a control circuit simulation."""
+    return {'actions': [files.nengo_cl + ' ' + files.nef_ctrl
+                        + ' %(control)s %(length)s %(nperd)s %(seed)s'],
+            'params': [{'name': 'control', 'short': 'c', 'default': 'adaptive'},
+                       {'name': 'length', 'short': 'l', 'default': '200'},
+                       {'name': 'nperd', 'short': 'n', 'default': '600'},
+                       {'name': 'seed', 'short': 's', 'default': '933'}]}
+
+
+def task_download_data():
+    """Downloads analyzed data from Figshare."""
+    def download_data(article_id='715887'):
+        import json
+        import requests
+        r = requests.get("http://api.figshare.com/v1/articles/" + article_id)
+        detail = json.loads(r.content)
+        for file_info in detail['items'][0]['files']:
+            outpath = os.path.join(files.simdata, file_info['name'])
+            if os.path.exists(outpath):
+                print outpath + " exists. Skipping."
+                continue
[…]

### 2013-11-14 run.py

@@ -1,302 +1,0 @@
-from glob import glob
-import os
-import os.path
-import subprocess
-import sys
-
-import numpy as np
-import matplotlib
-matplotlib.use('Agg')
-font = {'family': 'sans-serif',
-        'sans-serif': 'Arial'}
-matplotlib.rc('font', **font)
-matplotlib.rc('figure', dpi=100)
-from IPython import parallel
-
-from scripts.analyze import aggregate_ctrl, aggregate_dint
-from scripts.analyze import process_exp, process_sim_ctrl, process_sim_dint
-import scripts.combine as c
-import scripts.plot as p
-from scripts.sim_io import convert_to_h5
-
-unbuffered = os.fdopen(sys.stdout.fileno(), 'w', 0)
-#unbuffered = sys.stdout
-
-
-def download_data():
-    import json
-    try:
-        import requests
-    except:
-        print ("Requires requests. Please apt-get install python-requests,"
-               " or pip install requests.")
-        return
-
-    outdir = 'data_sim/'
-    article_id = '715887'
-
-    r = requests.get("http://api.figshare.com/v1/articles/" + article_id)
-    detail = json.loads(r.content)
-    for file_info in detail['items'][0]['files']:
-        outpath = os.path.join(outdir, file_info['name'])
-        if os.path.exists(outpath):
-            print outpath + " exists. Skipping."
-            continue
-        with open(outpath, 'wb') as outf:
-            print "Downloading " + outpath + "..."
-            dl = requests.get(file_info['download_url'])
-            outf.write(dl.content)
-
-def schedule(lview, func, path, *args, **kwargs):
-    if isinstance(path, list):
-        for pth in path:
-            if not os.path.exists(pth):
-                break
-        else:
-            return
-    elif isinstance(path, str):
-        if os.path.exists(path):
-            return
-
-    if isinstance(path, list):
-        res = lview.map(func, *args, **kwargs)
-    else:
-        res = lview.apply(func, *args, **kwargs)
-    return res
-
-
-def wait_for(jobs, name="Jobs"):
-    if len(jobs) == 0:
-        return
-    unbuffered.write("%s scheduled. Waiting for results..." % name)
-
-    for job in jobs:
-        job.wait()
-    unbuffered.write("done!\n")
-
-    # Any failures?
-    for job in jobs:
-        if not job.successful():
-            unbuffered.write("Failed. Printing first failure:\n")
[…]

### 2013-11-14 scripts/NEFBuilder.py

@@ -19,7 +19,7 @@ class BaseReactionTask(nef.FSMNode):
     The subject drinks for `t_drink` seconds.
     '''
     def __init__(self, name, t_foreperiod=1.0, t_timeout=2.0, t_response=0.6,
-                 t_tone=0.1, t_reward=1.0, t_iti=5.0):
+                 t_tone=0.1, t_reward=2.0, t_iti=5.0):
         self.reset()
         nef.FSMNode.__init__(self, name, 1)
 

### 2013-11-14 scripts/NEFControl.py

@@ -10,9 +10,9 @@ from NEFDoubleIntegrator import DoubleIntegrator
 class TopDownReactionTask(BaseReactionTask):
     def termination_lever(self, x, pstc=0.01, dimensions=2):
         if x[0] > 0.85:
-            self.lever[0] -= x[0] * 0.2  # dim0 = press
+            self.lever[0] -= x[0] * 0.19  # dim0 = press
         if x[1] > 0.75:
-            self.lever[0] += x[1] * 0.2  # dim1 = lift
+            self.lever[0] += x[1] * 0.19  # dim1 = lift
         self.lever[0] = min(1.0, self.lever[0])
         self.lever[0] = max(-1.0, self.lever[0])
 
@@ -27,10 +27,6 @@ class AdaptiveControl(DoubleIntegrator):
     def simple_make(self, net):
         net.make('Press/Release', self.nperd * 2, 2,
                  radius=self.radius, node_factory=self.alif_factory)
-        net.make('Holding', self.nperd * self.intscale * 2, 2,
-                 radius=self.radius, node_factory=self.alif_factory)
-        net.make('Trigger', self.nperd * 2, 2, radius=self.radius,
-                 node_factory=self.alif_factory)
         [net.make('P/R Delay %d' % i, self.nperd * 2, 2,
                   node_factory=self.alif_factory, radius=self.radius)
          for i in xrange(self.delaypops)]
@@ -40,9 +36,7 @@ class AdaptiveControl(DoubleIntegrator):
                     pstc=0.1, transform=[1, 0])
         net.connect('P/R Delay %d' % (self.delaypops - 1),
                     net.get('RTTask').getTermination('lever'))
-        net.connect('RTTask', 'Holding', origin_name='reward',
-                    pstc=0.1, transform=[0, -1])
-        net.connect('RTTask', 'Trigger', origin_name='trigger',
+        net.connect('RTTask', 'Press/Release', origin_name='trigger',
                     pstc=0.1, transform=[0, 3])
 
         net.connect('Press/Release', 'P/R Delay 0', pstc=0.01)
@@ -50,24 +44,14 @@ class AdaptiveControl(DoubleIntegrator):
         for i in xrange(self.delaypops - 1):
             net.connect('P/R Delay %d' % i, 'P/R Delay %d' % (i + 1),
                         pstc=0.05)
-        net.connect('Press/Release', 'Holding',
-                    pstc=0.01, transform=[[0.25, 0], [0, 0]])
[…]

### 2013-11-14 scripts/NEFDoubleIntegrator.py

@@ -127,48 +127,48 @@ class DIntReactionTask(BaseReactionTask):
 
 
 class DoubleIntegrator(MPFCModel):
-    def __init__(self, nperd=200,  mode=SimulationMode.DEFAULT, seed=None):
+    def __init__(self, nperd=200, seed=None):
         self.nperd = nperd
-        self.mode = mode
-
         self.radius = 1.1
         self.intscale = 2
 
         super(DoubleIntegrator, self).__init__(seed)
 
     @staticmethod
-    def oscillation(t, amp=1.0, freq=0.185, phase=-2.3):
+    def oscillation(t, amp=1.0, freq=0.185, phase=-2.28):
         return amp * math.sin(2 * math.pi * freq * t + phase)
 
     def oscillator_make(self, net):
         net.make_input('Oscillation', DoubleIntegrator.oscillation)
 
     def dint_make(self, net):
-        net.make('Delay state', self.nperd * 2, 2, radius=self.radius,
-                 node_factory=self.alif_factory)
-        net.make('Delaying', self.nperd * 2 * self.intscale, 2, noise=0.1,
-                 radius=self.radius, node_factory=self.alif_factory)
-        net.make('Timer', self.nperd * 2 * self.intscale, 2, noise=0.1,
-                 node_factory=self.alif_factory, radius=self.radius)
+        common = {'dimensions': 2,
+                  'node_factory': self.alif_factory,
+                  'radius': self.radius}
+
+        n = self.nperd * 2
+
+        net.make('Delay state', n, noise=1.5, **common)
+        net.make('Delaying', n * self.intscale, noise=0.15, **common)
+        net.make('Timer', n * self.intscale, **common)
 
     def _dint_connect(self, net):
         net.connect('RTTask', 'Delaying', origin_name='press',
-                    transform=[0.14, 0], pstc=0.01)
+                    transform=[0.16, 0], pstc=0.01)
         net.connect('RTTask', 'Timer', origin_name='press',
                     transform=[0, -0.01], pstc=0.01)
 
     def dint_connect(self, net):
         net.connect('RTTask', 'Delaying', origin_name='reward',
-                    transform=[-0.005, -0.2], pstc=0.01)
+                    transform=[-0.01, -0.2], pstc=0.1)
         net.connect('RTTask', 'Timer', origin_name='reward',
-                    transform=[0, -0.15], pstc=0.01)
[…]

### 2013-11-14 scripts/analyze.py

@@ -1,529 +1,0 @@
-from __future__ import absolute_import
-
-import os.path
-import random
-
-import numpy as np
-import quantities as pq
-from scipy import linalg
-from scipy import signal, stats
-from scipy.io import loadmat
-
-from .sim_io import get_seg
-from . import neo
-
-def correct_seq(e):
-    return (e['press'], e['toneon'], e['release'], e['pumpon'])
-
-
-def premature_seq(e):
-    return (e['press'], e['release'], e['lightsoff'])
-
-
-def late_seq(e):
-    return (e['press'], e['toneon'], e['lightsoff'])
-
-
-def reaction_seq(e):
-    return (e['toneon'], e['release'])
-
-
-def delay_seq(e):
-    return (e['press'], e['toneon'])
-
-
-def pca(x, comps=2, flip=True):
-    evecs, evals, _ = linalg.svd(x.T, full_matrices=False)
-    evecs = stats.zscore(evecs[:, :comps], ddof=1)
-    evals = (evals ** 2) / np.sum(evals ** 2)
-    loadings = np.dot(x, evecs)
-
-    if not flip:
-        return {'evecs': evecs, 'evals': evals, 'loadings': loadings}
-
-    for i in xrange(comps):
-        if evecs.T[i][int(evecs.T[i].shape[0] * 0.485)] < 0:
-            evecs.T[i] *= -1
-            loadings.T[i] *= -1
-
-    return {'evecs': evecs, 'evals': evals, 'loadings': loadings}
-
-
-def integrator_vf(x1, x2, press=0, reward=0, error=0, degrade=0):
-    alpha = 0.2
-    x1, x2 = x1.T, x2.T
-    dx1 = 0.45 * press + 0.25 * error - reward * x1 - degrade * x1
-    dx2 = alpha * x1 - 0.75 * reward * x2 - degrade * x2
-    return dx2, dx1
-
-
-def gaussian_sdf(hist, binsize, sigma=0.015 * pq.s, denoise=True):
-    # Put gaussian in terms of bins
-    sigma.units = binsize.units
-    sigma = float(sigma / binsize)
-    gauss = signal.gaussian(int(sigma * 7), sigma)  # +/- 3 stddevs
-    gauss /= np.sum(gauss)  # normalize to area 1 (important!)
-    gauss /= binsize  # then scale to size of bins
-    sdf = np.convolve(hist, gauss, mode='same')
-
[…]

### 2013-11-14 scripts/base.py

@@ -1,0 +1,185 @@
+try:
+    import cPickle as pickle
+except ImportError:
+    import pickle
+import subprocess
+
+import numpy as np
+import quantities as pq
+from scipy import linalg, stats
+import svgutils.transform as sg
+
+from . import files, neo
+
+
+### Pickle
+
+def load_pickle(pklf):
+    with open(pklf, 'rb') as pklpkl:
+        ret = pickle.load(pklpkl)
+    return ret
+
+
+def strip_dict(d):
+    for k in d:
+        d[k] = stripped(d[k])
+
+
+def strip_iterable(it):
+    for ix in xrange(len(it)):
+        it[ix] = stripped(it[ix])
+
+
+def stripped(el):
+    if type(el) == dict:
+        strip_dict(el)
+    elif type(el) == list:
+        strip_iterable(el)
+    elif hasattr(el, 'units'):
+        # print 'stripping',type(el)
+        return el.magnitude
+    # elif type(el) == tuple:
+    #     print 'tuple'
+    # else:
+    #     print 'not stripped',type(el)
+    return el
+
+
+def save_pickle(d, pklfp):
+    strip_dict(d)
+    with open(pklfp, 'wb') as pklf:
+        pickle.dump(d, pklf, protocol=pickle.HIGHEST_PROTOCOL)
+
+
+### Analysis
+
+def pca(x, comps=2):
+    evecs, evals, _ = linalg.svd(x.T, full_matrices=False)
+    evecs = stats.zscore(evecs[:, :comps], ddof=1)
+    evals = (evals ** 2) / np.sum(evals ** 2)
+    loadings = np.dot(x, evecs)
+    return {'evecs': evecs, 'evals': evals, 'loadings': loadings}
+
+
+def behaviour(seg, e):
+    merged = merge_all_eas(seg, e.values())
+
+    # How many correct, premature, late?
+    cor = merged.epochs_from(correct_seq(e), "C")
+    pre = merged.epochs_from(premature_seq(e), "P")
+    lat = merged.epochs_from(late_seq(e), "L")
+
+    # Filter for only long-delay correct trials
+    merged = merged.filter_for(correct_seq(e))
+    dly = merged.epochs_from(delay_seq(e))
+    react = merged.epochs_from(reaction_seq(e))
+    if e['toneon'] == 'eventTONEOFF':
+        react.times -= 0.1 * pq.s
+        react.durations += 0.1 * pq.s
[…]

### 2013-11-14 scripts/combine.py

@@ -1,149 +1,0 @@
-from glob import glob
-import os.path
-import random
-import subprocess
-
-import svgutils.transform as sg
-
-outdir = 'figures'
-sc = 1.25
-dims = {
-    'pc': {'w': 405 / sc, 'h': 540 / sc},
-    'traj': {'w': 360 / sc, 'h': 360 / sc},
-    'perf': {'w': 360 / sc, 'h': 360 / sc},
-    'rts': {'w': 360 / sc, 'h': 360 / sc},
-    'signals': {'w': 180 / sc, 'h': 450 / sc},
-}
-
-
-def el(char, path, x, y, w=None, h=None, scale=1, offset=(10, 30)):
-    toret = []
-    if char is not None:
-        toret.append(sg.TextElement(x + offset[0], y + offset[1],
-                        char, size=24, weight='bold', font='Arial'))
-    if path.endswith(".svg"):
-        svg = sg.fromfile(path)
-        if w is not None and h is not None:
-            svg.set_size((w, h))
-        svg = svg.getroot()
-        svg.moveto(str(x), str(y), scale)
-        return [svg] + toret
-
-
-def svgfig(w, h):
-    w = str(w)
-    h = str(h)
-    return sg.SVGFigure(w, h)
-
-
-def savefig(fig, out):
-    fig.save('%s.svg' % out)
-    print "Saved %s.svg" % out
-    subprocess.call(['inkscape',
-                     '--export-pdf=%s.pdf' % out, '%s.svg' % out])
-    subprocess.call(['inkscape', '--export-text-to-path',
-                     '--export-eps=%s.eps' % out, '%s.svg' % out])
-
-
-def fig4():
-    print 'fig4'
-    out = '%s/fig4' % outdir
-
-    cc = sorted(glob('plots/cc*-?_traj.svg'))
-    pc = sorted(glob('plots/pc*-?_traj.svg'))
-    lc = sorted(glob('plots/lc*-?_traj.svg'))
-
-    fig = svgfig(dims['traj']['w'] * 4, dims['traj']['h'] * 3)
-
-    l = 'A'
-    for i, c_f in enumerate(cc):
-        fig.append(el(l, c_f, dims['traj']['w'] * i, 0))
-        l = None
-
-    l = 'B'
-    for i, p_f in enumerate(pc):
-        fig.append(el(l, p_f, dims['traj']['w'] * i, dims['traj']['h']))
[…]

### 2013-11-14 scripts/ctrl.py

@@ -1,0 +1,119 @@
+from __future__ import absolute_import, division
+
+import numpy as np
+import quantities as pq
+
+from .sim_io import get_seg
+from . import base, files
+
+
+def perievent_signal(analog, evt, window):
+    t = analog.times
+    ix = np.logical_and(t > (evt + window[0]), t <= (evt + window[1]))
+    return [analog[ix].times, analog[ix]]
+
+
+def decoded_signals(seg, e, trial='C', with_x=True, window=(-1.0, 3.0) * pq.s):
+    evts = base.merge_all_eas(seg, e.values())
+    press_evts = base.valid_presses(seg, e, (trial,))
+
+    # Choose the 5th press event, if more than 5
+    ix = 4 if len(press_evts.times) >= 5 else -1
+    press = press_evts.times[ix]
+
+    if trial == 'C':
+        seq = base.correct_seq(e)
+    elif trial == 'P':
+        seq = base.premature_seq(e)
+    elif trial == 'L':
+        seq = base.late_seq(e)
+    s_evts = [evts[evts.times >= press].filter_for((evt,)).times[0]
+              for evt in seq]
+    s_evts = [-window[0] + (evt - press) for evt in s_evts]
+
+    evt_t = [-window[0]]
+    evt_l = ['$t_p$']
+
+    if trial == 'C' or trial == 'L':
+        evt_t.append(s_evts[1])
+        evt_l.append('$t_c$')
+    elif trial == 'P':
+        evt_t.append(s_evts[1])
+        evt_l.append('$t_r$')
+
+    if trial == 'L' or trial == 'P':
+        evt_t.append(s_evts[2])
+        evt_l.append('$t_{TO}$')
+    elif trial == 'C':
+        evt_t.extend(s_evts[2:4])
+        evt_l.extend(['$t_r$', '$t_{Rw}$'])
+
+    signals = []
+
+    # Get all signals
+    signals.append(["$u$"] + perievent_signal(
+        base.get_analogsignal(seg, 'Press/Release_0'), press, window))
+    signals.append(["$u_r$"] + perievent_signal(
+        base.get_analogsignal(seg, 'Press/Release_1'), press, window))
+    if with_x:
+        signals.append(["$x_1$"] + perievent_signal(
+            base.get_analogsignal(seg, 'Delay state_0'), press, window))
+        signals.append(["$x_2$"] + perievent_signal(
[…]

### 2013-11-14 scripts/ctrl_plot.py

@@ -1,0 +1,194 @@
+from glob import glob
+import os
+import random
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from . import base, files
+
+
+## Figure 9
+
+def signals(pklf, tr):
+    d = base.load_pickle(pklf)
+    if not d.has_key(tr + '_sig'):
+        return
+    sd = d[tr + '_sig']
+
+    plt.figure(1)
+    for ix in range(2):
+        plt.plot(sd['signals'][ix][1],
+                 sd['signals'][ix][2] - ix * 0.5,
+                 label=sd['signals'][ix][0])
+
+    plt.yticks(())
+    plt.xticks(sd['s_ticks'], sd['s_labels'])
+
+    if 'simple' in pklf and tr == 'c':
+        leg = plt.legend(loc='upper right', prop={'size':10}, frameon=True)
+        leg.get_frame().set_edgecolor('none')
+        plt.title("Cue-responding correct trial", fontsize='medium')
+    elif 'simple' in pklf:
+        plt.title("Cue-responding late trial", fontsize='medium')
+    elif 'adaptive' in pklf and tr == 'c':
+        plt.title("Adaptive control correct trial", fontsize='medium')
+    elif 'adaptive' in pklf:
+        plt.title("Adaptive control premature trial", fontsize='medium')
+
+    if 'adaptive' in pklf:
+        plt.gca().xaxis.get_major_ticks()[-2].set_pad(16)
+
+    plt.grid()
+    plt.gca().xaxis.set_ticks_position('none')
+
+    for side in ['left', 'top', 'right', 'bottom']:
+        plt.gca().spines[side].set_visible(False)
+
+    tight = {'rect': [0.05, -0.02, 0.95, 1.02]}
+    base.save_or_show(plt.gcf(), pklf, tr+'_signals', fs=(3, 1.5), tight=tight)
+
+
+def traj(pklf, tr):
+    d = base.load_pickle(pklf)
+    if not d.has_key(tr + '_sig'):
+        return
+    sd = d[tr + '_sig']
+
+    plt.figure(1)
+    plt.grid()
+    plt.axhline(0, color='0.2')
+    plt.axvline(0, color='0.2')
[…]

### 2013-11-14 scripts/dint.py

@@ -1,0 +1,92 @@
+from __future__ import absolute_import
+
+import os.path
+
+import numpy as np
+import quantities as pq
+from scipy import stats
+
+from .sim_io import get_seg
+from . import base
+
+
+def perievent_neural(seg, evts, window, binsize, stnames):
+    extend = 0.1 * pq.s
+    window_ext = (window[0] - extend, window[1] + extend)
+    bins = int((window_ext[1] - window_ext[0]) / binsize)
+    trim = int(extend / binsize)
+    sts = []
+
+    d = {k: [] for k in ['rates', 'sdfs']}
+
+    for st in seg.spiketrains:
+        for stname in stnames:
+            if st.name.startswith(stname):
+                sts.extend(st.perievent_slices(evts.times, window))
+                break
+
+    todelete = []
+    for ix, st in enumerate(sts):
+        sdf = st.gaussian_sdf(bins=bins)[trim:-trim]
+        if np.mean(sdf) < 1:  # filter out < 1 Hz spiketrains
+            todelete.append(ix)
+            continue
+        d['rates'].append(np.mean(sdf))
+        d['sdfs'].append(sdf)
+
+    if len(d['sdfs']) == 0:
+        raise Exception('No SDFs')
+
+    d['rates'] = np.array(d['rates'])
+    d['sdfs'] = stats.zscore(np.vstack(d['sdfs']), axis=1)
+    return d
+
+def analyze(path, comps=2, binsize=0.1*pq.s, window=(-4.0*pq.s, 4.1*pq.s)):
+    seg, e = get_seg(path)
+    d = {}
+    # fs contents:
+    #  dint, exptype, seed
+    fname = os.path.splitext(os.path.basename(path))[0]
+    fs = fname.split('-')
+    d['id'] = {'exp': fs[1], 'seed': fs[2]}
+    seq = fs[1].upper()
+    d['seq'] = 'cc' if seq == 'CC' else 'ec'
+
+    try:
+        evts = base.valid_presses(seg, e, seq)
+        evts[0]
+    except IndexError:
+        print base.merge_all_eas(seg, e.values())
+        raise ValueError("No valid presses in " + path)
+    evts.times += 0.3 * pq.s  # For delay to muscles
+
+    rand = np.random.RandomState(120)  # To keep analysis deterministic
+
+    for stname in ('dint', 'sint'):
+        # Only keep 174 neurons (as in experiment)
[…]

### 2013-11-14 scripts/dint_plot.py

@@ -1,0 +1,166 @@
+from glob import glob
+import os.path
+import random
+
+import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+from matplotlib.collections import LineCollection
+import numpy as np
+from scipy import stats
+
+from . import base, files
+
+## Figures 7 and 8
+
+def pcs(simf, expf, stname):
+    exp = base.load_pickle(expf)
+    sim = base.load_pickle(simf)
+    ids = sim['id']
+    name = '-'.join([ids['exp'], ids['seed']])
+
+    try:
+        comps = sim[stname]['evecs'].shape[1]
+    except:
+        if stname == 'sint':
+            return
+        raise
+    norm = plt.Normalize(vmin=-3, vmax=3)
+    w = 4.0
+    wr = [10] * comps
+    if 'lc' in simf:
+        wr.append(0.5)
+        w += 0.07 * w
+    if 'cc' in simf:
+        w += 0.125 * w
+
+    fig = plt.figure(1)
+    if stname == 'dint':
+        gs = GridSpec(3, len(wr), height_ratios=(1, 2, 2), width_ratios=wr)
+    elif stname == 'sint':
+        gs = GridSpec(2, len(wr), height_ratios=(1, 2), width_ratios=wr)
+
+    for comp in xrange(comps):
+        # Determine normal and flipped PR, use the right one
+        exp_ev = exp['evecs_%s' % sim['seq']].T[comp]
+        sim_ev = sim[stname]['evecs'].T[comp]
+        sim_sdf = sim[stname]['sdfs_%d' % comp]
+        pr, _ = stats.pearsonr(exp_ev, sim_ev)
+        if stats.pearsonr(exp_ev, sim_ev * -1)[0] > pr:
+            sim_ev *= -1
+            pr, _ = stats.pearsonr(exp_ev, sim_ev)
+            sim_sdf = sim_sdf[::-1]  # reverse SDFs
+        rsquared = pr ** 2
+        if stname == 'dint' and comp == 0:
+            print files.pkl(simf, True), 'R$^2$=%.2f ' % rsquared
+
+        # Top: PC
+        fig.add_subplot(gs[0, comp])
+        plt.plot(exp['t'], exp_ev, lw=2, c='r')
+        plt.plot(sim['t'], sim_ev, lw=2, c='k')
+        plt.axvline(0.0, ls=':', color='k')
[…]

### 2013-11-14 scripts/ds.py

@@ -1,0 +1,150 @@
+import random
+
+import numpy as np
+
+from . import base
+
+
+class SimpleRTSim(object):
+    def __init__(self, trial, E, R, beta, adaptive=True, dt=0.01):
+        self.trial = trial.lower()
+        self.dt = dt
+        self.E = E
+        self.R = R
+        self.beta = beta
+        self.adaptive = adaptive
+
+    @property
+    def dt(self):
+        return self._dt
+
+    @dt.setter
+    def dt(self, dt):
+        self._dt = dt
+        self.dt1 = dt * 3 * 0.5
+        self.dt2 = dt * 0.5
+
+    ## Constants
+
+    def cue(self, t):
+        if self.trial in 'cl':
+            return 1 if t > 2.0 and t <= 2.1 else 0
+        return 0
+
+    def lights(self, t):
+        if self.trial == 'p':
+            return 1 if t > 1.5 and t <= 3.5 else 0
+        elif self.trial == 'l':
+            return 1 if t > 2.6 and t <= 4.6 else 0
+        return 0
+
+    def press(self, t):
+        return 1 if t > 0.8 and t <= 1.0 else 0
+
+    def reward(self, t):
+        if self.trial == 'c':
+            return 1 if t > 2.6 and t <= 4.6 else 0
+        return 0
+
+    ## Dynamic
+
+    def d_lever(self, t, X, lever):
+        return 10 * (self.release(t, X, lever) - self.press(t))
+
+    def d_X(self, t, X):
+        dx1 = (2.5 * self.press(t)
+               - self.E * self.lights(t)
+               - self.R * self.reward(t) * X[0])
+        dx2 = self.beta * X[0] - self.R * self.reward(t) * X[1]
+        toret = np.array([dx1, dx2])
+        toret[toret * self.dt + X > 1] = 0
+        toret[toret * self.dt + X < -1] = 0
+        return toret
+
+    def release(self, t, X, lever):
+        if lever >= 1.0:
+            return 0
+
+        if self.trial == 'c':
+            ret = 1 if t > 2.0 and t <= 2.2 else 0
+        elif self.trial == 'p':
+            ret = 0 if t > 1.3 and t <= 1.5 else 0
+        elif self.trial == 'l':
[…]

### 2013-11-14 scripts/ds_plot.py

@@ -1,0 +1,299 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+from . import base, files
+
+
+## Figure 1
+def env_signals(pklf, trial):
+    d = base.load_pickle(pklf)
+
+    fig = plt.figure(1)
+    lu, = plt.plot(d['t'], d['u']+0.6, lw=2)
+    lc, = plt.plot(d['t'], d['u_c']+0.4, lw=2)
+    lr, = plt.plot(d['t'], d['u_r']+0.2, lw=2)
+    ll, = plt.plot(d['t'], d['u_l'], lw=2)
+    plt.xlim(right=5)
+    plt.ylim(-0.15, 1.75)
+
+    plt.yticks(())
+    if pklf[-5] == 'c':
+        plt.xticks((0, 1, 2, 2.6, 4.6),
+                   ('$t_s$', '$t_p$', '$t_c$', '$t_{Rw}$', '$t_{ITI}$'))
+    elif pklf[-5] == 'p':
+        plt.xticks((0, 1, 1.5, 3.5),
+                   ('$t_s$', '$t_p$', '$t_{TO}$', '$t_{ITI}$'))
+    elif pklf[-5] == 'l':
+        plt.xticks((0, 1, 2, 2.6, 4.6),
+                   ('$t_s$', '$t_p$', '$t_c$', '$t_{TO}$', '$t_{ITI}$'))
+
+    for side in ['left', 'top', 'right', 'bottom']:
+        plt.gca().spines[side].set_visible(False)
+    plt.gca().xaxis.set_ticks_position('none')
+    plt.grid()
+
+    figlegend = plt.figure(2)
+    figlegend.legend([lu, lc, lr, ll],
+                     ['Press', 'Cue', 'Reward', 'Timeout'],
+                     loc='center')
+
+    base.save_or_show(fig, pklf, trial + '_sig', fs=(2.75, 1))
+    base.save_or_show(figlegend, pklf, 'legend', fs=(1.25, 3))
+
+
+def fig1():
+    pad = 15
+    tr_w, tr_h = 209.74, 216.43
+    sig_w, sig_h = 2.75 * 72, 1 * 72
+    leg_w, leg_h = 1.25 * 72, 3.0 * 72
+    fsm_w, fsm_h = 497.74, 250.0
+    fig = base.svgfig(tr_w + sig_w + leg_w + pad, tr_h + fsm_h + pad*2)
[…]

### 2013-11-14 scripts/expt.py

@@ -1,0 +1,105 @@
+import numpy as np
+import quantities as pq
+from scipy import signal, stats
+from scipy.io import loadmat
+
+from .sim_io import get_seg
+from . import base
+
+
+def gaussian_sdf(hist, binsize, sigma=0.015 * pq.s, denoise=True):
+    # Put gaussian in terms of bins
+    sigma.units = binsize.units
+    sigma = float(sigma / binsize)
+    gauss = signal.gaussian(int(sigma * 7), sigma)  # +/- 3 stddevs
+    gauss /= np.sum(gauss)  # normalize to area 1 (important!)
+    gauss /= binsize  # then scale to size of bins
+    sdf = np.convolve(hist, gauss, mode='same')
+
+    if denoise:
+        sdf = (np.roll(sdf, -1) + sdf + np.roll(sdf, 1)) / 3.0
+
+    return sdf
+
+
+def perievent_neural_mat(mat, window, binsize):
+    d = {k: [] for k in ['sdfs_cc', 'rates_cc', 'sdfs_ec', 'rates_ec']}
+
+    for k in mat.keys():
+        if k.endswith('pC'):
+            seq = 'cc'
+        elif k.endswith('pE'):
+            seq = 'ec'
+        else:
+            continue
+
+        peh = mat[k].T
+        if peh.shape[1] > 80:
+            newpeh = np.zeros((peh.shape[0], 80))
+            scale = peh.shape[1] / 80
+            rnge = np.arange(0, peh.shape[1] + 1, scale)
+            rnge[-1] = peh.shape[1]
+            for i, s, e in zip(range(len(rnge)), rnge[:-1], rnge[1:]):
+                newpeh[:, i] = peh[:, s:e].sum(axis=1)
+            peh = newpeh
+
+        for hist in peh:
+            sdf = gaussian_sdf(hist, binsize)
+            if np.mean(sdf) < 1:  # filter out < 1 Hz spiketrains
+                continue
+            d['rates_%s' % seq].append(np.mean(sdf))
+            d['sdfs_%s' % seq].append(sdf)
+
+    for seq in ('cc', 'ec'):
+        sdf = 'sdfs_%s' % seq
+        d[sdf] = stats.zscore(np.vstack(d[sdf]), axis=1)[:, window[0]:window[1]]
+    return d
+
+
[…]

### 2013-11-14 scripts/files.py

@@ -1,0 +1,55 @@
+from glob import glob
+import os.path
+
+
+## Scripts
+home = os.path.expanduser("~")
+scripts = os.path.dirname(os.path.realpath(__file__))
+jneuro = os.path.realpath(os.path.join(scripts, '..'))
+data_sim = os.path.join(jneuro, 'data_sim')
+
+## Nengo
+nengo = os.path.join(home, 'nengo-latest')
+nengo_cl = os.path.join(nengo, 'nengo-cl')
+nengoscripts = os.path.join(nengo, 'trevor')
+nef_dint = os.path.join(nengoscripts, 'scripts', 'NEFDoubleIntegrator.py')
+nef_ctrl = os.path.join(nengoscripts, 'scripts', 'NEFControl.py')
+
+## Convert
+csv = glob(data_sim + '/*.csv')
+def h5(csv, basename=False):
+    if basename:
+        return os.path.basename(os.path.splitext(csv)[0] + '.h5')
+    return os.path.splitext(csv)[0] + '.h5'
+
+## Analysis
+exp_mat = os.path.join(jneuro, 'data_exp', 'PEHs_MFC_080312.mat')
+exp_nex = glob(os.path.join(jneuro, 'data_exp', '*.nex'))
+dintf = [os.path.splitext(f)[0] for f in glob(data_sim + '/dint-*')]
+ctrlf = [os.path.splitext(f)[0] for f in glob(data_sim + '/ctrl-*')]
+def pkl(h5, basename=False):
+    bname = os.path.splitext(os.path.basename(h5))[0]
+    if basename:
+        return bname + '.pkl'
+    return os.path.join(jneuro, 'analyzed', bname + '.pkl')
+
+
+## Plot
+def svg(pkl, suffix, basename=False):
+    bname = os.path.splitext(os.path.basename(pkl))[0]
+    if basename:
+        return bname + '_' + suffix + '.svg'
+    return os.path.join(jneuro, 'plots', bname + '_' + suffix + '.svg')
+
+
+## Combine
+def pdf(fname, basename=False):
+    toret = os.path.join(jneuro, 'figures', fname)
+    if basename:
+        return os.path.basename(toret)
+    return toret
+
+
+## Diagrams
+def diag(fname):
+    return os.path.join(jneuro, 'diagrams', fname + '.svg')

### 2013-11-14 scripts/nengocl.sh

@@ -1,4 +1,0 @@
-#!/bin/zsh
-NENGOCL=$HOME/nengo-latest/nengo-cl
-D=$HOME/nengo-latest/trevor/scripts
-$NENGOCL ${D}/$@

### 2013-11-14 scripts/neo.py

@@ -1,20 +1,19 @@
 """
-My tweaks to python-neo to make things easier.
+My tweaks to python-neo
 """
 from __future__ import absolute_import
 
-import quantities as pq
 import numpy as np
+import quantities as pq
 import scipy.signal
 
-from neo import Segment, AnalogSignal, SpikeTrain
-from neo import EventArray, EpochArray, Event
+from neo import Segment, AnalogSignal, SpikeTrain, EventArray, EpochArray, Event
 from neo.io.baseio import BaseIO
 from neo.io.neuroexplorerio import NeuroExplorerIO
 from neo.io.hdf5io import NeoHdf5IO
 
 
-### EpochArray additions
+### EpochArray
 
 def epa__eq__(self, other):
     return (np.all(self.times == other.times) and
@@ -68,7 +67,7 @@ def epa_merge_and_sort_epochs(epochs1, epochs2):
     return (map(lambda m: m[0], merged) * epochs1.times.units,
             map(lambda m: m[1], merged) * epochs1.durations.units,
             np.array(map(lambda m: m[2], merged), dtype='S'))
-EpochArray._merge_and_sort_epohcs = staticmethod(epa_merge_and_sort_epochs)
+EpochArray._merge_and_sort_epochs = staticmethod(epa_merge_and_sort_epochs)
 
 def epa_filter_for(self, seq, copy=True):
     """Filter the EpochArray so that only the seq `seq` remains."""
@@ -116,8 +115,8 @@ def epa_epochs_from(self, seq, name=None):
                                       dtype='S'))
 EpochArray.epochs_from = epa_epochs_from
 
+### EventArray
 
-### EventArray additions
 def eva__eq__(self, other):
     return (np.all(self.times == other.times) and
             np.all(self.labels == other.labels))
@@ -220,13 +219,6 @@ def eva_epochs_from(self, seq, name=None):
                                       dtype='S'))
 EventArray.epochs_from = eva_epochs_from
 
-def eva_during_epochs(self, epoch_array):
-    merged = EventArray()
-    for t, d in zip(epoch_array.times, epoch_array.durations):
-        merged += self.time_slice(t, t + d)
-    return merged
-EventArray.during_epochs = eva_during_epochs
-
 def eva_time_slice(self, t_start, t_stop):
     if len(self) == 0:
         return EventArray()
@@ -239,6 +231,12 @@ def eva_time_slice(self, t_start, t_stop):
[…]

### 2013-11-14 scripts/plot.py

@@ -1,399 +1,0 @@
-import matplotlib
-matplotlib.use('Agg')
-font = {'family': 'sans-serif',
-        'sans-serif': 'Arial'}
-matplotlib.rc('font', **font)
-matplotlib.rc('figure', dpi=300)
-
-import numpy as np
-import quantities as pq
-from scipy.stats import pearsonr
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from matplotlib.collections import LineCollection
-from matplotlib.patches import Polygon
-from matplotlib.colors import ListedColormap
-
-
-def save_or_show(path, ext='.svg', fig=1, save=True):
-    if save:
-        plt.savefig(path + ext, dpi=300)
-        print "Saved " + path + ext
-        plt.close(fig)
-    else:
-        plt.show()
-
-
-def plot_pc_summary(exp, sim, name='pc_summary', pos='left'):
-    print "Plotting PCs: %s" % name
-    comps = sim['evecs'].shape[1]  # Comps is always 2, but whatever
-    norm = plt.Normalize(vmin=-3, vmax=3)
-    w = 2 * comps
-    wr = [10] * comps
-    left = 0.025
-    right = 0.975
-    if pos == 'right':
-        wr += [1.0 / comps]
-        w += (wr[-1] / np.sum(wr)) * w
-        right -= (wr[-1] / np.sum(wr))
-    if pos == 'left':
-        w += 0.125 * w
-        left += 0.125
-
-    fig = plt.figure(figsize=(w, 6))
-    gs = gridspec.GridSpec(3, len(wr), height_ratios=(1, 2, 2), width_ratios=wr)
-
-    for comp in xrange(comps):
-        # Top: PC
-        ax = fig.add_subplot(gs[0, comp])
-
-        ax.plot(exp['t'], exp['evecs_%s' % sim['seq']].T[comp], lw=2, color='r')
-        ax.plot(sim['t'], sim['evecs'].T[comp], lw=2, color='k')
-        # ax.axhline(0, color='k')
-        ax.axvline(0.0, ls=':', color='k')
-        ax.axvline(1.0, ls=':', color='k')
-        ax.set_ylim((-2.5, 2.5))
-        ax.set_xlim((-4, 4))
-
-        # Legend with amount of variance explained
-        # ax.legend(("Expt, %.1f%% var" % (
-        #     exp['evals_%s' % sim['seq']][comp] * 100),
-        #            "Sim, %.1f%% var" % (sim['evals'][comp] * 100)),
[…]

### 2013-11-14 scripts/setup.sh

@@ -1,10 +1,0 @@
-#!/bin/zsh
-if [ ! -d "$HOME/nengo-latest" ]; then
-    echo "Moving nengo-* to nengo-latest"
-    mv $HOME/nengo-* $HOME/nengo-latest
-fi
-if [ ! -L "$HOME/nengo-latest/trevor" ]; then
-    echo "Linking scripts to /nengo-latest/trevor"
-    SCRIPTS=$(readlink -f ..)
-    ln -s $SCRIPTS $HOME/nengo-latest/trevor
-fi

### 2013-11-14 scripts/sim_io.py

@@ -1,6 +1,7 @@
 from __future__ import absolute_import
 
 import os
+import os.path
 import numpy as np
 import quantities as pq
 
@@ -125,7 +126,7 @@ class NengoIO(neo.BaseIO):
                             data[j] = [[] for _ in xrange(dims)]
                         elif t == "EventArray":
                             data[j] = [[] for _ in xrange(3)]
-                    # No continue, keep going
+                        # No continue, keep going
 
                 time = float(ll[0])
                 for j, lll in enumerate(ll):
@@ -136,11 +137,11 @@ class NengoIO(neo.BaseIO):
                             data[j][k].append(float(v))
                     elif types[j] == "EventArray":
                         curdata = float(lll)
-                        if curdata == 1.0 and curdata != lastdata[j]:
+                        if np.allclose(curdata, 1.0) and curdata != lastdata[j]:
                             data[j][0].append(time)
-                        elif curdata == 0.0 and curdata != lastdata[j]:
+                        elif np.allclose(curdata, 0.0) and curdata != lastdata[j]:
                             data[j][1].append(time)
-                        elif curdata == -1.0 and curdata != lastdata[j]:
+                        elif np.allclose(curdata, -1.0) and curdata != lastdata[j]:
                             data[j][2].append(time)
                         lastdata[j] = curdata
                     elif types[j] == "SpikeTrains":
@@ -167,17 +168,20 @@ class NengoIO(neo.BaseIO):
                                          name=col + '_' + str(i)))
             elif typ == "EventArray":
                 name = col[:-5] + 'on'
-                labels = np.array([name] * len(dat[0]), dtype='S')
                 seg.eventarrays.append(neo.EventArray(
-                    times=dat[0], labels=labels, channel_name=name))
+                    times=np.asarray(dat[0], dtype=np.float64) * pq.s,
+                    labels=np.array([name] * len(dat[0]), dtype='S'),
+                    channel_name=name))
                 name = col[:-5] + 'zero'
-                labels = np.array([name] * len(dat[0]), dtype='S')
                 seg.eventarrays.append(neo.EventArray(
-                    times=dat[1], labels=labels, channel_name=name))
+                    times=np.asarray(dat[1], dtype=np.float64) * pq.s,
+                    labels=np.array([name] * len(dat[0]), dtype='S'),
+                    channel_name=name))
                 name = col[:-5] + 'off'
[…]

### 2013-12-05 paper/jneurosci2013.tex
Removed figures and footnote from paper
In order to conform to JNeuroscience submission guidelines

@@
 
-\newcommand{\scalefigone}[3]{
-  \begin{figure}[ht!]
-    % Requires \usepackage{graphicx}
-    \centering
-    \includegraphics[width=#2\columnwidth]{#1}
-    \caption{#3}
-    \label{#1}
-  \end{figure}}
-
 \setlength\fboxrule{1pt}
@@
 
-\textbf{Pages:} 32. \textbf{Figures:} 10. \textbf{Tables:} 2. \textbf{Multimedia, 3D models:} 0
+\textbf{Pages:} 27. \textbf{Figures:} 10. \textbf{Tables:} 2. \textbf{Multimedia, 3D models:} 0
 
@@
 and is also penalized by a time-out with the house lights extinguished.
-Figure~\ref{fig1}A presents a schematic view of the task.
+Figure~1A presents a schematic view of the task.
 
@@
 the graphical depiction of
-the finite state machine in Figure~\ref{fig1}C.
+the finite state machine in Figure~1C.
 
@@
 the experimental principal components
-(shown in Figure~\ref{fig2}).
+(shown in Figure~2).
 
@@
 after correct trials.
-Figure~\ref{fig2}D indicates that
+Figure~2D indicates that
 the principal components return to their
@@
 the principal components
-in Figure~\ref{fig2}D are
+in Figure~2D are
 not at the origin point.
@@
 
-Finally, Figure~\ref{fig2}D shows that
+Finally, Figure~2D shows that
 the principal components are unstable
@@
 in the Nengo simulation environment
-\cite{BekolayInpress}.\footnote{The Nengo scripts and simulation
-  package simulating the networks and their environment are available
-  at \url{https://github.com/tbekolay/jneurosci2013/releases/tag/2013-11-29}.}
+\cite{BekolayInpress}.
 Using the NEF, each signal
@@
 between those populations
-(see Figure~\ref{fig3}).
+(see Figure~3).
 
@@
 contained 1200 adaptive leaky integrate-and-fire neurons
-(see Figure~\ref{fig3});
+(see Figure~3);
 174 of the neurons in an analyzed population
@@
 As previously described in \citetext{Narayanan2009}
-and summarized in Figure~\ref{fig2},
+and summarized in Figure~2,
 the first two principal components
@@
 during the post-correct trial
-(Figure~\ref{fig2}E).
+(Figure~2E).
 Since the first two principal components
 account for over half of the variance
-in the neural data (Figure~\ref{fig2}B),
+in the neural data (Figure~2B),
 we hypothesize that the mPFC
@@
 related to the two leading PCs,
-are shown for post-correct trials in Figure~\ref{fig2}F.
+are shown for post-correct trials in Figure~2F.
 Nearly equal fractions of neurons
@@
 Importantly, the trials used for this analysis,
-and shown in Figure~\ref{fig2}, were all correct trials.
+and shown in Figure~2, were all correct trials.
 The only difference between trial types
@@
     Appropriate values were found through simulation;
-    see Figure~\ref{fig4}.}
+    see Figure~4.}
   \label{tab:sim-param}
@@
 is to predict the time at which the cue occurs.
-Figure~\ref{fig4} shows that this is possible.
+Figure~4 shows that this is possible.
 The parameter $\beta$ controls how fast
@@
 resulting in different states at $t_c$
-(see Figure~\ref{fig4}B);
+(see Figure~4B);
 since this parameter is encoded in the strength
@@
 values at $t_c$ for $\beta=0.44$ are shown
-in Figure~\ref{fig4}A.
+in Figure~4A.
 
@@
 For sufficiently high values of $R$
-(see Figure~\ref{fig4}C),
+(see Figure~4C),
 the system starts the next trial
@@
 For sufficiently high values of $E$
-(see Figure~\ref{fig4}D),
+(see Figure~4D),
 the system starts the next trial near $(-1, -1)$,
 resulting in the system being unable to predict
-the time of the cue (see Figure~\ref{fig4}A),
+the time of the cue (see Figure~4A),
 and therefore employing the cue-responding strategy.
@@
 when the predicted cue approaches.
-Figure~\ref{fig5} shows a representative
+Figure~5 shows a representative
 correct trial from the cue-responding
@@
 following the error.
-Importantly, Figure~\ref{fig5}B
+Importantly, Figure~5B
 shows the adaptive control network
@@
 the double integrator model as well,
-as shown in Figure~\ref{fig6}.
+as shown in Figure~6.
 The only difference that cannot
@@
 
-Figure~\ref{fig7} compares
+Figure~7 compares
 the two leading principal components
@@
 the double integrator network.
-Figure~\ref{fig8}
+Figure~8
 shows the results of the analysis
@@
 
-Figure~\ref{fig9} shows that
+Figure~9 shows that
 the decoded values of spiking neural networks
@@
 in the theoretical results,
-as can be seen in Figure~\ref{fig9}A.
+as can be seen in Figure~9A.
 
@@
 
-Figure~\ref{fig10} shows the performance
+Figure~10 shows the performance
 of the experimental subjects,
@@
 despite all but one performing the task well
-(see Figure~\ref{fig10}).
+(see Figure~10).
 Also notable is that the model
@@
 temporal difference learning.
-As shown in Figure~\ref{fig6},
+As shown in Figure~6,
 the positive and negative reward prediction error
@@
 of the simple RT task
-(see Figures~\ref{fig7} and \ref{fig8}).
+(see Figures~7 and 8).
 Since we find neurons that are sensitive
@@
 
-\clearpage
-\section{Figures}
-
-\scalefigone{fig1}{1.0}{}
-
-\scalefigone{fig2}{1.0}{}
-
-\scalefigone{fig3}{0.8}{}
-
-\scalefigone{fig4}{1.0}{}
-
-\scalefigone{fig5}{1.0}{}
-
-\scalefigone{fig6}{0.5}{}
-
-\scalefigone{fig7}{1.0}{}
-
-\scalefigone{fig8}{1.0}{}
-
-\scalefigone{fig9}{0.5}{}
-
-\scalefigone{fig10}{1.0}{}
-
 \end{document}
